---
editor_options:
  markdown:
    wrap: 100
---

# Import

## Load libraries

```{r libraries, message=FALSE, warning=FALSE}
# General
library(ragg) # graphics device
library(here) # for file access
library(haven) # for importing .xpt

# Statistics
library(matrixStats) # used for weighted variance calculation
library(emmeans) # for pairwise comparisons
library(mgcv) # used to train model
library(survey) # to account for survey design
library(survival) # for Cox regressions

# Visualization
library(kableExtra) # for pretty tables
library(khroma) # for Paul Tol color schemes
library(ggthemes) # for theme_few()
library(ggtext) # for element_markdown()
library(ggh4x) # for element_rect()
library(ggridges) # for geom_density_ridges()
library(ggpubr) # for ggarrange()
library(cowplot) # for combined plots
library(sameplot) # for plotting consistently

# Data operations
library(flexsurv) # for calculating phenoage
library(data.table) # for speedy dataframes
library(tidyverse) # for general data operations
library(tidytable) # for fast data loading

# Multithreading
library(future)
library(future.apply)
```

## Import scripts

```{r get_helpers}
source(here("scripts", "helpers", "helpers.R"))
source(here("scripts", "bioage_package", "hd_calc.R"))
source(here("scripts", "bioage_package", "kdm_calc.R"))
source(here("scripts", "bioage_package", "phenoage_calc.R"))
```

## Print params for knitted reports

```{r print_params}
kable(
  as.data.frame(
    unclass(params)
  )
)
```

## Unpack covariates list

We take the param value for covariates and build one lists with all included interactions with age
and one with just the variables.

```{r covariate_list}
covariate_interactions <- trimws(
  strsplit(params$covariates, "_")[[1]]
)

covariate_vars <- covariate_interactions[!grepl(",", covariate_interactions)]
```

## Import data

### NHANES4

```{r import_data}
import_nhanes_data <- function(data_path) {
  files <- list.files(
    data_path,
    pattern = "\\.xpt$",
    full.names = TRUE,
    ignore.case = TRUE
  )

  tibble(
    dataset_id = tools::file_path_sans_ext(basename(files)),
    cycle = dplyr::case_when(
    str_detect(dataset_id, regex("^P_")) ~ "2017_2020",
    str_detect(dataset_id, regex("_B$")) ~ "2001_2002",
    str_detect(dataset_id, regex("_C$")) ~ "2003_2004",
    str_detect(dataset_id, regex("_D$")) ~ "2005_2006",
    str_detect(dataset_id, regex("_F$")) ~ "2009_2010",
    str_detect(dataset_id, regex("_I$")) ~ "2015_2016",
    TRUE ~ "1999_2000"
  ),
    data = map(
      files, ~ read_xpt(.x)
    )
  )
}

df_raw <- import_nhanes_data(here("data", "nhanes4"))
```

### Import NHANES4 Mortality data

```{r}
import_nhanes4_mortality <- function(data_path) {

  files <- list.files(
    data_path,
    pattern = "\\.dat$",
    full.names = TRUE,
    ignore.case = TRUE
  )
  
tibble(
    dataset_id = tools::file_path_sans_ext(basename(files)),
    cycle = str_extract(files, "\\d{4}_\\d{4}"),
    data = map2(files, dataset_id, function(dat, id) {

              read_fwf(
          file = dat,
          col_types = "iiiiiiii",
          fwf_cols(
            SEQN = c(1, 6),
            ELIGSTAT = c(15, 15),
            MORTSTAT = c(16, 16),
            UCOD_LEADING = c(17, 19),
            DIABETES = c(20, 20),
            HYPERTEN = c(21, 21),
            PERMTH_INT = c(43, 45),
            PERMTH_EXM = c(46, 48)
          ),
          na = c("", ".")
        ) %>% tibble() 
      })
  )
}

df_raw <- rbind(df_raw, import_nhanes4_mortality(here("data", "nhanes4")))
```

### NHANES3

#### Initial import

Below is the function to import NHANES 3 data from the .dat and .SAS files. Because the files are
large and importing is slow, it's run only once and raw data files are not stored in the repository
but can be downloaded from <https://wwwn.cdc.gov/nchs/nhanes/nhanes3/datafiles.aspx#core>.

The mortality data (<https://stacks.cdc.gov/view/cdc/117141>) also has a .SAS file in the
documentation, but this is intended for NHIS identifiers, not NHANES variables. Sample code by the
CDC in "R_ReadInProgramAllSurveys.R" is used to parse the mortality dataset.

```{r import_train_data_orig}
# import_nhanes3_data <- function(data_path) {
#   library(SAScii)
# 
#   files <- list.files(
#     data_path,
#     pattern = "\\.dat$",
#     full.names = TRUE,
#     ignore.case = TRUE
#   )
# 
#   tibble(
#     dataset_id = tools::file_path_sans_ext(basename(files)),
#     cycle = "1988_1994",
#     data = map2(files, dataset_id, function(dat, id) {
#       if (grepl("NHANES_III_MORT_2019_PUBLIC", id)) {
#         read_fwf(
#           file = dat,
#           col_types = "iiiiiiii",
#           fwf_cols(
#             SEQN = c(1, 6),
#             ELIGSTAT = c(15, 15),
#             MORTSTAT = c(16, 16),
#             UCOD_LEADING = c(17, 19),
#             DIABETES = c(20, 20),
#             HYPERTEN = c(21, 21),
#             PERMTH_INT = c(43, 45),
#             PERMTH_EXM = c(46, 48)
#           ),
#           na = c("", ".")
#         )
#       } else {
#       sas <- here(data_path, paste0(id, ".sas"))
#       read.SAScii(dat, sas)
#       }
#     })
#   )
# }
# 
# df_train_raw <- import_nhanes3_data(here("data", "nhanes3"))
# 
# saveRDS(
#      object = df_train_raw,
#      file = here("data", "nhanes3", "nhanes3.RDS"),
#      compress = "xz"
#    )
```

#### Import

```{r import_train_data}
df_train_raw <- readRDS(here("data", "nhanes3", "nhanes3.RDS"))
```
