---
editor_options:
  markdown:
    wrap: 100
---

# Data preprocessing

## Data selection & standardization

We make a list of files, store the year and filename from each file, then append the data to the
row. Next, we take all the files from one year and join the rows if they match or columns if they
don't match according to SEQN (ID) and create a `year` column and `cohort` column that takes the
first 4 characters from `year`. `imap_dfr()` applies this to each year (map), returns the index (i)
and row-binds into a dataframe (dfr).

### Select audiometry sample

We select the IDs in the 'Audiometry' files, then open the df from each file, return the df if the
file contains "Audiometry" otherwise, filter the IDs to the list of audiometry IDs.

```{r select_audiometry_sample}
select_audiometry_sample <- function(data_map) {
  # Extract all subject IDs from audiometry files
  audiometry_sample_ids <- data_map %>%
    filter(str_detect(dataset_id, "AUX")) %>%
    pull(data) %>%
    map(~ pull(.x, SEQN)) %>%
    reduce(union)

  # Filter non-audiometry files to audiometry subject IDs only
  data_map %>%
    mutate(
      data = map2(data, dataset_id, \(df, f) {
        if (str_detect(f, "AUX")) {
          df
        } else {
          filter(df, SEQN %in% audiometry_sample_ids)
        }
      })
    )
}
```

### Select and rename columns

We intentionally do not select creat_umol columns if we have to restandardize them later (1999 and
2005)

```{r select_cols}
select_columns <- function(data_map, column_path) {
  col_spec <- read.csv(column_path)

  col_spec <- col_spec %>%
    filter(include == TRUE) %>%
    select(label, variable, cycles) %>%
    separate_longer_delim(cycles, delim = ", ")

  # Merge data from the same year into one dataframe
  data_map <- data_map %>%
    group_by(cycle) %>%
    summarise(
      data = list(reduce(data, full_join, by = "SEQN")),
      .groups = "drop"
    )
  
    # Spec variables that aren't present in their cycle's merged data
  available <- data_map %>%
    transmute(cycle, cols = map(data, names)) %>%
    unnest_longer(cols)

  missing <- col_spec %>%
    anti_join(available, by = c("cycles" = "cycle", "variable" = "cols"))

  if (nrow(missing) > 0) {
    stop(
      "Spec variables missing from their cycle:\n",
      paste0("  ", missing$cycles, " : ", missing$variable, collapse = "\n"),
      call. = FALSE
    )
  }
  
  # Match variables with spec and rename to label. Per-cycle columns are obtained and named.
  data_map %>%
    mutate(data = map2(data, cycle, function(dat, cy) {
      keep <- col_spec %>% filter(cycles == cy)
      keep <- setNames(keep$variable, keep$label)
      keep <- c(SEQN = "SEQN", keep[keep != "SEQN"])

      dat %>%
      select(any_of(keep)) %>%
      mutate(cycle = cy)
    }))
  
}
```

### Merge dataframe and create ID vars

```{r merge_cycles}
merge_data <- function(data_map) {
  data_map %>%
    pull(data) %>%
    bind_rows() %>%
    relocate(SEQN, cycle)
}
```

### Standardize values

CDC recommends to use 4-year weights for 1999-2000 and 2001-2002, 3.2-year weights for 2017-2020
March, and 2-year weights for the others. Because different cycles collected audiometry data in
different age populations, we cannot standardize the weights simply by total number of years (15.2)
in the sample, but must standardize each age group by the number of years. Before 2005, audiometry
was performed in only half a sample, but this is corrected for already by the audiometry-specific
weights variable in those earlier years. In this study, people under 20 years old are excluded.
Audiometry weights in 1999-2000 and 2001-2002 contain NAs for missing data, which is set to 0 in the
corresponding MEC examination weight variable, so missing weights are set to 0 to match this. The
following ages were assessed in each cycle:

- 1999-2000: 20-69
- 2001-2002: 20-69
- 2003-2004: 20-69
- 2005-2006: 12-19, 70+ (above 84 is set to 84)
- 2009-2010: 12-19, 70+ (above 79 is set to 80)
- 2015-2016: 20-69
- 2017-2020 (up to March): 6-19, 70+ (above 79 is set to 80)

To calculate weights, we divide the number of years in the weights by the total number of years in
the age group.

We censor mortality data above 10 years, so the follow-up period is effectively capped at 10 years
(in line with the Levine paper).

Ages are binned into age groups. Mexican Americans (`race == 1`) and other hispanics (`race == 2`)
are merged into a single hispanics variable. BP is averaged over multiple measurements.

- CRP was measured in mg/L instead of mg/dL in 2015_2016 and 2017_2020 and was converted to mg/dL.
- Creatinine is in mg/dL (and must be used in mg/dL for correction/standardization) and is converted
  to umol/L after standardization to align with original PhenoAge paper. Standardized in 1999_2000,
  2005_2006, and 2017_2020.
- Albumin was converted from g/dL to g/L to align with PhenoAge paper and was backwards standardized
  in 2017_2020.
- The following others were standardized in 2017_2020 (due to new chemistry analyzer):
  - Alkaline phosphatase
  - Aspartate aminotransferase (additionally log-transformed, and no values are below 1 so no
    constant of one needed)
  - Alanine transaminase (additionally log-transformed, and no values are below 1 so no constant of
    one needed)
  - Blood urea nitrogen
  - Gamma glutamyl transpeptidase (additionally log-transformed, and no values are below 1 so no
    constant of one needed)
  - Iron
  - Lactate dehydrogenase
  - Uric acid

Creatinine in 1999 and 2005 and CRP in 2015 and 2017 are standardized. S

```{r standardize_vals}
standardize_values <- function(data) {
  data <- data %>%
    mutate(
      age_eligibility_group = case_when(
        age >= 6 & age < 12 ~ "6-11",
        (age >= 12 & age < 20) | age >= 70 ~ "12-19, 70+",
        age >= 20 & age < 70 ~ "20-69"
      ),
      cycle_years = case_when(
        cycle %in% c("1999_2000", "2001_2002") ~ 4,
        cycle == "2017_2020" ~ 3.2,
        TRUE ~ 2
      ),
      total_years_for_group = case_when(
        age_eligibility_group == "6-11" ~ 3.2,
        age_eligibility_group == "12-19, 70+" ~ 7.2,
        age_eligibility_group == "20-69" ~ 8
      ),
      weights = weights * (cycle_years / total_years_for_group)
    ) %>%
    select(
      -age_eligibility_group,
      -cycle_years,
      -total_years_for_group
    )

  data %>%
    mutate(
      age_group = case_when(
        age < 20 ~ "19-",
        age >= 20 & age < 30 ~ "20-30",
        age >= 30 & age < 40 ~ "30-40",
        age >= 40 & age < 50 ~ "40-50",
        age >= 50 & age < 60 ~ "50-60",
        age >= 60 & age < 70 ~ "60-70",
        isFALSE(params$telolength) & age >= 70 & age < 80 ~ "70-80",
        isFALSE(params$telolength) & age >= 80 ~ "80+"
      ),
        status = case_when(
      time > 120 ~ 0,
      TRUE ~ status
    ),
    time = case_when(
      time > 120 ~ 120,
      TRUE ~ time
    ),
      race = case_when(
        race == 2 ~ 1,
        TRUE ~ race
      ),
      alb = case_when(
        cycle == "2017_2020" ~ 1.044 * alb + 0.01128,
        TRUE ~ alb
      ),
      alb = alb * 10,
      alp = case_when(
        cycle == "2017_2020" ~ 1.001 * alp + 0.04294,
        TRUE ~ alp
      ),
      ast = case_when(
        cycle == "2017_2020" ~ 1.018 * ast + 3.762,
        TRUE ~ ast
      ),      
      ast_log = log(ast),
      alt = case_when(
        cycle == "2017_2020" ~ 1.013 * alt + 2.688,
        TRUE ~ alt
      ),    
      alt_log = log(alt),
      bun = case_when(
        cycle == "2017_2020" ~ 1.001 * bun + 0.4488,
        TRUE ~ bun
      ),    
      ggt = case_when(
        cycle == "2017_2020" ~ 0.8042 * ggt + 2.363,
        TRUE ~ ggt
      ),    
      ggt_log = log(ggt),
      iron = case_when(
        cycle == "2017_2020" ~ 0.9776 * iron - 4.494,
        TRUE ~ iron
      ),  
      ld = case_when(
        cycle == "2017_2020" ~ 0.8568 * ld + 2.062,
        TRUE ~ ld
      ),  
      ua = case_when(
        cycle == "2017_2020" ~ 0.9323 * ua + 0.2326,
        TRUE ~ ua
      ),  
      creat = case_when(
        cycle == "1999_2000" ~ 1.013 * creat + 0.147,
        cycle == "2005_2006" ~ -0.016 + 0.978 * creat,
        cycle == "2017_2020" ~ 1.051 * creat - 0.06945,
        TRUE ~ creat
      ),
      creat = creat * 88.4,
      crp = case_when(
        cycle %in% c("2015_2016", "2017_2020") ~ crp / 10,
        TRUE ~ crp
      ),
      crp_log = log(crp),
      telolength_log = log(telolength)
    ) %>%
    rowwise() %>%
    mutate(
      sbp = case_when(
        cycle == "2017_2020" ~ mean(c(sbp1, sbp2, sbp3), na.rm = TRUE),
        TRUE ~ mean(c(sbp1, sbp2, sbp3, sbp4), na.rm = TRUE)
      )
    ) %>%
    ungroup() %>%
    select(
      -sbp1,
      -sbp2,
      -sbp3,
      -sbp4
    )
}
```

### Calculate hearing metrics

Any values of "666" or "888" in audiometry are set to NA so that they're not used for calculating
the audiometric compound measures.

```{r calc_metrics}
calculate_hearing <- function(data) {
  audiometry_cols <- data %>%
    select(contains("audio")) %>%
    colnames()

  data %>%
    mutate(
      across(contains("audio"), ~ if_else(. %in% c(666, 888), NA_real_, as.numeric(.)))
    ) %>%
    rowwise() %>%
    mutate(
      test_retest_diff_left = abs(audio_left_1000 - audio_left_1000_retest),
      test_retest_diff_right = abs(audio_right_1000 - audio_right_1000_retest),
      audio_right_1000_mean = mean(c(audio_right_1000, audio_right_1000_retest)),
      audio_left_1000_mean = mean(c(audio_left_1000, audio_left_1000_retest)),
      pta_lf_r = mean(c(audio_right_500, audio_right_1000_mean, audio_right_2000)),
      pta_lf_l = mean(c(audio_left_500, audio_left_1000_mean, audio_left_2000)),
      pta_lf = mean(c(pta_lf_r, pta_lf_l)),
      pta_hf_r = mean(c(audio_right_4000, audio_right_6000, audio_right_8000)),
      pta_hf_l = mean(c(audio_left_4000, audio_left_6000, audio_left_8000)),
      pta_hf = mean(c(pta_hf_r, pta_hf_l))
    ) %>%
    ungroup()
}
```

### Data preprocessing pipeline

#### NHANES 4

1.  Only IDs in the audiometry sample of NHANES are retained.
2.  Only columns that are present in the specs.csv sheet are retained and data files from the same
    year are combined.
3.  Columns are renamed according to specs.csv.
4.  Data from different years is combined.
5.  Weights are standardized, an age group (10-year bins) column is made, and data is standardized
    according to NHANES guidelines.
6.  Hearing metrics used in this study are calculated.

```{r preprocess_pipe, message=FALSE, warning=FALSE, include=FALSE}
df_unfiltered <- df_raw %>%
  select_audiometry_sample() %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes4.csv")
  ) %>%
  merge_data() %>%
  standardize_values() %>%
  calculate_hearing()
```

#### NHANES 3

Training is only used for biological age algorithms and there's only a single cycle so no
standardization or audiometry subsampling required. Filtering will be performed in the biological
age calculation chunk.

```{r preprocess_pipe_train}
df_train <- df_train_raw %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes3.csv")
  ) %>%
  merge_data() %>%
  mutate(
    status = case_when(
      time > 120 ~ 0,
      TRUE ~ status
    ),
    time = case_when(
      time > 120 ~ 120,
      TRUE ~ time
    ),
    across(where(is.numeric),
                \(x) if_else(x %in% c(888L, 8888L, 88888L, 888888L), NA_real_, x)),
         crp_log = log(crp),
         creat = 0.96 * creat - 0.184,
    
    )
```

#### NHANES 4 HD Train

```{r}
df_train_hd <- df_raw %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes4.csv")
  ) %>%
  merge_data() %>%
  standardize_values() 
```

## Survey design

### Check completeness

Make a scrollable table and only show number of data present in ID, weights, and columns that have
no values at all.

```{r check_complete}
construct_table <- function(input, title) {
  input %>%
    kable(
      "html",
      caption = title
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}

df_unfiltered %>%
  group_by(cycle) %>%
  summarise(across(everything(), ~ sum(!is.na(.)), .names = "n_{.col}")) %>%
  select(cycle, n_SEQN, n_weights, where(~ any(. == 0))) %>%
  construct_table(title = "Check present values")
```

Telomere length was only assessed in early cohorts. Pregnancy is missing in 2009 and 2017 because it
was only assessed in women aged 20-44 and the audiometry sample only includes people aged 12-19 and
70+. Some IDs lack all audiometry data from 1999 and 2001. Because we have no data at all in these
subjects, we must exclude them. NHANES guidelines state that \< 10% missing is acceptable without
further adjustment needed.

### Exclude missing rows (no data at all)

```{r exclude_missing}
filter_data <- function(data) {
  verbose_filter <- function(data, ..., label) {
    start_n <- nrow(data)
    data_filtered <- data %>% filter(...)
    message(nrow(data_filtered), " remaining. ", start_n - nrow(data_filtered), "/", start_n, " excluded due to ", label)

    data_filtered
  }

  data <- data %>%
    verbose_filter(
      !is.na(weights),
      label = "missing weights"
    )
}

df_no_missing <- df_unfiltered %>% filter_data()
```

### Create survey design

We use the Survey package to account for the full survey design (weights, strata, PSUs)

```{r calc_descriptives}
design_no_missing <- svydesign(
  ids     = ~psu,
  strata  = ~stratum,
  weights = ~weights,
  data    = df_no_missing,
  nest    = TRUE
)
```

## Data filtering

We keep any columns we may need at this stage or have already used, such as the individual
audiometry columns. The design is stored intermittently in the pipe before any `if_any` statements
are used, because these don't work on the design object, only on the `design$variable` within the
design object, so in those cases, we store the design prior to the statement to ensure that
`design$variable` refers to the filtered `design`, set filter criteria on `design$variable`, which
is a dataframe/tidytable, and pass on a list of boolean values to select based on `design$variables`
to subset `design` on.

1.  Ages of 85 or higher before 2007 and 80 or higher after 2007 are cut off to 85/80, respectively,
    so we exclude these ages because we don't know the ages of the subjects. Because 2005-2006 is
    the only cycle with people aged 80-84 and because there are no people of "other" races, we
    excluded all people above 80. This also avoids weighting issues of the 70+ group in earlier and
    later cycles.
2.  People under 20 years of age aren't included in this study.
3.  Pregnant women are excluded (NAs are explicitly kept).
4.  Subjects with missing audiometry data are excluded.
5.  Subjects with large test-retest differences are excluded, as this suggests unreliable
    measurement.
6.  Subjects with missing or unknown covariate data are excluded.
7.  Subjects with missing biomarker data are excluded (as these are required for calculating
    biological ages).
8.  Subjects with any biomarker values \> 5 SDs above the sex-specific mean are excluded to avoid
    extreme values dominating the models.

### Filter data

```{r filter_df}
subset_data <- function(design, spec_path) {
  verbose_subset <- function(design, ..., label) {
    start_n <- nrow(design)
    design_subset <- design %>% subset(...)
    message(nrow(design_subset), " remaining. ", start_n - nrow(design_subset), "/", start_n, " excluded due to ", label)

    design_subset
  }

  audiometry_cols <- design$variables %>%
    select(contains("audio")) %>%
    colnames()

  col_spec <- read.csv(spec_path)

  biomarkers <- col_spec %>%
    filter(
      vartype == "biomarker", 
      include == TRUE, 
      !(label %in% c("sbp1", "sbp2", "sbp3", "sbp4", "crp", "ast", "alt", "ggt"))
      ) %>%
    pull(label)
  
  biomarkers <- unique(c(biomarkers, "sbp", "crp_log", "ast_log", "alt_log","ggt_log"))

  design <- design %>%
    verbose_subset(
      age < 80,
      label = "too old age."
    ) %>%
    verbose_subset(
      age >= 20,
      label = "too young age."
    ) %>%
    verbose_subset(
      pregnant != 1 | is.na(pregnant),
      label = "pregnancy."
    )

  design <- design %>%
    verbose_subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(audiometry_cols), ~ . == 888 | . == 666 | is.na(.))) %>%
        pull(keep),
      label = "missing audiometry data."
    ) %>%
    verbose_subset(
      test_retest_diff_left < 10 & test_retest_diff_right < 10,
      label = "too large test-retest difference."
    ) %>%
    verbose_subset(
      smoking %in% c(1, 2),
      label = "smoking refused or 'don't know'."
    ) %>%
    verbose_subset(
      !(education %in% c(7, 9)),
      label = "education refused or 'don't know'."
    )

  design <- design %>%
    verbose_subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(biomarkers), ~ is.na(.))) %>%
        pull(keep),
      label = "missing biomarkers."
    )

  design %>%
    verbose_subset(
      design$variables %>%
        group_by(sex) %>%
        mutate(keep = !if_any(
          all_of(biomarkers),
          ~ {
            m <- mean(.x, na.rm = TRUE)
            s <- sd(.x, na.rm = TRUE)
            abs(.x - m) > 5 * s
          }
        )) %>%
        ungroup() %>%
        pull(keep),
      label = "biomarker outlier (greater or smaller than 5 SDs)."
    )
}

design <- design_no_missing %>% subset_data(
  spec_path = here("specs", "col_spec_nhanes4.csv")
)
```

### Optionally only include telolength

```{r filt_telolength}
if (isTRUE(params$telolength)) {
  design <- design %>%
    subset(
      !is.na(telolength)
    )
}
```

### Check data issues (NHANES3 is unweighted)

```{r check_df_balance}
check_biomarker_standardization <- function(design, training_set, testing_spec_path, training_spec_path) {
  col_spec <- read.csv(testing_spec_path)

  biomarkers <- col_spec %>%
    filter(
      vartype == "biomarker", 
      include == TRUE, 
      !(label %in% c("sbp1", "sbp2", "sbp3", "sbp4", "crp", "ast", "alt", "ggt"))
      ) %>%
    pull(label)
  
  biomarkers <- unique(c(biomarkers, "crp_log", "ast_log", "alt_log","ggt_log"))

  all_vars <- c("age", "pta_hf", "pta_lf","sbp", biomarkers)
cycles <- unique(design$variables$cycle)

nhanes4 <- map_dfr(cycles, function(cyc) {
  des_cyc <- subset(design, cycle == cyc)

  stats <- map(all_vars, function(v) {
    fmla_v <- as.formula(paste("~", v))
    m <- as.numeric(svymean(fmla_v, des_cyc, na.rm = TRUE))
    s <- sqrt(as.numeric(svyvar(fmla_v, des_cyc, na.rm = TRUE)))
    paste0(round(m, 2), "(", round(s, 2), ")")
  }) %>%
    set_names(all_vars) %>%
    as_tibble()

  bind_cols(tibble(cycle = cyc), stats)
})

training_set[training_set == 88888] <- NA
training_set[training_set == 8888] <- NA
training_set[training_set == 888] <- NA

col_spec_train <- read.csv(training_spec_path)

biomarkers_train <- col_spec_train %>%
  filter(vartype == "biomarker",
         include == TRUE,
         label != "crp") %>%
  pull(label)
  all_vars_train <- c("age", "crp_log", biomarkers_train)


  nhanes3 <- map(all_vars_train, function(v) {
    m <- as.numeric(mean(training_set[[v]], na.rm = TRUE))
    s <- as.numeric(sd(training_set[[v]], na.rm = TRUE))
    paste0(round(m, 2), "(", round(s, 2), ")")
  }) %>%
    set_names(all_vars_train) %>%
    as_tibble() %>%
    mutate(cycle = "1988_1994")

  bind_rows(nhanes4, nhanes3) %>%
  arrange(cycle) %>%
  construct_table(title = "Check cohort age and biomarker balance")
}

design %>% check_biomarker_standardization(
  training_set = df_train,
  testing_spec_path = here("specs", "col_spec_nhanes4.csv"),
  training_spec_path = here("specs", "col_spec_nhanes3.csv")
)
```

ALP was higher in NHANES3, but could be in prat due to survey design (old people weigh too heavily)
and large SD. Creatinine is also a bit higher in NHANES3 but could be due to the same reason.

#### Verify normal distribution

```{r}
check_distribution <- function(data, spec_path, dataset) {
  col_spec <- read.csv(spec_path)
  biomarkers <- col_spec %>%
    filter(
      vartype == "biomarker", 
      include == TRUE, 
      !(label %in% c("sbp1", "sbp2", "sbp3", "sbp4", "crp", "ast", "alt", "ggt"))
      ) %>%
    pull(label)
  
  if (dataset == "NHANES4") {
      biomarkers <- unique(c(biomarkers, "sbp", "crp_log", "ast_log", "alt_log","ggt_log"))
  } else {
      biomarkers <- unique(c(biomarkers, "sbp", "crp_log"))
  }
  
  data %>%
    select(all_of(biomarkers)) %>%
    pivot_longer(names_to = "biomarker", values_to = "value") %>%
    ggplot(aes(x = value)) +
  geom_histogram(
    aes(y = ..density..),
    fill = "tomato",
    color = "white",
    alpha = 0.7
  ) +
  geom_density(
    color = "darkred",
    linewidth = 1,
    adjust = 3
  ) +
  facet_wrap(~ biomarker, scales = "free") +
  theme_few() +
  theme(
    axis.text = element_blank()
  )
  
}

sameplot(
design$variables %>% check_distribution(
     spec_path = here("specs", "col_spec_nhanes4.csv"),
     dataset = "NHANES4"
   ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "biomarker_distributions_NHANES4.png"
  )
)

sameplot(
df_train %>% check_distribution(
     spec_path = here("specs", "col_spec_nhanes3.csv"),
     dataset = "NHANES3"
   ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "biomarker_distributions_NHANES3.png"
  )
)
```

There is some kurtosis for several variables in NHANES3, but that should not be an issue.

## Calculate biological age

```{r calc_bioage, warning=FALSE}
insert_bioage <- function(design, train_data, train_data_hd, orig_phenoage = TRUE) {
  bioage_calc <- function(data, train_data, train_data_hd, orig_phenoage = TRUE) {
    biomarkers_phenoage <- c(
      "alb", "alp", "crp_log", "creat", "lymph", "mcv", "wbc", "rdw", "glu"
    )
    biomarkers_kdm <- c(
      "alb", "creat", "crp_log", "mono", "alp", "hba1c", "totchol", "bun", "sbp", "rbc"
    )
    biomarkers_hd <- c(
      "alb", "alp", "alt_log", "ast_log", "baso", "cal", "chlor", "ggt_log", "hdl", "hg", "htc", "iron", "ld", "lymph", "mch", "mchc", "mono", "neut", "phos", "pot", "rbc", "rdw", "sod", "totchol", "tp", "ua", "wbc"
    )

    # Calculate KDM
    train_fem_kdm <- kdm_calc(
      data = train_data %>%
        filter(
          sex == 2 & 
            age >= 20 & 
            age <= 79 & 
            (pregnant != 1 | is.na(pregnant))
          ),
      biomarkers = biomarkers_kdm,
      fit = NULL,
      s_ba2 = NULL,
      n_allowed_missing_biomarkers = 0
    )

    train_male_kdm <- kdm_calc(
      data = train_data %>%
        filter(
          sex == 1 & 
            age >= 20 & 
            age <= 79
          ),
      biomarkers = biomarkers_kdm,
      fit = NULL,
      s_ba2 = NULL,
      n_allowed_missing_biomarkers = 0
    )
    test_fem_kdm <- kdm_calc(
      data = data %>%
        filter(sex == 2),
      biomarkers = biomarkers_kdm,
      fit = train_fem_kdm$fit,
      s_ba2 = train_fem_kdm$fit$s_ba2,
      n_allowed_missing_biomarkers = 0
    )
    test_male_kdm <- kdm_calc(
      data = data %>%
        filter(sex == 1),
      biomarkers = biomarkers_kdm,
      fit = train_male_kdm$fit,
      s_ba2 = train_male_kdm$fit$s_ba2,
      n_allowed_missing_biomarkers = 0
    )

    test_kdm <- rbind(test_fem_kdm$data, test_male_kdm$data)

    data <- left_join(data, test_kdm[, c("SEQN", "kdm", "kdm_advance")], by = "SEQN")

    # Calculate PhenoAge
    train_phenoage <- phenoage_calc(
      data = train_data %>%
        filter(
            age >= 20 & 
            age <= 79 & 
            (pregnant != 1 | is.na(pregnant)) &
              !is.na(status) &
              !is.na(time) &
  !(rowSums(is.na(across(all_of(biomarkers_phenoage)))) > 0)
),
      biomarkers = biomarkers_phenoage,
      fit = NULL
    )

    test_phenoage <- phenoage_calc(
      data = data,
      biomarkers = biomarkers_phenoage,
      fit = train_phenoage$fit
    )

    data <- left_join(data, test_phenoage$data[, c("SEQN", "phenoage", "phenoage_advance")], by = "SEQN")

    # Calculate HD
    train_data_hd <- train_data_hd %>%
        filter(
            age <= 30 & 
              age >= 20 &
            (pregnant != 1 | is.na(pregnant)) &
  !(rowSums(is.na(across(all_of(biomarkers_hd)))) > 0)
)
    fem_hd <- hd_calc(
      data = data %>%
        filter(sex == 2),
      reference = train_data_hd %>%
        filter(sex == 2),
      biomarkers = biomarkers_hd
    )
    male_hd <- hd_calc(
      data = data %>%
        filter(sex == 1),
      reference = train_data_hd %>%
        filter(sex == 1),
      biomarkers = biomarkers_hd
    )

    hd <- rbind(fem_hd$data, male_hd$data)

    data <- left_join(data, hd[, c("SEQN", "hd", "hd_log")], by = "SEQN")

    data
  }

  df <- design$variables %>% bioage_calc(
    train_data = train_data,
    train_data_hd = train_data_hd
  )

  update(design,
    kdm = df$kdm,
    kdm_advance = df$kdm_advance,
    phenoage = df$phenoage,
    phenoage_advance = df$phenoage_advance,
    hd = df$hd,
    hd_log = df$hd_log
  )
}

design <- insert_bioage(design, train_data = df_train, train_data_hd = df_train_hd)
```

### Validate biological age calculations

We check the correlations between chronological and biological age and calculate odds ratios on the
differences/advances (corrected for age, sex, race, education, and smoking) and normalize to the
standard deviation and center the mean at 0. The hazard ratios are expressed as hazard per standard
deviation away.

```{r plot_aging_measures}
plot_bioage_correlations <- function(design, predictors, predictors_advance, covariates = c("age", "sex", "race", "education", "smoking")) {
  data <- design$variables

  # scale each predictor to whole-sample z-scores (overall)
  data <- data %>%
    mutate(across(all_of(predictors_advance), ~ as.numeric(scale(.))))

  # age-correlation per predictor (invariant to scaling)
  correlate   <- function(p) cor(data$age, data[[p]], use = "complete.obs")
  correlations <- vapply(predictors, correlate, numeric(1))

  # overall Cox HR per 1 SD of each scaled predictor, adjusted for covariates
  cox_hr <- function(p) {
    form <- as.formula(
      paste0("survival::Surv(time, status) ~ ", p, " + ",
             paste(covariates, collapse = " + "))
    )
    fit <- survival::coxph(form, data = dplyr::filter(data, !is.na(time), !is.na(status)))
    setNames(exp(c(coef(fit)[p], confint(fit)[p, ])), c("HR", "lo", "hi"))
  }
  hrs <- vapply(predictors_advance, cox_hr, numeric(3))   # 3 x n: rows HR/lo/hi, cols = predictors

  # facet labels: correlation + per-SD HR (95% CI)
  labs <- setNames(
    sprintf("%s\nr = %.2f\nHR = %.2f (%.2f, %.2f)",
            predictors, correlations,
            hrs["HR", ], hrs["lo", ], hrs["hi", ]),
    predictors
  )

  data %>%
    pivot_longer(all_of(predictors), names_to = "metric", values_to = "value") %>%
    mutate(sex = factor(sex)) %>%
    ggplot(aes(x = age, y = value)) +
    geom_point(aes(color = sex), alpha = 0.1, shape = 1) +
    geom_smooth(method = "lm", se = TRUE, color = "black", formula = y ~ x) +
    facet_wrap(~metric, scales = "free_y", labeller = as_labeller(labs)) +
    scale_color_manual(
      values = c("1" = color("sunset")(2)[1], "2" = color("sunset")(2)[2]),
      labels = c("1" = "Male", "2" = "Female")
    ) +
    labs(y = "Biological age advance (SD units)") +
    theme_few()
}

plot_bioage_correlations(design, predictors = c("kdm", "phenoage", "hd_log"), predictors_advance = c("kdm_advance", "phenoage_advance", "hd_log"))
```

## Descriptive statistics

### Calculate descriptive statistics

We use the full survey design to generate a descriptive statistics table on the filtered and
unfiltered audiometry samples across cycles. We calculate both weighted and unweighted proportions
and check n per stratum. We only include the strata in the unfiltered dataset that's in the filtered
dataset and remove NA values. Bioage descriptive statistics are also optionally calculated.

```{r calc_descriptives}
calculate_descriptive_statistics <- function(design, incl_bioage = FALSE) {
  design <- update(
    design,
    sex = dplyr::recode(
      sex,
      "1" = "Male",
      "2" = "Female",
      .default = NA_character_
    ),
    smoking = dplyr::recode(
      smoking,
      "1" = "Smoker",
      "2" = "Non-smoker",
      .default = NA_character_
    ),
    race = dplyr::recode(
      race,
      "1" = "Hispanic",
      "3" = "White",
      "4" = "Black",
      "5" = "Other",
      .default = NA_character_
    ),
    education = dplyr::recode(education,
      "1" = "< 9th grade",
      "2" = "9-11th grade",
      "3" = "High school",
      "4" = "Some college or AA degree",
      "5" = "College graduate or higher",
      .default = NA_character_
    )
  )

  groups <- c(Gender = "sex", Race = "race", Smoking = "smoking", Education = "education")
  age_groups <- factor(c("Overall", "20-30", "30-40", "40-50", "50-60", "60-70", if (isFALSE(params$telolength)) "70-80" else NULL))
  hearing_var <- as.formula(paste0("~", params$hearing_var))
  results <- list()

  design <- design %>% subset(
    !is.na(design$variables[[params$hearing_var]]) &
      age_group %in% age_groups &
      !is.na(sex) &
      !is.na(smoking) &
      !is.na(race) &
      !is.na(education)
  )

  for (age_group_val in age_groups) {
    design_subset <- if (age_group_val == "Overall") design else subset(design, age_group == age_group_val)

    if (age_group_val == "Overall") {
      proportion_w <- 1
    } else {
      proportion_w <- as.vector(
        coef(svymean(~age_group, design))[paste0("age_group", age_group_val)]
      )
    }

    rows <- bind_rows(
      data.frame(
        age_group = age_group_val,
        group = "Overall",
        category = "Overall",
        n = nrow(design_subset),
        proportion = nrow(design_subset) / nrow(design),
        proportion_w = proportion_w,
        mean_w_pta = coef(svymean(hearing_var, design_subset)),
        sd_w_pta = sqrt(coef(svyvar(hearing_var, design_subset))),
        mean_w_age = coef(svymean(~age, design_subset)),
        sd_w_age = sqrt(coef(svyvar(~age, design_subset)))
      ) %>%
        {
          if (isTRUE(incl_bioage)) {
            mutate(.,
              mean_w_kdm = coef(svymean(~kdm, design_subset)),
              mean_w_phenoage = coef(svymean(~phenoage, design_subset)),
              mean_w_hd_log = coef(svymean(~hd_log, design_subset)),
              mean_w_telolength_log = coef(svymean(~telolength_log, design_subset, na.rm = TRUE)),
              sd_w_kdm = sqrt(coef(svyvar(~kdm, design_subset))),
              sd_w_phenoage = sqrt(coef(svyvar(~phenoage, design_subset))),
              sd_w_hd_log = sqrt(coef(svyvar(~hd_log, design_subset))),
              sd_w_telolength_log = sqrt(coef(svyvar(~telolength_log, design_subset, na.rm = TRUE))),
            )
          } else {
            .
          }
        },
      lapply(names(groups), function(group) {
        grouping_variable <- as.formula(paste0("~", groups[[group]]))
        proportions_w <- coef(svymean(grouping_variable, design_subset))
        counts <- table(design_subset$variables[[groups[[group]]]])

        data.frame(
          age_group = age_group_val,
          group = group,
          category = svyby(hearing_var, grouping_variable, design_subset, svymean)[[groups[[group]]]],
          n = as.vector(counts),
          proportion = as.vector(counts) / nrow(design_subset),
          proportion_w = as.vector(proportions_w),
          mean_w_pta = svyby(hearing_var, grouping_variable, design_subset, svymean)[[params$hearing_var]],
          sd_w_pta = sqrt(svyby(hearing_var, grouping_variable, design_subset, svyvar)[[params$hearing_var]]),
          mean_w_age = svyby(~age, grouping_variable, design_subset, svymean)[["age"]],
          sd_w_age = sqrt(svyby(~age, grouping_variable, design_subset, svyvar)[["age"]])
        ) %>%
          {
            if (isTRUE(incl_bioage)) {
              mutate(.,
                mean_w_kdm = svyby(~kdm, grouping_variable, design_subset, svymean)$kdm,
                mean_w_phenoage = svyby(~phenoage, grouping_variable, design_subset, svymean)$phenoage,
                mean_w_hd_log = svyby(~hd_log, grouping_variable, design_subset, svymean)$hd_log,
                mean_w_telolength_log = svyby(~telolength_log, grouping_variable, design_subset, svymean, na.rm = TRUE)$telolength_log,
                sd_w_kdm = sqrt(svyby(~kdm, grouping_variable, design_subset, svyvar)$kdm),
                sd_w_phenoage = sqrt(svyby(~phenoage, grouping_variable, design_subset, svyvar)$phenoage),
                sd_w_hd_log = sqrt(svyby(~hd_log, grouping_variable, design_subset, svyvar)$hd_log),
                sd_w_telolength_log = sqrt(svyby(~telolength_log, grouping_variable, design_subset, svyvar, na.rm = TRUE)$telolength_log)
              )
            } else {
              .
            }
          }
      })
    )

    results[[age_group_val]] <- rows
  }

  bind_rows(results) %>%
    mutate(
      group = factor(group, levels = c("Overall", "Gender", "Race", "Smoking", "Education")),
      age_group = factor(age_group, levels = c("Overall", "20-30", "30-40", "40-50", "50-60", "60-70", if (isFALSE(params$telolength)) "70-80" else NULL))
    )
}

descriptives_unfiltered <- calculate_descriptive_statistics(design_no_missing)
descriptives <- calculate_descriptive_statistics(design, incl_bioage = TRUE)
```

### Check survey proportion representation

```{r plot_props}
plot_proportions <- function(data, y_variable, title) {
  ggplot(
    data,
    aes(x = category, y = .data[[y_variable]], group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_text(
      aes(y = .data[[y_variable]] / 2, label = signif(.data[[y_variable]], 2)),
      size = 2
    ) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "\n (dB HL)")
    ) +
    theme_few() +
    scale_y_continuous(breaks = seq(0, 1, 0.5), labels = c("0", "0.5", "1")) +
    coord_cartesian(ylim = c(0, 1)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion_w",
    title = "Weighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted unfiltered strata proportions.png"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion_w",
    title = "Weighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted filtered strata proportions.png"
  )
)

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion",
    title = "Unweighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted unfiltered strata proportions.png"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion",
    title = "Unweighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted filtered strata proportions.png"
  )
)
```

**Proportions remain roughly the same before and after filtering. Proportions are still roughly the
same if including missing audiometry data in proportions.**

### Check descriptive statistics before and after filtering

We plot survey design-adjusted hearing means and variances per-stratum of covariates in the full and
filtered dataset to assess whether the filtered dataset is representative of the general population.

```{r survey_balance_check}
descriptives_unfiltered %>%
  construct_table(
    title = "Descriptive statistics (before filtering)"
  )

descriptives %>%
  construct_table(
    title = "Descriptive statistics (after filtering)"
  )

plot_descriptives <- function(data, title) {
  ggplot(
    data,
    aes(x = category, y = mean_w_pta, group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_errorbar(aes(ymin = mean_w_pta - sd_w_pta, ymax = mean_w_pta + sd_w_pta), width = 0.2) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "\n (dB HL)")
    ) +
    theme_few() +
    coord_cartesian(ylim = c(0, 80)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_descriptives(
    data = descriptives_unfiltered,
    title = paste0("Mean Weighted Unfiltered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted unfiltered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".png")
  )
)

sameplot(
  plot_descriptives(
    data = descriptives,
    title = paste0("Mean Weighted Filtered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 600,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted filtered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".png")
  )
)
```

**Barely any selection bias after excluding asymmetric hearing loss.**

### Assessment of selection bias

```{r check_selection_bias}
plot_selection_bias <- function(ref, test) {
  ref <- ref %>% filter(group == "Overall")
  test <- test %>% filter(group == "Overall")

  diff <- ref$mean_w_pta - test$mean_w_pta

  ref %>%
    mutate(
      delta = diff
    ) %>%
    filter(
      age_group != "Overall"
    ) %>%
    ggplot(aes(x = age_group, y = delta)) +
    geom_point(aes(color = age_group), size = 4, show.legend = FALSE) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_few() +
    labs(x = "Age (Years)", y = paste0("Underestimation ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "", "\n(dB SPL)"))
}

sameplot(
  plot_selection_bias(
    ref = descriptives_unfiltered,
    test = descriptives
  ),
  width = params$double_column_width,
  height = 4,
  res = 600
)
```

## Prepare dataframe

### Create dataframe

We can discard the survey design from this point onward. Further analyses are a nested F-test
between models (in-sample) that share the same design, so inflation of variance due to PSU/strata is
symmetrical and cancels out. Weights are still used for model fitting. For cross-validation, we
estimate prediction error differences between models that share the same design, canceling out the
variance estimate due to PSU/strata. Moreover, gam() doesn't support PSU and strata uncertainty
adjustment.

```{r extract_df}
df <- design$variables
```

### Set data types

```{r set_data_types}
df <- df %>%
  mutate(
    SEQN = as.character(SEQN),
    age_group = factor(age_group),
    sex = factor(sex),
    pregnant = factor(pregnant),
    race = factor(race),
    education = factor(education),
    smoking = factor(smoking)
  )
```

## Clean up environment

From this point onward, we only keep what will actually be used for analysis. We remove df_raw and
df_unfiltered, but don't run `gc()`, as the objects being removed from the environment already
labels the RAM as free to be allocated, thus it's not needed to run garbage collection manually.

```{r cleanup_env}
rm(df_raw)
rm(df_unfiltered)
rm(df_no_missing)
rm(df_train)
rm(df_train_hd)
rm(df_train_raw)
rm(design_no_missing)
rm(descriptives_unfiltered)
```
