---
editor_options:
  markdown:
    wrap: 100
---

# Data preprocessing

## Data selection & standardization

### Select audiometry sample

We select the IDs in the 'Audiometry' files, then open the df from each file, return the df if the
file contains "Audiometry" otherwise, filter the IDs to the list of audiometry IDs.

```{r select_audiometry_sample}
select_audiometry_sample <- function(data_map) {
  # Extract all subject IDs from audiometry files
  audiometry_sample_ids <- data_map %>%
    filter(str_detect(dataset_id, "AUX")) %>%
    pull(data) %>%
    map(~ pull(.x, SEQN)) %>%
    reduce(union)

  # Filter non-audiometry files to audiometry subject IDs only
  data_map %>%
    mutate(
      data = map2(data, dataset_id, \(df, f) {
        if (str_detect(f, "AUX")) {
          df
        } else {
          filter(df, SEQN %in% audiometry_sample_ids)
        }
      })
    )
}
```

### Select and rename columns

Any columns that are missing in the data but are named in the spec for that given cycle will be
printed.

```{r select_cols}
select_columns <- function(data_map, column_path) {
  col_spec <- read.csv(column_path)

  col_spec <- col_spec %>%
    filter(include == TRUE) %>%
    select(label, variable, cycles) %>%
    separate_longer_delim(cycles, delim = ", ")

  # Merge data from the same year into one dataframe
  data_map <- data_map %>%
    group_by(cycle) %>%
    summarise(
      data = list(reduce(data, full_join, by = "SEQN")),
      .groups = "drop"
    )

  # Spec variables that aren't present in their cycle's merged data
  available <- data_map %>%
    transmute(cycle, cols = map(data, names)) %>%
    unnest_longer(cols)

  missing <- col_spec %>%
    anti_join(available, by = c("cycles" = "cycle", "variable" = "cols"))

  if (nrow(missing) > 0) {
    stop(
      "Spec variables missing from their cycle:\n",
      paste0("  ", missing$cycles, " : ", missing$variable, collapse = "\n"),
      call. = FALSE
    )
  }

  # Match variables with spec and rename to label. Per-cycle columns are obtained and named.
  data_map %>%
    mutate(data = map2(data, cycle, function(dat, cy) {
      keep <- col_spec %>% filter(cycles == cy)
      keep <- setNames(keep$variable, keep$label)
      keep <- c(SEQN = "SEQN", keep[keep != "SEQN"])

      dat %>%
        select(any_of(keep)) %>%
        mutate(cycle = cy)
    }))
}
```

### Merge dataframe and create ID vars

```{r merge_cycles}
merge_data <- function(data_map) {
  data_map %>%
    pull(data) %>%
    bind_rows() %>%
    relocate(SEQN, cycle)
}
```

### Standardize values

CDC recommends to use 4-year weights for 1999-2000 and 2001-2002, 3.2-year weights for 2017-2020
March, and 2-year weights for the others. Because different cycles collected audiometry data in
different age populations, we cannot standardize the weights simply by total number of years (15.2)
in the sample, but must standardize each age group by the number of years. Before 2005, audiometry
was performed in only half a sample, but this is corrected for already by the audiometry-specific
weights variable in those earlier years. In this study, people under 20 years old are excluded. The
following ages were assessed in each cycle:

- 1999-2000: 20-69
- 2001-2002: 20-69
- 2003-2004: 20-69
- 2005-2006: 12-19, 70+ (above 84 is set to 84)
- 2009-2010: 12-19, 70+ (above 79 is set to 80)
- 2015-2016: 20-69
- 2017-2020 (up to March): 6-19, 70+ (above 79 is set to 80)

To calculate weights, we divide the number of years in the weights by the total number of years in
the age group.

We censor mortality data above 10 years, so the follow-up period is effectively capped at 10 years
(in line with the Levine paper).

Ages are binned into age groups. Mexican Americans (`race == 1`) and other hispanics (`race == 2`)
are merged into a single hispanics variable. BP is averaged over multiple measurements.

- CRP was measured in mg/L instead of mg/dL in 2015_2016 and 2017_2020 and was converted to mg/dL.

- Creatinine is in mg/dL (and must be used in mg/dL for correction/standardization) and is converted
  to umol/L after standardization to align with original PhenoAge paper. Standardized in 1999_2000,
  2005_2006, and 2017_2020.

- Albumin was converted from g/dL to g/L to align with PhenoAge paper and was backwards standardized
  in 2017_2020.

- The following others were standardized in 2017_2020 (due to new chemistry analyzer):

  - Alkaline phosphatase
  - Aspartate aminotransferase (additionally log-transformed, and no values are below 1 so no
    constant of one needed)
  - Alanine transaminase (additionally log-transformed, and no values are below 1 so no constant of
    one needed)
  - Blood urea nitrogen
  - Gamma glutamyl transpeptidase (additionally log-transformed, and no values are below 1 so no
    constant of one needed)
  - Iron
  - Lactate dehydrogenase
  - Uric acid

```{r standardize_vals}
standardize_values <- function(data) {
  data <- data %>%
    mutate(
      age_eligibility_group = case_when(
        age >= 6 & age < 12 ~ "6-11",
        (age >= 12 & age < 20) | age >= 70 ~ "12-19, 70+",
        age >= 20 & age < 70 ~ "20-69"
      ),
      cycle_years = case_when(
        cycle %in% c("1999_2000", "2001_2002") ~ 4,
        cycle == "2017_2020" ~ 3.2,
        TRUE ~ 2
      ),
      total_years_for_group = case_when(
        age_eligibility_group == "6-11" ~ 3.2,
        age_eligibility_group == "12-19, 70+" ~ 7.2,
        age_eligibility_group == "20-69" ~ 8
      ),
      weights = weights * (cycle_years / total_years_for_group)
    ) %>%
    select(
      -age_eligibility_group,
      -cycle_years,
      -total_years_for_group
    )

  data <- data %>%
    mutate(
      age_group = case_when(
        age < 20 ~ "19-",
        age >= 20 & age < 30 ~ "20–29",
        age >= 30 & age < 40 ~ "30–39",
        age >= 40 & age < 50 ~ "40–49",
        age >= 50 & age < 60 ~ "50–59",
        age >= 60 & age < 70 ~ "60–69",
        isFALSE(params$telolength) & age >= 70 & age < 80 ~ "70–79",
        isFALSE(params$telolength) & age >= 80 ~ "80+"
      )
    )
  data <- data %>%
    mutate(
      status = case_when(
        time > 120 ~ 0,
        TRUE ~ status
      ),
      time = case_when(
        time > 120 ~ 120,
        TRUE ~ time
      )
    )

  data %>%
    mutate(
      race = case_when(
        race == 2 ~ 1,
        TRUE ~ race
      ),
      alb = case_when(
        cycle == "2017_2020" ~ 1.044 * alb + 0.01128,
        TRUE ~ alb
      ),
      alb = alb * 10,
      alp = case_when(
        cycle == "2017_2020" ~ 1.001 * alp + 0.04294,
        TRUE ~ alp
      ),
      ast = case_when(
        cycle == "2017_2020" ~ 1.018 * ast + 3.762,
        TRUE ~ ast
      ),
      ast_log = log(ast),
      alt = case_when(
        cycle == "2017_2020" ~ 1.013 * alt + 2.688,
        TRUE ~ alt
      ),
      alt_log = log(alt),
      bun = case_when(
        cycle == "2017_2020" ~ 1.001 * bun + 0.4488,
        TRUE ~ bun
      ),
      ggt = case_when(
        cycle == "2017_2020" ~ 0.8042 * ggt + 2.363,
        TRUE ~ ggt
      ),
      ggt_log = log(ggt),
      iron = case_when(
        cycle == "2017_2020" ~ 0.9776 * iron - 4.494,
        TRUE ~ iron
      ),
      ld = case_when(
        cycle == "2017_2020" ~ 0.8568 * ld + 2.062,
        TRUE ~ ld
      ),
      ua = case_when(
        cycle == "2017_2020" ~ 0.9323 * ua + 0.2326,
        TRUE ~ ua
      ),
      creat = case_when(
        cycle == "1999_2000" ~ 1.013 * creat + 0.147,
        cycle == "2005_2006" ~ -0.016 + 0.978 * creat,
        cycle == "2017_2020" ~ 1.051 * creat - 0.06945,
        TRUE ~ creat
      ),
      creat = creat * 88.4,
      crp = case_when(
        cycle %in% c("2015_2016", "2017_2020") ~ crp / 10,
        TRUE ~ crp
      ),
      crp_log = log(crp),
      telolength_log = log(telolength)
    ) %>%
    rowwise() %>%
    mutate(
      sbp = case_when(
        cycle == "2017_2020" ~ mean(c(sbp1, sbp2, sbp3), na.rm = TRUE),
        TRUE ~ mean(c(sbp1, sbp2, sbp3, sbp4), na.rm = TRUE)
      )
    ) %>%
    ungroup() %>%
    select(
      -sbp1,
      -sbp2,
      -sbp3,
      -sbp4
    )
}
```

### Calculate hearing metrics

Any values of "888" in audiometry are set to NA so that they're not used for calculating the
audiometric compound measures.

#### List audiometry cols

```{r}
audiometry_cols <- c(
  "audio_left_500",
  "audio_right_500",
  "audio_left_1000",
  "audio_right_1000",
  "audio_left_1000_retest",
  "audio_right_1000_retest",
  "audio_left_2000",
  "audio_right_2000",
  "audio_left_4000",
  "audio_right_4000",
  "audio_left_6000",
  "audio_right_6000",
  "audio_left_8000",
  "audio_right_8000"
)
```

#### Calculate metrics

We recode 'no response' to the ceiling value to avoid bias.

```{r calc_metrics}
calculate_hearing <- function(data, audiometry_cols) {
  data %>%
    mutate(
      across(
        all_of(audiometry_cols),
        ~ case_when(
          . == 888 ~ NA_real_,
          . == 666 & str_detect(cur_column(), "500|1000|2000|4000|6000") & cycle != "2017_2020" ~ 120,
          . == 666 & str_detect(cur_column(), "8000") & cycle != "2017_2020" ~ 110,
          . == 666 & str_detect(cur_column(), "500|1000|2000|4000|6000|8000") & cycle == "2017_2020" ~ 110,
          TRUE ~ as.numeric(.)
        )
      )
    ) %>%
    rowwise() %>%
    mutate(
      test_retest_diff_left = abs(audio_left_1000 - audio_left_1000_retest),
      test_retest_diff_right = abs(audio_right_1000 - audio_right_1000_retest),
      audio_right_1000_mean = mean(c(audio_right_1000, audio_right_1000_retest)),
      audio_left_1000_mean = mean(c(audio_left_1000, audio_left_1000_retest)),
      pta_lf_r = mean(c(audio_right_500, audio_right_1000_mean, audio_right_2000)),
      pta_lf_l = mean(c(audio_left_500, audio_left_1000_mean, audio_left_2000)),
      pta_lf = mean(c(pta_lf_r, pta_lf_l)),
      pta_hf_r = mean(c(audio_right_4000, audio_right_6000, audio_right_8000)),
      pta_hf_l = mean(c(audio_left_4000, audio_left_6000, audio_left_8000)),
      pta_hf = mean(c(pta_hf_r, pta_hf_l))
    ) %>%
    ungroup()
}
```

### Data preprocessing pipeline

#### NHANES 4

1.  Only IDs in the audiometry sample of NHANES are retained.
2.  Only columns that are present in the specs.csv sheet are retained and data files from the same
    year are combined.
3.  Columns are renamed according to specs.csv.
4.  Data from different years is combined.
5.  Weights are standardized, mortality is censored, an age group (10-year bins) column is made, and
    data is standardized according to NHANES guidelines.
6.  Hearing metrics used in this study are calculated.

```{r preprocess_pipe, message=FALSE, warning=FALSE, include=FALSE}
df_unfiltered <- df_raw %>%
  select_audiometry_sample() %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes4.csv")
  ) %>%
  merge_data() %>%
  standardize_values() %>%
  calculate_hearing(audiometry_cols = audiometry_cols)
```

#### NHANES 3

Training is only used for biological age algorithms and there's only a single cycle so no
standardization or audiometry subsampling required. Filtering will be performed in the biological
age calculation chunk. Missing data is recoded to NA.

```{r preprocess_pipe_train}
df_train <- df_train_raw %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes3.csv")
  ) %>%
  merge_data() %>%
  mutate(
    status = case_when(
      time > 120 ~ 0,
      TRUE ~ status
    ),
    time = case_when(
      time > 120 ~ 120,
      TRUE ~ time
    ),
    sbp = case_when(
      sbp == 888 ~ NA,
      TRUE ~ sbp
    ),
    alb = case_when(
      alb == 888 ~ NA,
      TRUE ~ alb
    ),
    glu = case_when(
      glu == 888888 ~ NA,
      TRUE ~ glu
    ),
    creat = case_when(
      creat == 888888 ~ NA,
      TRUE ~ creat
    ),
    crp = case_when(
      crp == 88888 ~ NA,
      TRUE ~ crp
    ),
    mono = case_when(
      mono == 88888 ~ NA,
      TRUE ~ mono
    ),
    lymph = case_when(
      lymph == 88888 ~ NA,
      TRUE ~ lymph
    ),
    mcv = case_when(
      mcv == 88888 ~ NA,
      TRUE ~ mcv
    ),
    rdw = case_when(
      rdw == 88888 ~ NA,
      TRUE ~ rdw
    ),
    alp = case_when(
      alp == 8888 ~ NA,
      TRUE ~ alp
    ),
    wbc = case_when(
      wbc == 88888 ~ NA,
      TRUE ~ wbc
    ),
    hba1c = case_when(
      hba1c == 8888 ~ NA,
      TRUE ~ hba1c
    ),
    totchol = case_when(
      totchol == 888 ~ NA,
      TRUE ~ totchol
    ),
    bun = case_when(
      bun == 888 ~ NA,
      TRUE ~ bun
    ),
    rbc = case_when(
      rbc == 8888 ~ NA,
      TRUE ~ rbc
    ),
    crp_log = log(crp),
    creat = 0.96 * creat - 0.184,
  )
```

#### NHANES 4 HD Train

This is the full unfiltered NHANES 4 dataset from the included cycles. For training HD, we do not
require only audiometry data.

```{r}
df_train_hd <- df_raw %>%
  select_columns(
    column_path = here("specs", "col_spec_nhanes4.csv")
  ) %>%
  merge_data() %>%
  standardize_values()
```

## Survey design

### Check completeness

Make a scrollable table and only show number of data present in ID, weights, and columns that have
no values at all.

```{r check_complete}
construct_table <- function(input, title) {
  input %>%
    kable(
      "html",
      caption = title
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}

df_unfiltered %>%
  group_by(cycle) %>%
  summarise(across(everything(), ~ sum(!is.na(.)), .names = "n_{.col}")) %>%
  select(cycle, n_SEQN, n_weights, where(~ any(. == 0))) %>%
  construct_table(title = "Check present values")
```

Telomere length was only assessed in early cohorts. Pregnancy is missing in 2009 and 2017 because it
was only assessed in women aged 20-44 and the audiometry sample only includes people aged 12-19 and
70+. Mortality data is missing in 2017-2020. Some IDs lack all audiometry data from 1999 and 2001.
Because we have no data at all in these subjects (weights are also missing), we must exclude them.
NHANES guidelines state that \< 10% missing is acceptable without further adjustment needed.

### Exclude missing rows (no data at all)

```{r exclude_missing}
filter_data <- function(data) {
  verbose_filter <- function(data, ..., label) {
    start_n <- nrow(data)
    data_filtered <- data %>% filter(...)
    message(nrow(data_filtered), " remaining. ", start_n - nrow(data_filtered), "/", start_n, " excluded due to ", label)

    data_filtered
  }

  data <- data %>%
    verbose_filter(
      !is.na(weights),
      label = "missing weights"
    )
}

df_no_missing <- df_unfiltered %>% filter_data()
```

### Create survey design

We use the Survey package to account for the full survey design (weights, strata, PSUs)

```{r create_design}
design_no_missing <- svydesign(
  ids     = ~psu,
  strata  = ~stratum,
  weights = ~weights,
  data    = df_no_missing,
  nest    = TRUE
)
```

## Data filtering

### List biomarkers

This list will be reused throughout the scripts whenever we need to do something using all included
biomarkers.

```{r list_bm}
list_bm <- function(spec_path) {
  specs <- read.csv(spec_path)

  bm_labels <- unique(
    specs %>% filter(
      include == TRUE,
      vartype == "biomarker",
      !label %in% c(
        "sbp1",
        "sbp2",
        "sbp3",
        "sbp4",
        "alt",
        "ast",
        "ggt",
        "crp"
      )
    ) %>%
      pull(label)
  )
  c(bm_labels, "sbp", "alt_log", "ast_log", "ggt_log", "crp_log")
}

biomarker_vars <- list_bm(here("specs", "col_spec_nhanes4.csv"))
```

We keep any columns we may need at this stage or have already used, such as the individual
audiometry columns. The design is stored intermittently in the pipe before any `if_any` statements
are used, because these don't work on the design object, only on the `design$variable` within the
design object. Therefore, in those cases, we store the design prior to the statement to ensure that
`design$variable` refers to the filtered `design`, set filter criteria on `design$variable`, which
is a dataframe/tidytable, and pass on a list of boolean values to select based on `design$variables`
to subset `design` on.

1.  Ages of 80 or higher are cut off to 80, so we exclude these ages because we don't know the ages
    of the subjects. Because 2005-2006 is the only cycle with people with ages marked as age 80-84
    and because there are no people of non-white races among these, we excluded all people above 80.
    This also avoids weighting issues of the 70+ group in earlier and later cycles.
2.  People under 20 years of age aren't included in this study.
3.  Pregnant women are excluded (NAs are explicitly kept).
4.  Subjects with missing audiometry data for frequencies used in PTA are excluded ('no response
    values' have been recoded to ceiling).
5.  Subjects with large test-retest differences are excluded, as this suggests unreliable
    measurement.
6.  Subjects with missing or unknown covariate data are excluded.
7.  Subjects with missing biomarker data are excluded (as these are required for calculating
    biological ages).
8.  Subjects with any biomarker values \> 5 SDs away from the sex-specific mean are excluded to
    avoid extreme values dominating the models.

### Filter data

```{r filter_df}
subset_data <- function(design, biomarker_vars, audiometry_cols) {
  verbose_subset <- function(design, ..., label) {
    start_n <- nrow(design)
    design_subset <- design %>% subset(...)
    message(nrow(design_subset), " remaining. ", start_n - nrow(design_subset), "/", start_n, " excluded due to ", label)

    design_subset
  }

  design <- design %>%
    verbose_subset(
      age < 80,
      label = "too old age."
    ) %>%
    verbose_subset(
      age >= 20,
      label = "too young age."
    ) %>%
    verbose_subset(
      pregnant != 1 | is.na(pregnant),
      label = "pregnancy."
    )

  design <- design %>%
    verbose_subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(audiometry_cols), ~ . == 888 | . == 666 | is.na(.))) %>%
        pull(keep),
      label = "missing audiometry data."
    ) %>%
    verbose_subset(
      test_retest_diff_left < 10 & test_retest_diff_right < 10,
      label = "too large test-retest difference."
    ) %>%
    verbose_subset(
      smoking %in% c(1, 2),
      label = "smoking refused or 'don't know'."
    ) %>%
    verbose_subset(
      !(education %in% c(7, 9)),
      label = "education refused or 'don't know'."
    )

  design <- design %>%
    verbose_subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(biomarker_vars), ~ is.na(.))) %>%
        pull(keep),
      label = "missing biomarkers."
    )

  design %>%
    verbose_subset(
      design$variables %>%
        group_by(sex) %>%
        mutate(keep = !if_any(
          all_of(biomarker_vars),
          ~ {
            m <- mean(.x, na.rm = TRUE)
            s <- sd(.x, na.rm = TRUE)
            abs(.x - m) > 5 * s
          }
        )) %>%
        ungroup() %>%
        pull(keep),
      label = "biomarker outlier (greater or smaller than 5 SDs)."
    )
}

design <- design_no_missing %>% subset_data(
  biomarker_vars = biomarker_vars,
  audiometry_cols = audiometry_cols
)
```

### Optionally only include telolength

```{r filt_telolength}
if (isTRUE(params$telolength)) {
  design <- design %>%
    subset(
      !is.na(telolength)
    )
}
```

### Check data issues (NHANES3 is unweighted)

```{r check_df_balance}
check_biomarker_standardization <- function(design, training_set, biomarker_vars, training_spec_path) {
  all_vars <- c("age", "pta_hf", "pta_lf", biomarker_vars)
  cycles <- unique(design$variables$cycle)

  nhanes4 <- map_dfr(cycles, function(cyc) {
    des_cyc <- subset(design, cycle == cyc)

    stats <- map(all_vars, function(v) {
      fmla_v <- as.formula(paste("~", v))
      m <- as.numeric(svymean(fmla_v, des_cyc, na.rm = TRUE))
      s <- sqrt(as.numeric(svyvar(fmla_v, des_cyc, na.rm = TRUE)))
      paste0(round(m, 2), "(", round(s, 2), ")")
    }) %>%
      set_names(all_vars) %>%
      as_tibble()

    bind_cols(tibble(cycle = cyc), stats)
  })

  col_spec_train <- read.csv(training_spec_path)

  biomarkers_train <- col_spec_train %>%
    filter(
      vartype == "biomarker",
      include == TRUE,
      label != "crp"
    ) %>%
    pull(label)
  all_vars_train <- c("age", "crp_log", biomarkers_train)


  nhanes3 <- map(all_vars_train, function(v) {
    m <- as.numeric(mean(training_set[[v]], na.rm = TRUE))
    s <- as.numeric(sd(training_set[[v]], na.rm = TRUE))
    paste0(round(m, 2), "(", round(s, 2), ")")
  }) %>%
    set_names(all_vars_train) %>%
    as_tibble() %>%
    mutate(cycle = "1988_1994")

  bind_rows(nhanes4, nhanes3) %>%
    arrange(cycle) %>%
    construct_table(title = "Check cohort age and biomarker balance")
}

design %>% check_biomarker_standardization(
  training_set = df_train,
  biomarker_vars = biomarker_vars,
  training_spec_path = here("specs", "col_spec_nhanes3.csv")
)
```

ALP was higher in NHANES3, but could be in part due to survey design (older participants receive
higher weights) and large SD. Creatinine is also higher in NHANES3 but could be due to the same
reason. CRP being lower could be due to different detection limit.

#### Verify normal distribution

```{r}
check_distribution <- function(data, spec_path, dataset) {
  col_spec <- read.csv(spec_path)
  biomarkers <- col_spec %>%
    filter(
      vartype == "biomarker",
      include == TRUE,
      !(label %in% c("sbp1", "sbp2", "sbp3", "sbp4", "crp", "ast", "alt", "ggt"))
    ) %>%
    pull(label)

  if (dataset == "NHANES4") {
    biomarkers <- unique(c(biomarkers, "sbp", "crp_log", "ast_log", "alt_log", "ggt_log"))
  } else {
    biomarkers <- unique(c(biomarkers, "sbp", "crp_log"))
  }

  data %>%
    select(all_of(biomarkers)) %>%
    pivot_longer(names_to = "biomarker", values_to = "value") %>%
    ggplot(aes(x = value)) +
    geom_histogram(
      aes(y = ..density..),
      fill = "tomato",
      color = "white",
      alpha = 0.7
    ) +
    geom_density(
      color = "darkred",
      linewidth = 1,
      adjust = 3
    ) +
    facet_wrap(~biomarker, scales = "free") +
    theme_few() +
    theme(
      axis.text = element_blank()
    )
}

sameplot(
  design$variables %>% check_distribution(
    spec_path = here("specs", "col_spec_nhanes4.csv"),
    dataset = "NHANES4"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "biomarker_distributions_NHANES4.tif"
  )
)

sameplot(
  df_train %>% check_distribution(
    spec_path = here("specs", "col_spec_nhanes3.csv"),
    dataset = "NHANES3"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "biomarker_distributions_NHANES3.tif"
  )
)
```

There is some kurtosis for several variables in NHANES3, but that should not be an issue.

#### Check PSU variance

```{r}
design$variables %>%
  distinct(stratum, psu) %>%
  count(stratum, name = "psu_n") %>%
  count(psu_n, name = "stratum_n") %>%
  construct_table(title = "Number of PSUs within strata")
```

**One stratum only has one PSU after filtering, so we can't estimate variance within it.**

## Calculate biological age

We train PhenoAge on ages 20-79 (full sample) in line with the Levine paper. We train KDM on ages
30-75 to ensure people were old enough to have detectable age-related biomarkers but not so old that
the sample represents people with above-average health due to survivor bias. We train HD on people
aged 20-45, because we want a younger sample with good health but not so young that they represent a
very different demographic from the total population. We additionally exclude people with missing
biomarker data and pregnant women in all measures and people with missing mortality data for
PhenoAge.

```{r calc_bioage, warning=FALSE}
insert_bioage <- function(design, train_data, train_data_hd) {
  bioage_calc <- function(data, train_data, train_data_hd) {
    biomarkers_phenoage <- c(
      "alb", "alp", "crp_log", "creat", "lymph", "mcv", "wbc", "rdw", "glu"
    )
    biomarkers_kdm <- c(
      "alb", "creat", "crp_log", "mono", "alp", "hba1c", "totchol", "bun", "sbp", "rbc"
    )
    biomarkers_hd <- c(
      "alb", "alp", "alt_log", "ast_log", "baso", "cal", "chlor", "ggt_log", "hdl", "hg", "htc", "iron", "ld", "lymph", "mch", "mchc", "mono", "neut", "phos", "pot", "rbc", "rdw", "sod", "totchol", "tp", "ua", "wbc"
    )

    # KDM
    train_data_kdm <- train_data %>%
      filter(
        age >= 30 &
          age <= 75 &
          !(is.na(sex)) &
          (pregnant != 1 | is.na(pregnant)) &
          !(rowSums(is.na(across(all_of(biomarkers_kdm)))) > 0)
      )

    message("N train KDM = ", nrow(train_data_kdm))

    train_fem_kdm <- kdm_calc(
      data = train_data_kdm %>%
        filter(sex == 2),
      biomarkers = biomarkers_kdm,
      verbose = FALSE
    )

    train_male_kdm <- kdm_calc(
      data = train_data_kdm %>%
        filter(sex == 1),
      biomarkers = biomarkers_kdm,
      verbose = FALSE
    )

    test_fem_kdm <- kdm_calc(
      data = data %>%
        filter(sex == 2),
      biomarkers = biomarkers_kdm,
      fit = train_fem_kdm$fit,
      s_ba2 = train_fem_kdm$fit$s_ba2,
      verbose = FALSE
    )

    test_male_kdm <- kdm_calc(
      data = data %>%
        filter(sex == 1),
      biomarkers = biomarkers_kdm,
      fit = train_male_kdm$fit,
      s_ba2 = train_male_kdm$fit$s_ba2,
      verbose = FALSE
    )

    test_kdm <- rbind(test_fem_kdm$data, test_male_kdm$data)

    data <- left_join(data, test_kdm[, c("SEQN", "kdm", "kdm_advance")], by = "SEQN")

    # PhenoAge
    train_data_phenoage <- train_data %>%
      filter(
        age >= 20 &
          age <= 79 &
          (pregnant != 1 | is.na(pregnant)) &
          !is.na(status) &
          !is.na(time) &
          !(rowSums(is.na(across(all_of(biomarkers_phenoage)))) > 0)
      )

    message("N train PhenoAge = ", nrow(train_data_phenoage))

    train_phenoage <- phenoage_calc(
      data = train_data_phenoage,
      biomarkers = biomarkers_phenoage
    )

    test_phenoage <- phenoage_calc(
      data = data,
      biomarkers = biomarkers_phenoage,
      fit = train_phenoage$fit
    )

    data <- left_join(data, test_phenoage$data[, c("SEQN", "phenoage", "phenoage_advance")], by = "SEQN")

    # HD
    train_data_hd <- train_data_hd %>%
      filter(
        age <= 45 &
          age >= 20 &
          !(is.na(sex)) &
          (pregnant != 1 | is.na(pregnant)) &
          !(rowSums(is.na(across(all_of(biomarkers_hd)))) > 0)
      )

    message("N train HD = ", nrow(train_data_hd))

    fem_hd <- hd_calc(
      data = data %>%
        filter(sex == 2),
      reference = train_data_hd %>%
        filter(sex == 2),
      biomarkers = biomarkers_hd
    )
    male_hd <- hd_calc(
      data = data %>%
        filter(sex == 1),
      reference = train_data_hd %>%
        filter(sex == 1),
      biomarkers = biomarkers_hd
    )

    hd <- rbind(fem_hd$data, male_hd$data)

    data <- left_join(data, hd[, c("SEQN", "hd", "hd_log")], by = "SEQN")

    data
  }

  df <- design$variables %>% bioage_calc(
    train_data = train_data,
    train_data_hd = train_data_hd
  )

  update(design,
    kdm = df$kdm,
    kdm_advance = df$kdm_advance,
    phenoage = df$phenoage,
    phenoage_advance = df$phenoage_advance,
    hd = df$hd,
    hd_log = df$hd_log
  )
}

design <- insert_bioage(design, train_data = df_train, train_data_hd = df_train_hd)
```

### Validate biological age calculations (supplementary)

We check the correlations between chronological and biological age and calculate Cox hazard ratios
on the differences/advances (corrected for age, sex, race, education, and smoking) and normalize to
the standard deviation so that 1 unit away means 1 standard deviation. The hazard ratios are
expressed as hazard per standard deviation away. The hazard ratios are unweighted.

```{r plot_aging_measures}
plot_bioage_correlations <- function(design, predictors, labels, predictors_advance, covariates = c("age", "sex", "race", "education", "smoking")) {
  data <- design$variables

  data <- data %>%
    mutate(
      across(
        all_of(predictors_advance),
        ~ .x / sd(.x, na.rm = TRUE)
      )
    )

  correlate <- function(p) cor(data$age, data[[p]], use = "complete.obs")
  correlations <- vapply(predictors, correlate, numeric(1))

  cox_hr <- function(p) {
    form <- as.formula(
      paste0(
        "survival::Surv(time, status) ~ ", p, " + ",
        paste(covariates, collapse = " + ")
      )
    )
    test_set <- data %>%
      filter(
        !is.na(time),
        !is.na(status),
        p != "telolength_log" | !is.na(telolength_log)
      ) %>%
      mutate(
        "{p}" := if (p == "telolength_log") {
          -.data[[p]]
        } else {
          .data[[p]]
        }
      )

    message(p, " test set size = ", nrow(test_set))

    fit <- survival::coxph(form, data = test_set)
    setNames(exp(c(coef(fit)[p], confint(fit)[p, ])), c("HR", "ci_low", "ci_high"))
  }
  hazard_ratios <- vapply(predictors_advance, cox_hr, numeric(3))

  labs <- setNames(
    sprintf(
      "<b>%s</b><br><i>r</i> = %.2f<br>HR = %.2f<br>HR 95%% CI [%.2f, %.2f]",
      labels,
      correlations,
      hazard_ratios["HR", ],
      hazard_ratios["ci_low", ],
      hazard_ratios["ci_high", ]
    ),
    predictors
  )

  data %>%
    pivot_longer(all_of(predictors), names_to = "metric", values_to = "value") %>%
    mutate(sex = factor(sex)) %>%
    ggplot(aes(x = age, y = value)) +
    geom_point(aes(color = sex, alpha = 0.04 * weights / mean(weights)), shape = 20) +
    geom_smooth(method = "lm", se = TRUE, color = "black", formula = y ~ x) +
    facet_wrap(~metric, scales = "free_y", labeller = as_labeller(labs), nrow = 1) +
    scale_alpha_identity() +
    scale_color_manual(
      values = c("1" = color("sunset")(2)[1], "2" = color("sunset")(2)[2]),
      labels = c("1" = "Male", "2" = "Female")
    ) +
    labs(y = "Biological Age", x = "Age (Years)") +
    theme_few() +
    theme(
      axis.text.x = element_markdown(
        "sans",
        size = 8
      ),
      axis.text.y = element_markdown(
        "sans",
        size = 9
      ),
      strip.text.x = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = 9,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
}


sameplot(
  plot_bioage_correlations(design,
    predictors = c("kdm", "phenoage", "hd_log", "telolength_log"),
    labels = c("KDM", "PhenoAge", "HD (Log)", "Telomere Length (Log)"),
    predictors_advance = c("kdm_advance", "phenoage_advance", "hd_log", "telolength_log")
  ),
  width = params$double_column_width,
  height = 4,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Correlations.tif"
  )
)
```

## Descriptive statistics

### Calculate descriptive statistics

We use the full survey design to generate a descriptive statistics table on the filtered and
unfiltered audiometry samples across cycles. We calculate both weighted and unweighted proportions
and check n per stratum. We only include the strata in the unfiltered dataset that's in the filtered
dataset and remove NA values. Bioage descriptive statistics are also optionally calculated.

```{r calc_descriptives}
calculate_descriptive_statistics <- function(design, audiometry_cols, incl_bioage = FALSE) {
  design <- update(
    design,
    sex = dplyr::recode(
      sex,
      "1" = "Male",
      "2" = "Female",
      .default = NA_character_
    ),
    smoking = dplyr::recode(
      smoking,
      "1" = "Smoker",
      "2" = "Non-smoker",
      .default = NA_character_
    ),
    race = dplyr::recode(
      race,
      "1" = "Hispanic",
      "3" = "White",
      "4" = "Black",
      "5" = "Other",
      .default = NA_character_
    ),
    education = dplyr::recode(education,
      "1" = "< 9th grade",
      "2" = "9-11th grade",
      "3" = "High school",
      "4" = "Some college or AA degree",
      "5" = "College graduate or higher",
      .default = NA_character_
    )
  )

  groups <- c(Sex = "sex", Race = "race", Smoking = "smoking", Education = "education")
  age_groups <- factor(c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
  results <- list()

  design <- design %>%
    subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(audiometry_cols), ~ . == 888 | . == 666 | is.na(.))) %>%
        pull(keep) &
        age_group %in% age_groups &
        !is.na(sex) &
        !is.na(smoking) &
        !is.na(race) &
        !is.na(education)
    )


  for (age_group_val in age_groups) {
    design_subset <- if (age_group_val == "Overall") design else subset(design, age_group == age_group_val)

    if (age_group_val == "Overall") {
      proportion_w <- 1
    } else {
      proportion_w <- as.vector(
        coef(svymean(~age_group, design))[paste0("age_group", age_group_val)]
      )
    }

    rows <- bind_rows(
      data.frame(
        age_group = age_group_val,
        group = "Overall",
        category = "Overall",
        n = nrow(design_subset),
        proportion = nrow(design_subset) / nrow(design),
        proportion_w = proportion_w,
        mean_w_pta_hf = coef(svymean(~pta_hf, design_subset)),
        sd_w_pta_hf = sqrt(coef(svyvar(~pta_hf, design_subset))),
        mean_w_pta_lf = coef(svymean(~pta_lf, design_subset)),
        sd_w_pta_lf = sqrt(coef(svyvar(~pta_lf, design_subset))),
        mean_w_age = coef(svymean(~age, design_subset)),
        sd_w_age = sqrt(coef(svyvar(~age, design_subset)))
      ) %>%
        {
          if (isTRUE(incl_bioage)) {
            mutate(.,
              mean_w_kdm = coef(svymean(~kdm, design_subset)),
              mean_w_phenoage = coef(svymean(~phenoage, design_subset)),
              mean_w_hd_log = coef(svymean(~hd_log, design_subset)),
              mean_w_telolength_log = coef(svymean(~telolength_log, design_subset, na.rm = TRUE)),
              sd_w_kdm = sqrt(coef(svyvar(~kdm, design_subset))),
              sd_w_phenoage = sqrt(coef(svyvar(~phenoage, design_subset))),
              sd_w_hd_log = sqrt(coef(svyvar(~hd_log, design_subset))),
              sd_w_telolength_log = sqrt(coef(svyvar(~telolength_log, design_subset, na.rm = TRUE))),
            )
          } else {
            .
          }
        },
      lapply(names(groups), function(group) {
        grouping_variable <- as.formula(paste0("~", groups[[group]]))
        proportions_w <- coef(svymean(grouping_variable, design_subset))
        counts <- table(design_subset$variables[[groups[[group]]]])

        data.frame(
          age_group = age_group_val,
          group = group,
          category = svyby(~pta_hf, grouping_variable, design_subset, svymean)[[groups[[group]]]],
          n = as.vector(counts),
          proportion = as.vector(counts) / nrow(design_subset),
          proportion_w = as.vector(proportions_w),
          mean_w_pta_hf = svyby(~pta_hf, grouping_variable, design_subset, svymean)[["pta_hf"]],
          sd_w_pta_hf = sqrt(svyby(~pta_hf, grouping_variable, design_subset, svyvar)[["pta_hf"]]),
          mean_w_pta_lf = svyby(~pta_lf, grouping_variable, design_subset, svymean)[["pta_lf"]],
          sd_w_pta_lf = sqrt(svyby(~pta_lf, grouping_variable, design_subset, svyvar)[["pta_lf"]]),
          mean_w_age = svyby(~age, grouping_variable, design_subset, svymean)[["age"]],
          sd_w_age = sqrt(svyby(~age, grouping_variable, design_subset, svyvar)[["age"]])
        ) %>%
          {
            if (isTRUE(incl_bioage)) {
              mutate(.,
                mean_w_kdm = svyby(~kdm, grouping_variable, design_subset, svymean)$kdm,
                mean_w_phenoage = svyby(~phenoage, grouping_variable, design_subset, svymean)$phenoage,
                mean_w_hd_log = svyby(~hd_log, grouping_variable, design_subset, svymean)$hd_log,
                mean_w_telolength_log = svyby(~telolength_log, grouping_variable, design_subset, svymean, na.rm = TRUE)$telolength_log,
                sd_w_kdm = sqrt(svyby(~kdm, grouping_variable, design_subset, svyvar)$kdm),
                sd_w_phenoage = sqrt(svyby(~phenoage, grouping_variable, design_subset, svyvar)$phenoage),
                sd_w_hd_log = sqrt(svyby(~hd_log, grouping_variable, design_subset, svyvar)$hd_log),
                sd_w_telolength_log = sqrt(svyby(~telolength_log, grouping_variable, design_subset, svyvar, na.rm = TRUE)$telolength_log)
              )
            } else {
              .
            }
          }
      })
    )

    results[[age_group_val]] <- rows
  }

  bind_rows(results) %>%
    mutate(
      group = factor(group, levels = c("Overall", "Sex", "Race", "Smoking", "Education")),
      age_group = factor(age_group, levels = c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
    )
}

descriptives_unfiltered <- calculate_descriptive_statistics(design_no_missing, audiometry_cols = audiometry_cols)
descriptives <- calculate_descriptive_statistics(design, audiometry_cols = audiometry_cols, incl_bioage = TRUE)
```

### Show descriptives tables

```{r show_descr_table}
descriptives_unfiltered %>%
  construct_table(
    title = "Descriptive statistics (before filtering)"
  )

descriptives %>%
  construct_table(
    title = "Descriptive statistics (after filtering)"
  )
```

#### Show summary descriptives table

```{r summary_descr_table}
descriptive_summary_table <- bind_rows(
  descriptives %>%
    filter(age_group == "Overall", group == "Sex") %>%
    arrange(match(category, c("Male", "Female"))) %>%
    transmute(
      Characteristic = as.character(category),
      Value = sprintf(
        "%.1f%% (<i>n</i> = %s)",
        100 * proportion_w,
        format(n, big.mark = ",")
      )
    ) %>%
    add_row(Characteristic = "<b>Sex</b>", Value = "", .before = 1),
  descriptives %>%
    filter(group == "Overall", category == "Overall") %>%
    transmute(
      Characteristic = if_else(
        age_group == "Overall",
        "<b>Chronological Age (years)</b>",
        as.character(age_group)
      ),
      Value = if_else(
        age_group == "Overall",
        sprintf("%.1f ± %.1f", mean_w_age, sd_w_age),
        sprintf(
          "%.1f%% (<i>n</i> = %s)",
          100 * proportion_w,
          format(n, big.mark = ",")
        )
      )
    ),
  descriptives %>%
    filter(age_group == "Overall", group == "Race") %>%
    arrange(match(category, c("White", "Black", "Hispanic", "Other"))) %>%
    transmute(
      Characteristic = recode(
        as.character(category),
        White = "White, non-Hispanic",
        Black = "Black, non-Hispanic"
      ),
      Value = sprintf(
        "%.1f%% (<i>n</i> = %s)",
        100 * proportion_w,
        format(n, big.mark = ",")
      )
    ) %>%
    add_row(
      Characteristic = "<b>Race/Ethnicity</b>",
      Value = "",
      .before = 1
    ),
  descriptives %>%
    filter(
      age_group == "Overall",
      group == "Overall",
      category == "Overall"
    ) %>%
    transmute(
      Characteristic = list(sprintf("<b>%s</b>", c(
        "HF-PTA (dB HL)",
        "LF-PTA (dB HL)",
        "KDM",
        "PhenoAge",
        "HD (log)",
        "Telomere Mean Length (log)"
      ))),
      Value = list(c(
        sprintf("%.1f ± %.1f", mean_w_pta_hf, sd_w_pta_hf),
        sprintf("%.1f ± %.1f", mean_w_pta_lf, sd_w_pta_lf),
        sprintf("%.1f ± %.1f", mean_w_kdm, sd_w_kdm),
        sprintf("%.1f ± %.1f", mean_w_phenoage, sd_w_phenoage),
        sprintf("%.1f ± %.1f", mean_w_hd_log, sd_w_hd_log),
        sprintf(
          "%.2f ± %.2f (<i>n</i> = %s)",
          mean_w_telolength_log,
          sd_w_telolength_log,
          format(
            design$variables %>%
              filter(!is.na(telolength_log)) %>%
              nrow(),
            big.mark = ","
          )
        )
      ))
    ) %>%
    unnest(c(Characteristic, Value))
) %>%
  kable(
    format = "html",
    col.names = c(
      "Characteristic",
      "Percentage or<br>Mean ± Standard Deviation"
    ),
    align = c("c", "c"),
    escape = FALSE,
    table.attr = paste0(
      'cellspacing="0" cellpadding="0" ',
      'style="width:159mm; table-layout:fixed; border-collapse:collapse; ',
      "border:1px solid black; margin-left:auto; margin-right:auto; ",
      "font-family:'Times New Roman', Times, serif; ",
      'font-size:10pt; color:black;"'
    )
  ) %>%
  row_spec(
    0,
    extra_css = paste0(
      "border-bottom:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  row_spec(
    if (isFALSE(params$telolength)) c(3, 10, 15:21) else c(3, 9, 14:20),
    extra_css = paste0(
      "border-bottom:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  column_spec(
    1,
    include_thead = TRUE,
    extra_css = paste0(
      "border-right:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  column_spec(
    2,
    extra_css = paste0(
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  )

if (isTRUE(params$save_figures)) {
  save_kable(descriptive_summary_table, here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "descriptive_summary.html"
  ))
}

descriptive_summary_table
```

### Check survey proportion representation

```{r plot_props}
plot_proportions <- function(data, y_variable, title) {
  ggplot(
    data,
    aes(x = category, y = .data[[y_variable]], group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_text(
      aes(y = .data[[y_variable]] / 2, label = signif(.data[[y_variable]], 2)),
      size = 2
    ) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = "Proportion"
    ) +
    theme_few() +
    scale_y_continuous(breaks = seq(0, 1, 0.5), labels = c("0", "0.5", "1")) +
    coord_cartesian(ylim = c(0, 1)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion_w",
    title = "Weighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted unfiltered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion_w",
    title = "Weighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted filtered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion",
    title = "Unweighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted unfiltered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion",
    title = "Unweighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted filtered strata proportions.tif"
  )
)
```

**Proportions remain roughly the same before and after filtering. Gender skews slightly towards men
after filtering due to excluding pregnant women. Relative black people drops by \~12% after
filtering but absolute change is under 3% for any stratum.**

### Check descriptive statistics before and after filtering

We plot survey design-adjusted hearing means and variances per-stratum of covariates in the full and
filtered dataset to assess whether the filtered dataset is representative of the general population.

```{r survey_balance_check}
plot_descriptives <- function(data, title) {
  paste0("mean_w_", params$hearing_var)
  ggplot(
    data,
    aes(x = category, y = .data[[paste0("mean_w_", params$hearing_var)]], group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_errorbar(aes(ymin = .data[[paste0("mean_w_", params$hearing_var)]] - .data[[paste0("sd_w_", params$hearing_var)]], ymax = .data[[paste0("mean_w_", params$hearing_var)]] + .data[[paste0("sd_w_", params$hearing_var)]]), width = 0.2) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "\n (dB HL)")
    ) +
    theme_few() +
    coord_cartesian(ylim = c(0, 80)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_descriptives(
    data = descriptives_unfiltered,
    title = paste0("Mean Weighted Unfiltered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted unfiltered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".tif")
  )
)

sameplot(
  plot_descriptives(
    data = descriptives,
    title = paste0("Mean Weighted Filtered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted filtered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".tif")
  )
)
```

### Assessment of selection bias

```{r check_selection_bias}
plot_selection_bias <- function(ref, test) {
  ref <- ref %>% filter(group == "Overall")
  test <- test %>% filter(group == "Overall")

  diff <- ref[[paste0("mean_w_", params$hearing_var)]] - test[[paste0("mean_w_", params$hearing_var)]]

  ref %>%
    mutate(
      delta = diff
    ) %>%
    filter(
      age_group != "Overall"
    ) %>%
    ggplot(aes(x = age_group, y = delta)) +
    geom_point(aes(color = age_group), size = 4, show.legend = FALSE) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_few() +
    labs(x = "Age (Years)", y = paste0("Underestimation ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "", "\n(dB HL)"))
}

sameplot(
  plot_selection_bias(
    ref = descriptives_unfiltered,
    test = descriptives
  ),
  width = params$double_column_width,
  height = 4,
  res = 1200
)
```

Selection bias is minimal in both LF-PTA and HF-PTA.

## Prepare dataframe

### Create dataframe

We can discard the survey design from this point onward. Further analyses are a nested F-test
between models (in-sample) that share the same design, so inflation of variance due to PSU/strata
should be fairly symmetrical and not of high impact. Weights are still used for model fitting. For
cross-validation, we estimate prediction error differences between models that share the same
design, not a population estimate. Moreover, gam() doesn't support PSU and strata uncertainty
adjustment/complex survey design. We will use PSUs within strata to select bootstrapping samples
later on, but we don't need the survey object for this.

```{r extract_df}
df <- design$variables

df <- df %>%
  select(
    SEQN,
    weights,
    psu,
    stratum,
    age,
    age_group,
    all_of(covariate_vars),
    kdm,
    kdm_advance,
    phenoage,
    phenoage_advance,
    hd_log,
    telolength_log,
    all_of(params$hearing_var),
    all_of(biomarker_vars)
  )
```

### Set data types

```{r set_data_types}
df <- df %>%
  mutate(
    SEQN = as.character(SEQN),
    age_group = factor(age_group),
    sex = factor(sex),
    race = factor(race),
    education = factor(education),
    smoking = factor(smoking)
  )
```

## Clean up environment

From this point onward, we only keep what will actually be used for analysis. We remove df_raw and
df_unfiltered, but don't run `gc()`, as the objects being removed from the environment already
labels the RAM as free to be allocated, thus it's not needed to run garbage collection manually.

```{r cleanup_env}
rm(df_raw)
rm(df_unfiltered)
rm(df_no_missing)
rm(df_train)
rm(df_train_hd)
rm(df_train_raw)
rm(design_no_missing)
rm(descriptives_unfiltered)
rm(audiometry_cols)
```
