---
editor_options:
  markdown:
    wrap: 100
---

# Modelling

## K fold splitting

### Fold-maker function

We sort by PSU and stratum for reproducability. Then for each repeat, we assign fold labels to the
PSUs (the `mutate()` statement). In theses assignments, we assign labels to PSUs within each
stratum, assigning as many labels as there are PSUs within that stratum, so that each PSU gets one
fold label per repeat. The numbers that the PSUs can get are determined by the number of folds
(e.g., 1 or 2 for 2-fold or 1-5 for 5-fold). We sample without replacement so that we distribute all
PSUs across folds. E.g., with 2-fold 2-repeat CV, for a stratum with 3 PSUs, we could get 1, 2, 1
for repeat 1 and 2, 1, 1 for repeat 2 (indicating the fold number for the 3 PSUs). For a 5-fold
2-repeat CV for a stratum with 2 PSUs, we could get 5, 2 for repeat 1 and 3, 4 for repeat 2. Then
when fold 1 is the testing fold, no PSUs from this stratum are tested. This is the tradeoff. With
2-fold CV we have smaller training datasets, but with 5-fold or 10-fold CV we may get more variation
if performance per fold depends on which stratum is included. However, within each repeat, every
participant is still tested exactly once.

```{r make_folds}
make_repeated_folds <- function(data, k, repeats) {
  psus <- data %>%
    distinct(stratum, psu) %>%
    arrange(stratum, psu)

  for (i in seq_len(repeats)) {
    psus <- psus %>%
      mutate(
        !!paste0("fold_r", i) := sample(
          rep(sample.int(k), length.out = n())
        ),
        .by = stratum
      )
  }

  data %>%
    select(-starts_with("fold_r")) %>%
    left_join(psus, by = c("stratum", "psu"))
}

set.seed(2)
df <- make_repeated_folds(
  data = df,
  k = params$global_folds,
  repeats = params$global_repeats
)
```

## Formula building

### Formula builder function

The `k` determines the number of base functions (linear or non-linear).The parameter with the lowest
degrees of freedom (deg_free) determines the maximum basis dimensions (k in mgcv). Albumin has 26
distinct levels, so the maximal k value/number of basic functions, equivalent to the deg_free, is 25
if each level is in each fold. We must compute the deg_free for each fold and pick the lowest to
avoid errors. The GAM is a composite of all these underlying functions, with different weights for
each function. Max degrees of freedom are only computed once per variable and cached for retrieval.

- `k`: number of basis functions.
- `max_k`: sets k to the theoretical limit for each parameter (deg_free), note that for
  non-integers, like the bioage measures, this will be equivalent to the length of the dataframe
  minus one.
- `auto_limit_k`: if k exceeds the deg_free for a parameter (e.g., if k \> 25 for albumin), it will
  automatically be limited to the deg_free.
- `equal_wiggliness`: gives an individual id to each sum-to-zero interaction term when TRUE which
  ensures they have the same smoothing parameter. Setting to FALSE marginally improves fit but uses
  many degrees of freedom.
- `stepdown`: additionally creates formulas with age as predictor that each miss one covariate for a
  nested-CV comparison (i.e., compare, for example, the impact of adding 'sex' to a model that
  already includes other covariates).

```{r fun_formula_gen}
formula_generator <- function(
  data,
  formulas,
  covariates = NULL,
  k = 25,
  max_k = FALSE,
  auto_limit_k = FALSE,
  equal_wiggliness = TRUE,
  stepdown = FALSE,
  reverse = FALSE
) {
  # Initialize the repeat columns with folds and unique fold numbers in folds
  fold_columns <- grep("^fold_r", names(data), value = TRUE)
  folds <- 1:params$global_folds

  # Initialize a list for compiled formulas
  formula_list <- list()

  deg_free_store <- tibble(
    predictor = character(),
    deg_free = numeric()
  )

  for (formula in formulas) {
    # Unlist, split by comma, and get rid of white spaces for the current predictors string
    predictors <- trimws(
      unlist(
        strsplit(formula, ",")
      )
    )

    # Initialize a list for terms to compile formula from
    sm_terms <- list()

    # make a term for each predictor and load degrees of freedom if already computed for predictor
    for (predictor in predictors) {
      if (predictor %in% deg_free_store$predictor) {
        # Load degrees of freedom from cache if already computed else create list
        deg_free <- deg_free_store$deg_free[deg_free_store$predictor == predictor]
      } else {
        deg_free_list <- list()

        # Store the deg_free for each fold in each repeat in a list
        for (rep in fold_columns) {
          for (fold in folds) {
            deg_free_list <- append(
              deg_free_list,
              length(
                unique(
                  data[data[[rep]] != fold, ][[predictor]]
                )
              )
            )
          }
        }

        # Pick the smallest value in the list and subtract 1 to obtain degrees of freedom
        deg_free <- min(
          unlist(deg_free_list)
        ) - 1
        deg_free_store <- bind_rows(
          deg_free_store,
          tibble(predictor = predictor, deg_free = deg_free)
        )
      }

      # Overwrite k with max allowed k if desired
      k_formula <- if (max_k) deg_free else if (auto_limit_k && k > deg_free) deg_free else k

      if ((predictor == "age" && isFALSE(reverse)) || (predictor != "age" && isTRUE(reverse))) {
        # Construct age term
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(", predictor, ", k = ",
            k_formula,
            ")"
          )
        )

        if (!is.null(covariates)) {
          # construct Age * covariate terms
          for (i in seq_along(covariates)) {
            covariate <- covariates[i]
            if (isTRUE(equal_wiggliness)) {
              # Combine an age term with smooth terms with a unique ID for each
              sm_terms <- append(
                sm_terms,
                paste0(
                  "s(", predictor, ", ",
                  covariate,
                  ", bs = 'sz', k = ",
                  k,
                  ", id = ",
                  i,
                  ")"
                )
              )
            } else {
              # Combine an age term with smooth terms
              sm_terms <- append(
                sm_terms,
                paste0(
                  "s(", predictor, ", ",
                  covariate,
                  ", bs = 'sz', k = ",
                  k,
                  ")"
                )
              )
            }
          }
        }
      } else {
        # Construct predictor terms
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(",
            predictor,
            ", k = ",
            k_formula,
            ")"
          )
        )
      }
    }

    # Put predictor terms in formula and separate by +
    formula_list <- append(
      formula_list,
      paste(
        params$hearing_var,
        "~",
        paste(unlist(sm_terms), collapse = " + ")
      )
    )

    # Create covariate stepdown vars
    if (isTRUE(stepdown) && formula == "age") {
      # Age will already have been created by the previous loop
      deg_free <- deg_free_store$deg_free[deg_free_store$predictor == "age"]

      k_formula <- if (max_k) deg_free else if (auto_limit_k && k > deg_free) deg_free else k

      # Remove one covariate at a time, construct the formula, and add it to the formula list
      for (i in seq_along(covariates)) {
        sm_terms <- list()

        # Add age term
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(age, k = ",
            k_formula,
            ")"
          )
        )
        constrained_covariates <- covariates[-i]

        # Remove one covariate from formula at a time and create terms (with or without ID)
        for (j in seq_along(constrained_covariates)) {
          constrained_covariate <- constrained_covariates[j]
          if (isTRUE(equal_wiggliness)) {
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                constrained_covariate,
                ", bs = 'sz', k = ",
                k_formula,
                ", id = ",
                j,
                ")"
              )
            )
          } else {
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                constrained_covariate,
                ", bs = 'sz', k = ",
                k_formula,
                ")"
              )
            )
          }
        }

        formula_list <- append(
          formula_list,
          paste(
            params$hearing_var,
            "~",
            paste(
              unlist(sm_terms),
              collapse = " + "
            )
          )
        )
      }
    }
  }

  formula_list
}
```

### Extract variables from formula function

Helper used to turn a formula back into a list of variables.

```{r fun_extr_pred}
extract_predictors <- function(formula, collapse = "") {
  s_terms <- unlist(
    str_extract_all(
      formula,
      "s\\([^\\)]*\\)"
    )
  )
  keep <- !str_detect(
    s_terms,
    paste0(
      "\\b(",
      paste(covariate_vars, collapse = "|"),
      ")\\b"
    )
  )
  str_match(
    s_terms[keep],
    "^s\\(\\s*([^,\\)\\s]+)"
  )[, 2]
}
```

### Generate formulas bioage

```{r build_bioage_forms}
formulas_bioage <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c(
    "age",
    "age, phenoage",
    "age, kdm",
    "age, hd_log",
    if (isTRUE(params$telolength)) "age, telolength_log" else NULL
  ),
  covariates = covariate_interactions
)

tibble(
  Formulas = unlist(
    formulas_bioage
  )
) %>% construct_table(
  title = "Bioage Formulas"
)
```

### Generate formulas reverse bioage

```{r}
if (isFALSE(params$telolength)) {
  formulas_bioage_reverse <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = c(
      "phenoage",
      "kdm",
      "hd_log",
      "phenoage, age",
      "kdm, age",
      "hd_log, age"
    ),
    covariates = covariate_interactions,
    reverse = TRUE
  )

  tibble(
    Formulas = unlist(
      formulas_bioage_reverse
    )
  ) %>% construct_table(
    title = "Bioage Formulas"
  )
}
```

### Generate formulas biomarkers

```{r build_biomark_forms}
if (isFALSE(params$telolength)) {
  formulas_biomarkers <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = c(
      "age",
      paste0("age, ", biomarker_vars)
    ),
    covariates = covariate_interactions
  )

  tibble(
    Formulas = unlist(
      formulas_biomarkers
    )
  ) %>% construct_table(
    title = "Biomarker Formulas"
  )
}
```

### Generate formulas covariates

```{r build_conf_forms}
if (isFALSE(params$telolength)) {
  formulas_covariates <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = "age",
    covariates = covariate_vars,
    stepdown = TRUE
  )

  tibble(
    Formulas = unlist(
      formulas_covariates
    )
  ) %>% construct_table(
    title = "Covariate Formulas"
  )
}
```

## Repeated k-fold modeling

### Repeated k-fold validation function per age group

Fits a BAM() to the formulas generated splitting data into folds. Predictions are made using the
models and MSEs are calculated with these predictions per age group. A list is returned with a
tibble with results (MSE per fold per formula), predictions (each individual prediction made), and
optionally the actual models (uses very large amounts of memory when discrete = TRUE). When discrete
= TRUE, discretizes fitting using fREML (much faster but slightly worse fit). Covariate_analysis
does nothing to the fitting but returns names based on nested covariate models instead of the
regular predictors (which exclude covariates in the label). Model is fitted using normalized weights
`weights / mean(weights)`. Weights can have large numbers (thousands), which mgcv would recognize as
frequencies, making model estimates too small. By dividing by the mean we rescale the mean weight
to 1. Each fold gets its weights normalized to 1 so that mgcv does not overestimate variance. We can
correct for different total weight per test sets across folds later.

```{r fun_CV}
cross_validation <- function(
  data,
  formulas,
  folds = 10,
  repeats = 10,
  discrete = FALSE,
  covariate_analysis = FALSE,
  threads = 1
) {
  # Initialize age groups to stratify by and repeat column names that contain fold specification
  age_groups <- c("20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79")
  repeat_columns <- paste0("fold_r", seq_len(repeats))

  # Store predictor column names for prediction dataset
  predictor_columns <- unique(
    unlist(
      lapply(formulas, extract_predictors)
    )
  )

  # Set predictor combination names
  if (isTRUE(covariate_analysis)) {
    predictor_labels <- c(
      paste(c("age", covariate_vars), collapse = ", "),
      sapply(
        seq_along(covariate_vars),
        function(i) paste(c("age", covariate_vars[-i]), collapse = ", ")
      )
    )
  } else {
    predictor_labels <- unique(
      sapply(formulas, function(f) {
        paste(
          extract_predictors(f),
          collapse = ", "
        )
      })
    )
  }

  # Only select the columns that we need to reduce memory footprint per thread
  data <- data %>% select(
    all_of(
      c(
        "SEQN",
        "weights",
        params$hearing_var,
        "age_group",
        predictor_columns,
        covariate_vars,
        repeat_columns
      )
    )
  )

  splits <- expand.grid(
    formula = formulas,
    repeat_column = repeat_columns,
    fold = seq_len(folds)
  ) %>% mutate(
    predictors = map(formula, extract_predictors)
  )

  run_split <- function(split) {
    formula <- split[["formula"]]
    repeat_column <- split[["repeat_column"]]
    fold <- split[["fold"]]
    predictors <- split[["predictors"]]

    predictor_label <- predictor_labels[which(formulas == formula)]
    repeat_value <- as.numeric(
      sub(".*fold_r(\\d+).*", "\\1", repeat_column)
    )

    results_list <- list()
    predicted_list <- list()

    # Subset training and testing data
    train <- data %>%
      filter(
        .data[[repeat_column]] != fold
      ) %>%
      select(
        -starts_with("fold"),
        -all_of(setdiff(predictor_columns, predictors))
      )
    test <- data %>%
      filter(
        .data[[repeat_column]] == fold
      ) %>%
      select(
        -starts_with("fold"),
        -all_of(setdiff(predictor_columns, predictors))
      )

    # Fit GAM
    mod <- bam(
      as.formula(formula),
      data = train,
      family = gaussian(),
      weights = weights / mean(weights),
      method = if (isTRUE(discrete)) "fREML" else "REML",
      discrete = if (isTRUE(discrete)) TRUE else FALSE,
      nthreads = 1
    )

    map(age_groups, function(age_group_val) {
      # Subset testing and training data for age group
      train_subset <- train %>%
        filter(
          age_group == age_group_val
        )
      test_subset <- test %>%
        filter(
          age_group == age_group_val
        )

      # Use GAM to predict responses in current age group
      pred <- predict(mod, newdata = test_subset, type = "response")

      # Make a tibble with predictions and sample sizes  used in the model for predictions (n_train must be constructed by combining across all age groups)
      test_subset %>%
        as_tibble() %>%
        mutate(
          predictors = predictor_label,
          age_group = age_group_val,
          r = repeat_value,
          k = fold,
          n_train = nrow(train_subset),
          n_test = nrow(test_subset),
          predicted = pred
        )
    })
  }

  plan(multisession, workers = threads)
  cv_output <- future_apply(
    splits,
    MARGIN = 1,
    FUN = run_split,
    future.seed = TRUE,
    future.globals = list(
      data = data,
      params = params,
      age_groups = age_groups,
      discrete = discrete,
      predictor_labels = predictor_labels,
      predictor_columns = predictor_columns,
      formulas = formulas
    ),
    future.packages = c(
      "mgcv",
      "tidyverse",
      "tidytable"
    )
  )
  plan(sequential)

  cv_output %>%
    unlist(recursive = FALSE) %>%
    map_dfr(~.) %>%
    mutate(
      predictors = factor(
        predictors,
        levels = predictor_labels
      ),
      age_group = factor(age_group, levels = age_groups)
    )
}
```

### Generate model path function

```{r fun_gen_m_paths}
generate_model_path <- function(output_type) {
  folder_name <- paste0(
    params$hearing_var,
    if (isTRUE(params$telolength)) "_telolength" else NULL
  )

  file_name <- paste0(
    output_type,
    "_k=",
    params$global_k,
    "_discrete=",
    params$global_discrete,
    ".RDS"
  )

  model_path <- here("models", folder_name, file_name)

  dir.create(
    dirname(model_path),
    recursive = TRUE,
    showWarnings = FALSE
  )

  model_path
}
```

### Model bioage

Cross-validations rerun if an .RDS file with the corresponding name does not exist or if
load_model_results is FALSE. Results are stored if store_model_results is TRUE.

#### Bioage K-fold CV

```{r bioage_cv, warning=FALSE}
system.time({
  pred_bioage <- if (
    isTRUE(params$load_model_results) &&
      file.exists(generate_model_path("bioage_predictions"))
  ) {
    readRDS(generate_model_path("bioage_predictions"))
  } else {
    cross_validation(
      df,
      formulas_bioage,
      folds = params$global_folds,
      repeats = params$global_repeats,
      discrete = params$global_discrete,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      )
    )
  }
})
```

#### Store bioage modeling results

```{r store_bioage_res}
if (isTRUE(params$store_model_results)) {
  saveRDS(
    object = pred_bioage,
    file = generate_model_path("bioage_predictions"),
    compress = "xz"
  )
}
```

#### Reverse bioage K-fold CV

```{r bioage_cv_reverse, warning=FALSE}
if (isFALSE(params$telolength)) {
  system.time({
    pred_bioage_reverse <- if (
      isTRUE(params$load_model_results) &&
        file.exists(generate_model_path("bioage_reverse_predictions"))
    ) {
      readRDS(generate_model_path("bioage_reverse_predictions"))
    } else {
      cross_validation(
        df,
        formulas_bioage_reverse,
        folds = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        threads = min(
          params$n_threads,
          as.numeric(future::availableCores()) - 1
        )
      )
    }
  })
}
```

#### Store bioage reverse modeling results

```{r store_bioage_reverse_res}
if (isFALSE(params$telolength) && isTRUE(params$store_model_results)) {
  saveRDS(
    object = pred_bioage_reverse,
    file = generate_model_path("bioage_reverse_predictions"),
    compress = "xz"
  )
}
```

### Model biomarkers

#### Biomarkers K-fold CV

```{r biomarkers_cv, warning=FALSE}
if (isFALSE(params$telolength)) {
  system.time({
    pred_biomarkers <- if (
      isTRUE(params$load_model_results) &&
        file.exists(generate_model_path("biomarkers_predictions"))
    ) {
      readRDS(generate_model_path("biomarkers_predictions"))
    } else {
      cross_validation(
        df,
        formulas_biomarkers,
        folds = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        threads = min(
          params$n_threads,
          as.numeric(future::availableCores()) - 1
        )
      )
    }
  })
}
```

#### Store bioage modeling results

```{r store_biomarkers_cv}
if (isFALSE(params$telolength) && isTRUE(params$store_model_results)) {
  saveRDS(
    object = pred_biomarkers,
    file = generate_model_path("biomarkers_predictions"),
    compress = "xz"
  )
}
```

### Model covariates

#### Covariates K-fold CV

```{r covariates_cv, warning=FALSE}
if (isFALSE(params$telolength)) {
  system.time({
    pred_covariates <- if (
      isTRUE(params$load_model_results) &&
        file.exists(generate_model_path("covariates_predictions"))
    ) {
      readRDS(generate_model_path("covariates_predictions"))
    } else {
      cross_validation(
        df,
        formulas_covariates,
        folds = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        covariate_analysis = TRUE,
        threads = min(
          params$n_threads,
          as.numeric(future::availableCores()) - 1
        )
      )
    }
  })
}
```

#### Store covariates modeling results

```{r store_conf_res}
if (isFALSE(params$telolength) && isTRUE(params$store_model_results)) {
  saveRDS(
    object = pred_covariates,
    file = generate_model_path("covariates_predictions"),
    compress = "xz"
  )
}
```

## Model age rate

### Build GAM and linear age models in a list

When solely wanting to predict an outcome variable, the wiggliness penalty of a GAM prevents
overfitting, so k can be as large as we'd like. However, with a larger k, we have more parameters,
so if we take the derivative of the GAM and we have more parameters, each with its own uncertainty,
the total uncertainty will grow, even if the extra base functions don't contribute. Therefore, a
larger k past the number that captures all global trends decreases statistical power. This is a
problem when assessing the rate of aging. Therefore, we need to have a penalty for practically
useless base functions so that they shrink to 0. Default shrinkage solution still permits large
complexity despite the RMSE improving little over a simpler model.

We have three options for constructing the smooth. Option 1 was used in this analysis.

1.  `pta ~ age + I(age^2) + s(age, m = 3)`: we can create a smooth model with a linear term for age,
    a quadratic term for age, and a smooth term for age. The smooth term then also contains a
    penalized quadratic and unpenalized linear component, so the structure is complex and the model
    is highly concurve (multicolinear). By comparing between the quadratic model and GAM at specific
    ages, the null hypothesis is: at age $a$, the estimated rate of hearing loss is the same under
    the GAM and quadratic model. As a better alternative, we could also use `pta ~ s(age, m = 3)`
    and perform the same analysis but without the double linear and quadratic terms. We can estimate
    the standard error with bootstrapping. The delta method (using gratia) is used to calculate
    derivatives.
2.  `pta ~ age + I(age^2) + s(age, m = c(3,0))`: here we allow there to be a penalty null space in
    the first and second order polynomials (smooth starts penalizing third or higher-order
    polynomials) but we exclude the polynomials in the penalty null space from the smooth. This
    means that the smooth does not contain constants, linear, or quadratic components. The smooth
    term then represents deviation beyond quadratic. However, we then need to be able to estimate
    the smooth term on its own. With this model, we can identify the entire age trajectory well, but
    we cannot identify how much of the shape belongs to the quadratic coefficients and how much
    belongs to the smooth terms (higher order polynomials) because many of the smooth polynomials
    resemble the quadratic term (concurvity). So we can be certain that non-quadratic structure
    improves the model globally, we can estimate the entire derivative of the model, but we cannot
    estimate a smooth-only derivative for a particular age with any certainty. The uncertainty is
    thus not about what the model looks like but about whether a shape should be attributed to the
    quadratic or coefficients or to the smooth basis functions. The null hypothesis is: at age $a$,
    the rate of hearing loss in the complete GAM does not differ from the rate implied by the
    quadratic component within that same GAM.
3.  `pta ~ s(age, m = 3)`: here we have unpenalized linear and quadratic terms (in the null space),
    so the smooth contains a linear and a quadratic component. Here, if the data does not support a
    complex shape, the model shrinks towards a quadratic shape without competing with explicit
    terms. Because the smooth model still contains a quadratic and linear term, we can still use the
    approximate F-test to compare against a quadratic model and assume it is nested. We can still do
    pairwise comparisons. If a model is purely quadratic, then the 2nd order derivative is just a
    constant/intercept and the 3rd derivative is 0. Let's say we want to know whether a change in
    rate of HL during two years is quadratic. We can then calculate the 1st order derivative for a
    quadratic function, $f'(a) = \beta_1 + 2\beta_2 a$, in which age is $a$. We can calculate the
    difference between two years (interval size defined as $h$) as $f'(a) - f'(a - h)$. If the rate
    of HL in year 1 $a - h$ is 6dB/year and in year 2 is 9dB/year, then the difference is 3dB/year
    (the rate increased with 3 dB). Now, if the model is indeed quadratic, then the rate of HL
    increases linearly, so if we'd take a difference for a higher year, e.g. $f'(a + h) - f'(a)$,
    then the difference would also be 3 dB/year. If we'd take the 2nd difference (the difference
    between the two differences), we'd get 3 - 3 = 0. So, if the model is indeed quadratic, then the
    2nd difference of the derivative must equal zero, which we can write as
    $f'(a-h) - 2f'(a) + f'(a+h) = 0$. The null hypothesis is: around age $a$, the rate of hearing
    loss changes linearly with age, as required by a quadratic trajectory.

In option 1, we must account for covariance between full models with high multicolinearity.
Estimating this covariance is hard, but if we bootstrap differences between derivatives for each
age, then we include covariance. In option 2, we have a nice model structure, but we need to truly
isolate the smooth component at specific ages, which is not identifiable. In option 3, we have an
even simpler model structure and instead ask whether the total GAM (which is well-identifiable)
behaves quadratically at a given age range. Because option 2 is not well-identifiable, it's not
viable. Option 1 requires bootstrapping to estimate model covariances (computationally intensive).
Option 3 does not require bootstrapping but does require the coefficient-covariance matrix (vcov),
which we can compute, to calculate a difference. However, option 1 and option 3 ask fundamentally
different questions in pairwise comparisons. Option 1 asks whether there are any ages at which the
GAM derivative is different from the best-fitting quadratic derivative. Option 3 asks whether for
any age range $2h$, a local quadratic fit could explain the difference, as we're fitting a quadratic
model just to a 6-year period and seeing if the rate of change for 2 3-year periods is constant. The
downside of option 3 is that you don't get a true difference between the GAM and a single
best-fitting quadratic model, and you also manually need to pick the $h$ value. If $h$ is too small,
then changes in the rate of HL that perpetuate over a longer time may be missed, but if $h$ is too
big, we may include a linear portion with a nonlinear portion and miss the effect, or the 'binwidth'
falls just wrong. We can test different $h$ values but that would cause false positives. Moreover,
pairwise comparisons (albeit through confidence intervals) between a single best-fitting quadratic
model and a single GAM fits better with the (approximate) F-test. Therefore, option 1 was used.

Probability weights are used.

```{r aging_models}
models_age_rate <- list()

model_age_rate <- function(terms) {
  bam(
    reformulate(terms, response = params$hearing_var),
    data = df,
    weights = weights / mean(weights),
    method = "REML",
    family = gaussian()
  )
}

models_age_rate <- append(
  models_age_rate,
  list(
    linear = model_age_rate("age"),
    quadratic = model_age_rate(c("age", "I(age^2)")),
    smooth = model_age_rate(paste0("s(age, k = ", params$age_rate_k, ", m = 3)"))
  )
)

set.seed(1)
k.check(models_age_rate$smooth)
```

### Build covariate age rate models

bs = "sz" ensures that the individual factor levels sum to zero, so that we can estimate both the
main effect of age as well as the levels of individual strata. However, we don't need to estimate
the main effect of age because we already assess this in a separate analysis. The purpose of these
models is solely to visualize the per-stratum predictions, so it's fine if the main effect of age
gets absorbed.

```{r aging_covariate_models}
models_covariates <- list()
models_covariates <- append(
  models_covariates,
  list(
    smooth_sex = model_age_rate(
      terms = c(
        paste0("s(age, k = ", params$global_k, ")"),
        paste0("s(age, sex, bs = 'sz', k = ", params$global_k, ", id = 1)")
      )
    ),
    smooth_smoking = model_age_rate(
      terms = c(
        paste0("s(age, k = ", params$global_k, ")"),
        paste0("s(age, smoking, bs = 'sz', k = ", params$global_k, ", id = 1)")
      )
    ),
    smooth_race = model_age_rate(
      terms = c(
        paste0("s(age, k = ", params$global_k, ")"),
        paste0("s(age, race, bs = 'sz', k = ", params$global_k, ", id = 1)")
      )
    ),
    smooth_education = model_age_rate(
      terms = c(
        paste0("s(age, k = ", params$global_k, ")"),
        paste0("s(age, education, bs = 'sz', k = ", params$global_k, ", id = 1)")
      )
    )
  )
)
```

## Optional tests

### (Optional) Inspect folds

We check if each fold number is assigned equal amounts of times in two repeats for each age group.
Thereafter, we save the row number and plot the fold number of points (with some jitter to spread
things out) to check if bands appear homogenous. We know it should be homogenous in the y-axis
within a band because the jitter is random. If there are similarities between bands, then the
shuffle isn't randomized for each repeat. If the bands don't appear of similar intensity over the
x-axis, then the number assignment is not random.

```{r inspect_folds, fig.height = 10, fig.width = 10}
if (isTRUE(params$run_optional_chunks)) {
  df %>%
    group_by(age_group) %>%
    mutate(total_fold1 = n()) %>%
    group_by(age_group, fold_r1, total_fold1) %>%
    summarize(
      n_fold1 = n(),
      .groups = "drop"
    ) %>%
    rename("fold" = fold_r1) %>%
    left_join(
      df %>%
        group_by(age_group) %>%
        mutate(total_fold2 = n()) %>%
        group_by(age_group, fold_r2, total_fold2) %>%
        summarize(
          n_fold2 = n(),
          .groups = "drop"
        ) %>%
        select(age_group, fold_r2, total_fold2, n_fold2, ) %>%
        rename("fold" = fold_r2),
      by = c("age_group", "fold")
    ) %>%
    kable(
      "html",
      caption = "Inspect Fold Number Frequencies"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")

  p <- df %>%
    mutate(row_index = row_number()) %>%
    pivot_longer(
      cols = starts_with("fold_r"),
      names_to = "r",
      values_to = "fold"
    ) %>%
    ggplot(aes(x = row_index, y = fold)) +
    geom_point(
      size = 1,
      alpha = 0.1,
      color = "grey80",
      position = position_jitter(width = 0, height = 0.4)
    ) +
    facet_wrap(~r) +
    theme_few()

  plot(p)
}
```

Fold numbers are consistent between groups and correctly split upon multiple repeats. each fold
number has the same (within 1) frequency, so there are no imbalances between folds and repeats. Plot
shows no non-random changes in intensity over the X dimension and random dissimilarities between
bands in the Y dimension (indicating repeats have different samples).

### (Optional) Compare training support for two and five PSU folds

Check whether observed age-group and covariate combinations remain represented in every training
split. Compare training counts, distinct ages, and age-range coverage without fitting models or
examining outcomes.

```{r compare_psu_fold_support}
if (isTRUE(params$run_optional_chunks)) {
  compare_fold_support <- function(data, psus, repeats) {
    map_dfr(c(2, 5, 10), function(folds) {
      set.seed(2)
      fold_data <- make_repeated_folds(
        data = data,
        k = folds,
        repeats = repeats
      )

      map_dfr(unlist(covariate_interactions), function(term) {
        cells <- fold_data %>%
          unite(
            "cell",
            all_of(trimws(strsplit(term, ",")[[1]])),
            sep = " / "
          )

        reference <- cells %>%
          summarize(
            age_min = min(age),
            age_max = max(age),
            .by = c(age_group, cell)
          )

        map_dfr(seq_len(repeats), function(r) {
          map_dfr(seq_len(folds), function(fold) {
            training <- cells %>%
              filter(.data[[paste0("fold_r", r)]] != fold) %>%
              summarize(
                n_train = n(),
                unique_ages = n_distinct(age),
                train_age_min = min(age),
                train_age_max = max(age),
                .by = c(age_group, cell)
              )

            reference %>%
              left_join(training, by = c("age_group", "cell")) %>%
              mutate(
                folds = folds,
                term = term,
                r = r,
                fold = fold,
                n_train = replace_na(n_train, 0L),
                unique_ages = replace_na(unique_ages, 0L),
                lower_age_loss = train_age_min - age_min,
                upper_age_loss = age_max - train_age_max
              )
          })
        })
      })
    })
  }

  compare_fold_support(
    data = df,
    repeats = params$global_repeats
  ) %>%
    summarize(
      min_train_n = min(n_train),
      min_unique_ages = min(unique_ages),
      missing_cell_splits = sum(n_train == 0),
      max_lower_age_loss = max(c(0, lower_age_loss), na.rm = TRUE),
      max_upper_age_loss = max(c(0, upper_age_loss), na.rm = TRUE),
      .by = c(folds, term, age_group)
    ) %>%
    arrange(term, age_group) %>%
    construct_table(title = "Two versus five vs ten PSU folds: training support")
}
```

### (Optional) Check covariate two-way interactions vs offsets

We want to get a sense which covariates improve the model and which interact with age.

```{r assess_covariates}
if (isTRUE(params$run_optional_chunks)) {
  assess_covariates <- function(df, covariates, k) {
    f0 <- paste0(params$hearing_var, " ~ s(age, k = ", k, ")")
    m0 <- bam(as.formula(f0),
      data = df,
      weights = weights / mean(weights),
      method = "REML"
    )

    results <- data.frame()

    for (covariate in covariates) {
      f1 <- paste0(
        params$hearing_var,
        " ~ ",
        covariate,
        " + s(age, k = ",
        k,
        ")"
      )
      f2 <- paste0(
        params$hearing_var,
        " ~ s(age, k = ",
        k,
        ") + s(age, ",
        covariate,
        ", bs = 'sz', k = ",
        k,
        ")"
      )

      m1 <- bam(
        as.formula(f1),
        data = df,
        weights = weights / mean(weights),
        method = "REML",
        family = gaussian()
      )
      m2 <- bam(
        as.formula(f2),
        data = df,
        weights = weights / mean(weights),
        method = "REML",
        family = gaussian()
      )

      aov <- as.data.frame(
        anova(m0, m1, m2)
      )
      aic <- as.data.frame(
        AIC(m0, m1, m2)
      )
      bic <- as.data.frame(
        BIC(m0, m1, m2)
      )

      results <- rbind(
        results,
        cbind(
          tibble(
            covariate = covariate,
            formula = c(f0, f1, f2)
          ),
          aic,
          bic,
          aov
        )
      )
    }

    return(results)
  }

  assess_covariates(df, covariates = covariate_vars, k = params$global_k)
}
```

**We see that for each covariate, the interaction model fits better than the intercept-only model,
so we should use interaction terms instead of intercepts for covariates.**

### (Optional) Investigate additional information

First we see if the covariates are correlated (e.g., smoking is more prevalent in males than females
so sex is partly coded in smoking), to see if all covariates actually contribute.

```{r compare_added_info}
if (isTRUE(params$run_optional_chunks)) {
  compare_covariates <- function(df, k) {
    m0 <- bam(
      reformulate(
        "s(age, k = k)",
        response = params$hearing_var
      ),
      data = df,
      weights = weights / mean(weights),
      method = "REML",
      family = gaussian()
    )
    m1 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, sex, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      weights = weights / mean(weights),
      method = "REML",
      family = gaussian()
    )
    m2 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, sex, bs = 'sz', k = k)",
          "s(age, race, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      weights = weights / mean(weights),
      method = "REML",
      family = gaussian()
    )
    m3 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, sex, bs = 'sz', k = k)",
          "s(age, race, bs = 'sz', k = k)",
          "s(age, smoking, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      weights = weights / mean(weights),
      method = "REML",
      family = gaussian()
    )
    m4 <- bam(
      reformulate(c(
        "s(age, k = k)",
        "s(age, sex, bs = 'sz', k = k)",
        "s(age, race, bs = 'sz', k = k)",
        "s(age, smoking, bs = 'sz', k = k)",
        "s(age, education, bs = 'sz', k = k)"
      ), response = params$hearing_var),
      data = df,
      weights = weights / mean(weights),
      method = "REML",
      family = gaussian()
    )

    print(
      cbind(
        as.data.frame(
          AIC(m0, m1, m2, m3, m4)
        ),
        as.data.frame(
          BIC(m0, m1, m2, m3, m4)
        ),
        as.data.frame(
          anova(m0, m1, m2, m3, m4)
        )
      )
    )
    print(
      summary(m4)
    )
    plot(m4)
  }

  compare_covariates(df, k = params$global_k)
}
```

**Each factor significantly improves the model and seems to interact with age in some way**

### (Optional) Check shared wiggliness

We could insist that different smooths have the same amount of smoothness, i.e., that there's no
difference in the wiggliness of the GAM between men and women. The plots from the previous chunk
(where different wiggliness was permitted) suggest that there may be equal amount of wiggliness, so
that each smooth has the same basis functions but with different coefficients. There's no specific
reason to assume that wiggliness is the same in many cases, so if aiming to maximize model
flexibility/predictive power, we should leave this unconstrained, but we do eat up a lot of degrees
of freedom and may be 'wasting' the data estimating these global trends. It may also be better to
spend the processing time on a higher k rather than estimating smooth parameters.

See the link below for more info on constructing interaction terms:
<https://stats.stackexchange.com/questions/637423/conceptual-interpretation-of-bs-fs-and-by-term-in-gam>.

```{r compare_wiggliness}
if (isTRUE(params$run_optional_chunks)) {
  compare_wiggliness <- function(df, k, discrete = FALSE) {
    m1 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          ' ~
      s(age, k = k) +
      s(age, sex, bs = "sz", k = k, id = 1) +
      s(age, race, bs = "sz", k = k, id = 2) +
      s(age, smoking, bs = "sz", k = k, id = 3) +
      s(age, education, bs = "sz", k = k, id = 4)'
        )
      ),
      data = df,
      weights = weights / mean(weights),
      method = if (isTRUE(discrete)) "fREML" else "REML",
      discrete = if (isTRUE(discrete)) TRUE else FALSE,
      family = gaussian()
    )

    m2 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          ' ~
      s(age, k = k) +

      s(age, sex, bs = "sz", k = k) +
      s(age, race, bs = "sz", k = k) +
      s(age, smoking, bs = "sz", k = k) +
      s(age, education, bs = "sz", k = k)'
        )
      ),
      data = df,
      weights = weights / mean(weights),
      method = if (isTRUE(discrete)) "fREML" else "REML",
      discrete = if (isTRUE(discrete)) TRUE else FALSE,
      family = gaussian()
    )

    print(
      cbind(
        as.data.frame(
          AIC(m1, m2)
        ),
        as.data.frame(
          BIC(m1, m2)
        ),
        as.data.frame(
          anova(m1, m2)
        )
      )
    )
  }

  compare_wiggliness(
    df,
    k = params$global_k,
    discrete = params$global_discrete
  )
}
```

**The free wiggliness model takes about 15 times longer to compute and barely has a drop in residual
deviance. The improvement in fit is barely significant, so enforce shared smoothing parameters.**

### (Optional) Check factor three-way interactions

```{r conf_interactions}
if (isTRUE(params$run_optional_chunks)) {
  compare_covariate_interactions <- function(df, k, discrete = FALSE, return_model = FALSE) {
    # Pull covariates, cross them, keep copy with var1 being first (alphabetical), paste by comma
    terms <- c(
      covariate_vars,
      crossing(var1 = covariate_vars, var2 = covariate_vars) %>%
        filter(var1 < var2) %>%
        mutate(pair = paste(var1, var2, sep = ", ")) %>%
        pull(pair)
    )

    # Combine an age term with smooth terms with a unique ID for each
    smooth_terms <- c(
      paste0("s(age, k = ", k, ")"),
      paste0(
        "s(",
        "age, ",
        terms,
        ", bs = 'sz', k = ",
        k,
        ", id = ",
        seq_along(terms),
        ")"
      )
    )


    m0 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          " ~ ",
          paste(smooth_terms, collapse = " + ")
        )
      ),
      data = df,
      weights = weights / mean(weights),
      family = gaussian(),
      method = if (isTRUE(discrete)) "fREML" else "REML",
      discrete = if (isTRUE(discrete)) TRUE else FALSE
    )

    print(summary(m0))

    if (return_model) {
      m0
    }
  }

  system.time({
    compare_covariate_interactions(
      df,
      k = params$global_k,
      discrete = params$global_discrete
    )
  })
}
```

We see that `age*sex*race` and `age*education*sex` are a significant three-way interaction but it
takes 3 times longer to compute. Bam() speeds things up enough over gam() that this processing cost
is acceptable.

### (Optional) Check three-way interactions in all 100 folds

```{r check_signif_interactions_per_fold}
if (isTRUE(params$run_optional_chunks)) {
  check_terms <- function(
    data,
    covariate_interactions,
    k,
    folds,
    repeats,
    discrete = FALSE,
    return_model = FALSE,
    threads = 1,
    skip = FALSE
  ) {
    if (isFALSE(skip)) {
      smooth_terms <- c(
        paste0("s(age, k = ", k, ")"),
        paste0(
          "s(",
          "age, ",
          covariate_interactions,
          ", bs = 'sz', k = ",
          k,
          ", id = ",
          seq_along(covariate_interactions),
          ")"
        )
      )

      formula <- as.formula(
        paste0(
          params$hearing_var,
          " ~ ",
          paste(smooth_terms, collapse = " + ")
        )
      )

      repeat_columns <- paste0("fold_r", seq_len(repeats))

      included_vars <- c("age", params$hearing_var, covariate_vars, "weights")

      splits <- expand.grid(
        repeat_column = repeat_columns,
        fold = seq_len(folds)
      ) %>% mutate(
        fold_n = seq_len(folds * repeats)
      )

      run_split <- function(split) {
        repeat_column <- split[["repeat_column"]]
        fold <- split[["fold"]]

        repeat_value <- as.numeric(
          sub(".*fold_r(\\d+).*", "\\1", repeat_column)
        )

        train <- data %>%
          filter(
            .data[[repeat_column]] != as.integer(fold)
          ) %>%
          select(
            all_of(included_vars)
          )

        mod <- bam(
          formula,
          data = train,
          weights = weights / mean(weights),
          family = gaussian(),
          method = if (isTRUE(discrete)) "fREML" else "REML",
          discrete = if (isTRUE(discrete)) TRUE else FALSE
        )

        summary_table <- as.data.frame(summary(mod)$s.table)
        summary_table[["predictor"]] <- rownames(summary_table)
        summary_table[["fold"]] <- as.integer(split[["fold_n"]])

        summary_table
      }

      plan(multisession, workers = threads)
      cv_output <- future_apply(
        splits,
        MARGIN = 1,
        FUN = run_split,
        future.seed = TRUE,
        future.globals = list(
          data = data,
          discrete = discrete,
          included_vars = included_vars,
          formula = formula
        ),
        future.packages = c(
          "mgcv",
          "tidyverse",
          "tidytable"
        )
      )
      plan(sequential)

      output_summary <- bind_rows(cv_output) %>%
        filter(`p-value` <= 0.05) %>%
        group_by(fold) %>%
        summarize(n = n())

      if (all(output_summary$n == 7)) {
        message("All folds have equal numbers of significant variables")
      } else {
        message("Folds have unequal numbers of significant variables")
        print(bind_rows(cv_output) %>%
          filter(`p-value` > 0.05))
      }
    }
  }

  system.time({
    check_terms(
      data = df,
      covariate_interactions = unlist(covariate_interactions),
      k = params$global_k,
      folds = params$global_folds,
      repeats = params$global_repeats,
      discrete = params$global_discrete,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      ),
      skip = TRUE
    )
  })
}
```

**One fold does not have smoking as a significant term, so there is minor information leakage**.

### (Optional) Validate excluded terms

```{r check_signif_excluded_terms}
if (isTRUE(params$run_optional_chunks)) {
  verify_covariate_interactions <- function(df, k, discrete = FALSE, return_model = FALSE, covariate_interactions) {
    # Combine an age term with smooth terms with a unique ID for each
    smooth_terms <- c(
      paste0("s(age, k = ", k, ")"),
      paste0(
        "s(",
        "age, ",
        covariate_interactions,
        ", bs = 'sz', k = ",
        k,
        ", id = ",
        seq_along(covariate_interactions),
        ")"
      )
    )

    m0 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          " ~ ",
          paste(smooth_terms, collapse = " + ")
        )
      ),
      data = df,
      weights = weights / mean(weights),
      family = gaussian(),
      method = if (isTRUE(discrete)) "fREML" else "REML",
      discrete = if (isTRUE(discrete)) TRUE else FALSE
    )

    print(summary(m0))
    plot(m0)

    if (return_model) {
      m0
    }
  }

  system.time({
    verify_covariate_interactions(
      df,
      k = params$global_k,
      discrete = params$global_discrete,
      covariate_interactions = unlist(covariate_interactions)
    )
  })
}
```

### (Optional) Inspect k-values

To assess k-value, we fit a model on the whole dataset using the formula lists, build a list of
models, make a helper function to run the k.check function from mgcv, then turn it in a dataframe
and return the output. The k-value determines the number of base functions that the GAM is
constructed from. A GAM adds these different base functions together, some of which capture linear
trends, some of which are just splines or polynomials with a bump in one region. The GAM gives
different weights to each model to get as good a fit as possible, but a GAM also measures wiggliness
and gives a penalty to wiggliness. This is what makes it generalized. I.e., to perform well in
general, even to unseen data. Therefore, a GAM optimizes to get as good a fit as possible with as
little wiggliness as possible. Because of this wiggliness penalty, adding more base functions to the
GAM doesn't make it overfit. Therefore, k should be as large as possible within computation power or
degrees of freedom. If k isn't large enough, there will be patterns left in the residuals.
Therefore, by measuring patterns in the residuals, we can assess if k is large enough. We want to
capture all global patterns without fitting to any noise. This is what we assess when we look at
k-index.

Models are fit on whole dataset.

```{r inspect_k}
if (isTRUE(params$run_optional_chunks)) {
  inspect_k_values <- function(data, formulas, discrete = FALSE, threads = 1) {
    run_formula <- function(formula) {
      set.seed(2)
      mod <- bam(
        as.formula(formula),
        data = data,
        weights = weights / mean(weights),
        family = gaussian(),
        method = if (isTRUE(discrete)) "fREML" else "REML",
        discrete = if (isTRUE(discrete)) TRUE else FALSE
      )
      k_result <- k.check(mod)
      predictors <- paste(
        extract_predictors(formula),
        collapse = ", "
      )
      as.data.frame(k_result) %>%
        rownames_to_column(var = "Predictors") %>%
        mutate(model = predictors) %>%
        filter(!is.na(`k-index`))
    }

    plan(multisession, workers = threads)
    k_list <- future_lapply(
      formulas,
      FUN = run_formula,
      future.seed = TRUE,
      future.globals = list(
        data = data,
        discrete = discrete,
        extract_predictors = extract_predictors,
        covariate_vars = covariate_vars
      ),
      future.packages = c(
        "mgcv",
        "tidyverse",
        "tidytable"
      )
    )
    plan(sequential)

    bind_rows(k_list)
  }

  inspect_k_values(df,
    formulas_bioage,
    discrete = params$global_discrete,
    threads = min(
      params$n_threads,
      as.numeric(future::availableCores()) - 1
    )
  ) %>%
    construct_table(title = "Inspect K Bioage")
  if (isFALSE(params$telolength)) {
    inspect_k_values(df,
      formulas_biomarkers,
      discrete = params$global_discrete,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      )
    ) %>%
      construct_table(title = "Inspect K Biomarkers")
  }
}
```

### (Optional) K stability analysis

We observe at which K the model stops improving by fitting models with different K values. The
largest of the two smallest Ks that is close to the ceiling for the HF-PTA and LF-PTA is used for
both sets of models.

```{r assess_k_stability}
if (isTRUE(params$run_optional_chunks)) {
  assess_stability_k <- function(data, threads = 1) {
    run_k <- function(k) {
      set.seed(1)
      m <- model_age_rate(paste0(
        "s(age, k = ",
        k,
        ", m = 3)"
      ))
      k.check(m) %>%
        as_tibble() %>%
        mutate(k = k)
    }

    plan(multisession, workers = threads)
    k_list <- future_lapply(
      seq(4, 30, 1),
      FUN = run_k,
      future.seed = TRUE,
      future.globals = TRUE,
      future.packages = c(
        "mgcv",
        "tidyverse",
        "tidytable"
      )
    )
    plan(sequential)

    k_df <- bind_rows(k_list)

    p <- ggplot(k_df, aes(x = k, y = edf)) +
      geom_line() +
      geom_point(size = 1.5) +
      geom_point(data = k_df %>% filter(k == params$age_rate_k), color = "red", shape = 1, size = 3) +
      scale_x_continuous(breaks = c(3, 5, 10, 15, 20, 25, 30)) +
      scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5)) +
      coord_cartesian(ylim = c(0, 5.2)) +
      labs(x = "K (Smooth Basis Dimensions)", y = "Effective degrees of freedom") +
      theme_few() +
      theme(
        panel.grid.major.x = element_line(linewidth = 0.1, color = "grey50"),
        panel.border = element_part_rect(
          colour = "grey30",
          side = "lb",
          linewidth = 0.5
        ),
        plot.margin = margin(1, 1, 0, 1, unit = "pt"),
        axis.text.x = element_markdown(
          "sans",
          size = 8
        ),
        axis.text.y = element_markdown(
          "sans",
          size = 8
        ),
        axis.title.x = element_markdown(
          "sans",
          size = 9,
          lineheight = 1.1
        ),
        axis.title.y = element_markdown(
          "sans",
          size = 9,
          lineheight = 1.1,
          color = if (params$hearing_var == "pta_lf") "white" else "black"
        )
      )

    p <- ggarrange(
      p,
      labels = if (params$hearing_var == "pta_hf") "A" else "B",
      label.x = -0.02,
      label.y = 1.01
    )

    p +
      draw_label(
        if (params$hearing_var == "pta_hf") "PTA-HF" else "PTA-LF",
        x = 0.5,
        y = 0.99,
        hjust = 0.5,
        vjust = 1,
        size = 9
      )
  }

  sameplot(
    assess_stability_k(
      data = df,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      )
    ),
    width = params$single_column_width,
    height = 3,
    res = 1200,
    save = params$save_figures,
    file = here(
      "output",
      "figures",
      if (isTRUE(params$telolength)) {
        paste0(params$hearing_var, "_telolength_log")
      } else {
        params$hearing_var
      },
      paste0("k stability analysis", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".tif")
    )
  )
}
```
