---
editor_options:
  markdown:
    wrap: 100
---

# Data Processing

## Build predictions dataframe age rate

We sample 1 point for each age and predict the PTA using the smooth-only model and generate
simultaneous confidence intervals with 'mvt'. We could ask for a single point and ask where the true
mean would be with 95% certainty, but if we'd apply that to all 60 points, there's a large chance
that somewhere we get a false positive. With the simultaneous interval or with multivariate t
distribution error correction, we can say that we're 95% sure that the **entire** derivative in the
true population that was sampled from (whole US population) is within the band.

To calculate the quadratic derivative, we can either calculate the slope at each point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same. However, using
emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

### Main models

```{r predict_aging}
predict_age_rate <- function(
  df,
  model,
  predictor = "age",
  type = c("predict", "predict_derivative")
) {
  ages <- seq(
  min(df[[predictor]], na.rm = TRUE),
  max(df[[predictor]], na.rm = TRUE),
  by = 1
)

  formula <- as.formula(
    paste("~", predictor)
  )

  if (type == "predict") {
    em <- emmeans(
      model,
      formula,
      at = list(age = ages),
      type = "response",
      data = df
    )
  } else {
    em <- emtrends(
      model,
      formula,
      var = predictor,
      at = list(age = ages),
      data = df
    )
  }

  predictions <- tibble(
    confint(em, adjust = "mvt")
  )

  if (type == "predict") {
    predictions %>%
      rename(
        "estimate" = emmean,
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  } else {
    predictions %>%
      rename(
        "derivative" = !!paste0(predictor, ".trend"),
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  }
}

pred_age_rate <- list()

pred_age_rate <- append(
  pred_age_rate,
  list(
    linear = predict_age_rate(
      df,
      models_age_rate$linear,
      type = "predict"
    ),
    linear_derivative = predict_age_rate(
      df,
      models_age_rate$linear,
      type = "predict_derivative"
    ),
    quadratic = predict_age_rate(
      df,
      models_age_rate$quadratic,
      type = "predict"
    ),
    quadratic_derivative = predict_age_rate(
      df,
      models_age_rate$quadratic,
      type = "predict_derivative"
    ),
    smooth = predict_age_rate(
      df,
      models_age_rate$smooth,
      type = "predict"
    ),
    derivative_smooth = predict_age_rate(
      df,
      models_age_rate$smooth,
      type = "predict_derivative"
    )
  )
)
```

### Quadratic-smooth derivative difference

To calculate the difference between the smooth and quadratic derivatives, we can't simply sum their
uncertainties, because need to account for their covariance.

We take a nonparametric bootstrap, meaning we sample from the actual data. Within each replicate, we
standardize the mean weight to 1, so that each replicate is equal in information, even if it happens
to have a lower total weight, because mgcv is sensitive to the scale of weights. We take samples
based on the primary sampling units within strata, rather than individuals. Within each replicate,
we fit a GAM and quadratic model to the same data, calculate the derivative/coefficient for each
age, and subtract the quadratic coefficient from the smooth to obtain a difference. If the sample at
a given age is randomly higher and this bump raises the GAM's derivative but drops the derivative of
the quadratic (if, for example, the bump leads to a higher intercept or linear component with lower
quadratic component to compensate, i.e., the derivative rotates), then there is negative covariance
and we get a higher total variance for that age. This means that both within-model variance and
across-model covariance is accounted for when we compute the standard error from the bootstrap.

```{r}
nonparametric_difference_bootstrap <- function(df, n_boot = 2000, threads = 1) {
  ages <- seq(
    min(df[["age"]], na.rm = TRUE),
    max(df[["age"]], na.rm = TRUE),
    by = 1
  )

  set.seed(2)

  plan(multisession, workers = threads)
  
  bootstrap_derivative_diff <- simplify2array(future_lapply(
    seq_len(n_boot),
    function(i) {
      indices <- sample.int(nrow(df), replace = TRUE)
      data_subset <- df[indices, ]
      weights_subset <- df$weights[indices]

      fit <- function(terms) {
        gam(
          reformulate(terms, response = params$hearing_var),
          data = data_subset,
          weights = weights_subset / mean(weights_subset),
          method = "REML",
          family = gaussian()
        )
      }

      trend <- function(model) {
        as.data.frame(
          emtrends(
            model, ~age,
            var = "age",
            at = list(age = ages),
            data = data_subset
          )
        )$age.trend
      }

      smooth_coefs <- trend(fit(paste0("s(age, k = ", params$age_rate_k, ", m = 3)")))
      parametric_coefs <- trend(fit(c("age", "I(age^2)")))

      smooth_coefs - parametric_coefs
    },
    future.seed = TRUE,
    future.globals = TRUE,
    future.packages = c(
      "emmeans",
      "mgcv",
      "tidytable"
    )
  ))
  
  plan(sequential)

  SE_adj <- apply(bootstrap_derivative_diff, 1, sd)

  t_crit_max_t <- unname(quantile(
    apply(
      abs(sweep(bootstrap_derivative_diff, 1, rowMeans(bootstrap_derivative_diff), "-") / SE_adj),
      2,
      max
    ),
    0.95
  ))

  cbind(SE_adj, t_crit_max_t)
}

system.time(
  pred_age_rate$derivative_smooth <- cbind(
    pred_age_rate$derivative_smooth,
    df %>% nonparametric_difference_bootstrap(
      n_boot = params$n_boot,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      )
    )
  ) %>%
    mutate(
      derivative_adj = derivative - pred_age_rate$quadratic_derivative$derivative,
      ci_lower_adj = derivative_adj - t_crit_max_t * SE_adj,
      ci_upper_adj = derivative_adj + t_crit_max_t * SE_adj
    )
)
```

### Covariates

We predict the `hearing_var` for each age in the dataset (20-79) for each stratum level.

```{r pred_covariate_pta}
predict_age_rate_covariate <- function(
  df,
  model,
  predictor = "age",
  covariate
) {
  ages <- setNames(
    list(
      seq(
        min(df[[predictor]]),
        max(df[[predictor]]),
        by = 1
      )
    ),
    predictor
  )

  em <- emmeans(
    model,
    as.formula(
      paste0("~ ", predictor, " + ", covariate)
    ),
    at = ages,
    by = covariate,
    type = "response",
    data = df
  )

  predictions <- as.data.frame(
    confint(em)
  )

  predictions %>%
    rename(
      "estimate" = emmean,
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}

pred_covariates_full <- list()

pred_covariates_full <- append(
  pred_covariates_full,
  list(
    smooth_sex = predict_age_rate_covariate(
      df,
      models_covariates$smooth_sex,
      covariate = "sex"
    ),
    smooth_smoking = predict_age_rate_covariate(
      df,
      models_covariates$smooth_smoking,
      covariate = "smoking"
    ),
    smooth_race = predict_age_rate_covariate(
      df,
      models_covariates$smooth_race,
      covariate = "race"
    ),
    smooth_education = predict_age_rate_covariate(
      df,
      models_covariates$smooth_education,
      covariate = "education"
    )
  )
)
```
