---
editor_options:
  markdown:
    wrap: 100
---

# Data Processing

## Build predictions dataframe age rate

We sample 1 point for each age and predict the PTA using the smooth-only model and generate
simultaneous confidence intervals with 'mvt'. We could ask for a single point and ask where the true
mean would be with 95% certainty, but if we'd apply that to all 60 points, there's a large chance
that somewhere we get a false positive. With the simultaneous interval or with multivariate t
distribution error correction, we can say that we're 95% sure that the **entire** derivative in the
true population that was sampled from (whole US population) is within the band.

To calculate the quadratic derivative, we can either calculate the slope at each point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same. However, using
emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

### Main models

```{r predict_aging}
predict_age_rate_parametric <- function(
  df,
  model,
  predictor = "age",
  type = c("predict", "predict_derivative")
) {
  ages <- seq(
    min(df[[predictor]], na.rm = TRUE),
    max(df[[predictor]], na.rm = TRUE),
    by = 1
  )

  formula <- as.formula(
    paste("~", predictor)
  )

  if (type == "predict") {
    em <- emmeans(
      model,
      formula,
      at = list(age = ages),
      type = "response",
      data = df
    )
  } else if (type == "predict_derivative") {
    em <- emtrends(
      model,
      formula,
      var = predictor,
      at = list(age = ages),
      data = df
    )
  }

  predictions <- tibble(
    confint(em, adjust = "mvt")
  )

  if (type == "predict") {
    predictions %>%
      rename(
        "estimate" = emmean,
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  } else if (type == "predict_derivative") {
    predictions %>%
      rename(
        "derivative" = !!paste0(predictor, ".trend"),
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  }
}

pred_age_rate <- list()

pred_age_rate <- append(
  pred_age_rate,
  list(
    linear = predict_age_rate_parametric(
      df,
      models_age_rate$linear,
      type = "predict"
    ),
    linear_derivative = predict_age_rate_parametric(
      df,
      models_age_rate$linear,
      type = "predict_derivative"
    ),
    quadratic = predict_age_rate_parametric(
      df,
      models_age_rate$quadratic,
      type = "predict"
    ),
    quadratic_derivative = predict_age_rate_parametric(
      df,
      models_age_rate$quadratic,
      type = "predict_derivative"
    ),
    smooth = confint(
      models_age_rate$smooth,
      parm = "s(age)",
      data = data.frame(age = seq(
        min(df[["age"]], na.rm = TRUE),
        max(df[["age"]], na.rm = TRUE),
        by = 1
      )),
      interval = "simultaneous",
      shift = TRUE
    ) %>%
      select(
        age,
        estimate = .estimate,
        SE = .se,
        ci_lower = .lower_ci,
        ci_upper = .upper_ci
      ),
    derivative_smooth = derivatives(
      models_age_rate$smooth,
      select = "s(age)",
      data = data.frame(age = seq(
        min(df[["age"]], na.rm = TRUE),
        max(df[["age"]], na.rm = TRUE),
        by = 1
      )),
      interval = "simultaneous"
    ) %>%
      select(
        age,
        derivative = .derivative,
        SE = .se,
        ci_lower = .lower_ci,
        ci_upper = .upper_ci
      )
  )
)
```

### Quadratic-smooth derivative difference

To calculate the difference between the smooth and quadratic derivatives, we can't simply sum their
uncertainties, because we need to account for their covariance.

We take a nonparametric bootstrap, meaning we sample from the actual data. Within each replicate, we
standardize the mean weight to 1, so that each replicate is equal in information, even if it happens
to have a lower total weight, because mgcv is sensitive to the scale of weights. We take samples
based on the primary sampling units within strata, rather than individuals. Within each replicate,
we fit a GAM and quadratic model to the same data, calculate the derivative for each age, and
subtract the quadratic derivative from the smooth to obtain a difference. If the sample at a given
age is randomly higher and this bump raises the GAM's derivative but drops the derivative of the
quadratic (if, for example, the bump leads to a higher intercept or linear component with lower
quadratic component to compensate, i.e., the derivative rotates), then there is negative covariance
and we get a higher total variance for that age. This means that both within-model variance and
across-model covariance is accounted for when we compute the standard error from the bootstrap.

The PSU-based sampling creates an interaction variable for PSU \* stratum, so that we identify the
PSUs within each stratum. We then identify the rows/IDs pertaining to each PSU. We then identify
which PSUs belong to which stratum. We then randomly sample PSUs within each stratum. For example,
if have 3 PSUs within one stratum, one bootstrap replicate we may have 1 2 3, another 1 1 3, another
2 2 2, another 1 3 1, etc. When a PSU is sampled, we include all rows/IDs within that PSU.
Individuals inside the same PSU are not independent, as they're from the same, for example,
geographical location. A single stratum may be, for example, high socioeconomic rural Texan black
females, and they might have 3 centers where this demographic is sampled. Each center is then a
single PSU. By sampling individuals, the model sees noise at a person level (sometimes a person is
duplicated, some are dropped), assuming people are independent, which they're not because people
within a PSU share geographical context, but with a PSU bootstrap, the model sees noise at the
cluster level. Some replicates a PSU with systematically more hearing loss gets sampled twice. The
bootstrapping then tells us the uncertainty based on 'what if this whole sampled cluster had more or
less influence' instead of 'what if these isolated people had more or less influence'. This mimics
'what if NHANES had sampled different PSUs' instead of 'what if NHANES had picked different people
from the same PSUs', which is less generalizable.

```{r diff_smooth_boot}
nonparametric_difference_bootstrap <- function(df, n_boot = 2000, threads = 1) {
  ages <- seq(
    min(df[["age"]], na.rm = TRUE),
    max(df[["age"]], na.rm = TRUE),
    by = 1
  )

  psu_id <- interaction(df$stratum, df$psu, drop = TRUE)
  psu_rows <- split(seq_len(nrow(df)), psu_id, drop = TRUE)
  stratum_psus <- split(names(psu_rows), df$stratum[match(names(psu_rows), as.character(psu_id))])

  set.seed(2)
  plan(multisession, workers = threads)

  bootstrap_derivative_diff <- simplify2array(future_lapply(
    seq_len(n_boot),
    function(i) {
      indices <- unlist(lapply(stratum_psus, function(psus) {
        unlist(psu_rows[sample(psus, replace = TRUE)], use.names = FALSE)
      }), use.names = FALSE)

      data_subset <- df[indices, ]
      weights_subset <- df$weights[indices]

      fit <- function(terms) {
        bam(
          reformulate(terms, response = params$hearing_var),
          data = data_subset,
          weights = weights_subset / mean(weights_subset),
          method = "REML",
          family = gaussian()
        )
      }

      derivative_smooth <- function(model) {
        derivatives(
          model,
          select = "s(age)",
          data = data.frame(age = ages)
        ) %>%
          as.data.frame() %>%
          pull(.derivative)
      }

      derivative_parametric <- function(model) {
        as.data.frame(
          emtrends(
            model, ~age,
            var = "age",
            at = list(age = ages),
            data = data_subset
          )
        )$age.trend
      }

      smooth_coefs <- derivative_smooth(fit(paste0("s(age, k = ", params$age_rate_k, ", m = 3)")))
      parametric_coefs <- derivative_parametric(fit(c("age", "I(age^2)")))

      smooth_coefs - parametric_coefs
    },
    future.seed = TRUE,
    future.globals = TRUE,
    future.packages = c(
      "emmeans",
      "mgcv",
      "tidytable",
      "gratia"
    )
  ))

  plan(sequential)

  SE_adj <- apply(bootstrap_derivative_diff, 1, sd)

  t_crit_max_t <- unname(quantile(
    apply(
      abs(sweep(bootstrap_derivative_diff, 1, rowMeans(bootstrap_derivative_diff), "-") / SE_adj),
      2,
      max
    ),
    0.95
  ))

  cbind(SE_adj, t_crit_max_t)
}

system.time(
  pred_age_rate$derivative_smooth <- cbind(
    pred_age_rate$derivative_smooth,
    df %>% nonparametric_difference_bootstrap(
      n_boot = params$n_boot,
      threads = min(
        params$n_threads,
        as.numeric(future::availableCores()) - 1
      )
    )
  ) %>%
    mutate(
      derivative_adj = derivative - pred_age_rate$quadratic_derivative$derivative,
      ci_lower_adj = derivative_adj - t_crit_max_t * SE_adj,
      ci_upper_adj = derivative_adj + t_crit_max_t * SE_adj
    )
)
```

### Covariates

We predict the `hearing_var` for each age in the dataset (20-79) for each stratum level with
pointwise confidence intervals.

```{r pred_covariate_pta}
predict_age_rate_covariate <- function(
  df,
  model,
  predictor = "age",
  covariate
) {
  ages <- seq(
    min(df[[predictor]], na.rm = TRUE),
    max(df[[predictor]], na.rm = TRUE),
    by = 1
  )

  newdata <- expand.grid(
    age = ages,
    covariate_level = unique(df[[covariate]])
  )

  names(newdata) <- c(predictor, covariate)

  fitted_values(
    model,
    data = newdata,
    scale = "response",
    ci_level = 0.95
  ) %>% rename(
    estimate = .fitted,
    ci_lower = .lower_ci,
    ci_upper = .upper_ci
  )
}

pred_covariates_full <- list()

pred_covariates_full <- append(
  pred_covariates_full,
  list(
    smooth_sex = predict_age_rate_covariate(
      df,
      models_covariates$smooth_sex,
      covariate = "sex"
    ),
    smooth_smoking = predict_age_rate_covariate(
      df,
      models_covariates$smooth_smoking,
      covariate = "smoking"
    ),
    smooth_race = predict_age_rate_covariate(
      df,
      models_covariates$smooth_race,
      covariate = "race"
    ),
    smooth_education = predict_age_rate_covariate(
      df,
      models_covariates$smooth_education,
      covariate = "education"
    )
  )
)
```
