---
editor_options:
  markdown:
    wrap: 100
---

# Data Processing

## Calculate age rate and covariate curves

### Predict age rate and covariate point estimates

We sample 1 point for each integer age. To calculate the quadratic derivative, we can either
calculate the slope at each point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same. We use
`derivatives()` from gratia to calculate the smooth derivative. The function is called in the
bootstrap, so that point estimates and confidence intervals are returned simultaneously.

```{r extract_age_rate_prediction}
extract_age_rate_prediction <- function(model, newdata, data, type) {
  if (type == "predict") {
    as.numeric(
      predict(
        model,
        newdata = newdata,
        type = "response"
      )
    )
  } else if (length(model$smooth) > 0) {
    derivatives(
      model,
      select = "s(age)",
      data = newdata
    ) %>%
      as.data.frame() %>%
      pull(.derivative)
  } else {
    as.data.frame(
      emtrends(
        model, ~age,
        var = "age",
        at = list(age = newdata$age),
        data = data
      )
    )$age.trend
  }
}
```

### Age rate and covariate design-based bootstrap

The PSU-based sampling creates an interaction variable for PSU \* stratum, so that we identify the
PSUs within each stratum. We then identify the rows/IDs pertaining to each PSU. We then identify
which PSUs belong to which stratum. We then randomly sample PSUs within each stratum. For example,
if have 3 PSUs within one stratum, one bootstrap replicate we may have 1 2 3, another 1 1 3, another
2 2 2, another 1 3 1, etc. When a PSU is sampled, we include all rows/IDs within that PSU.
Individuals inside the same PSU are not independent, as they're from the same, for example,
geographical location. A single stratum may be, for example, high socioeconomic rural Texan black
females, and they might have 3 centers where this demographic is sampled. Each center is then a
single PSU. By sampling individuals, the model sees noise at a person level (sometimes a person is
duplicated, some are dropped), assuming people are independent, which they're not because people
within a PSU share geographical context, but with a PSU bootstrap, the model sees noise at the
cluster level. Some replicates a PSU with systematically more hearing loss gets sampled twice. The
bootstrapping then tells us the uncertainty based on 'what if this whole sampled cluster had more or
less influence' instead of 'what if these isolated people had more or less influence'. This mimics
'what if NHANES had sampled different PSUs' instead of 'what if NHANES had picked different people
from the same PSUs', which is less generalizable.

We correct the variance at the end, taking into account the number of PSUs per stratum. We then
correct the variance with $\frac{n}{(n-1)}$. This is needed because we sample with replacement, if
we have PSU A and B in a strata, we have 50% chance to draw AA or BB and 50% to draw AB or BA (the
original mean). The mean contribution is then about half the true sampling variance. If each stratum
had 2 PSUs, we could've multiplied by $\sqrt2$, but since some strata have 3 PSUs, we need to
calculate the harmonic mean.

To calculate the difference between the smooth and quadratic derivatives, we can't simply sum their
uncertainties, because we need to account for their covariance.

We take a nonparametric bootstrap, meaning we sample from the actual data. Within each replicate, we
standardize the mean weight to 1, so that each replicate is equal in information, even if it happens
to have a lower total weight, because mgcv is sensitive to the scale of weights. We take samples
based on the primary sampling units within strata, rather than individuals. Within each replicate,
we fit a GAM and quadratic model to the same data, calculate the derivative for each age, and
subtract the quadratic derivative from the smooth to obtain a difference. If the sample at a given
age is randomly higher and this bump raises the GAM's derivative but drops the derivative of the
quadratic (if, for example, the bump leads to a higher intercept or linear component with lower
quadratic component to compensate, i.e., the derivative rotates), then there is negative covariance
and we get a higher total variance for that age. This means that both within-model variance and
across-model covariance is accounted for when we compute the standard error from the bootstrap. The
same procedure applies to the other models.

The confidence intervals are simultaneous, not pairwise confidence intervals. We calculate the mean
prediction across all 10000 replicates at each age and subtract the age-specific mean from each
individual replicate. With this, we calculate the deviation from the mean for each age for each
replicate and express the deviation standardized by the deviation of that age. We then find the
worst standardized deviation among any age for each replicate. We then find the threshold that 95%
of the maximum deviations do not exceed. The critical value based on the largest deviation is then
used for each ages. If the variance of an age is lower, then the width of the confidence intervals
will still be lower, since we multiply by the standard error of each age. To illustrate, if age 50
has a mean standard deviation of 4 dB and a bootstrap replicate has an SD of 2 or 6 dB at that age
(2 different from the mean), the standardized deviation is 0.5. If at age 25 the mean SD is 0.2 dB
and a bootstrap replicate has an SD of 0.8 (0.6 difference), the standardized difference is 3. We
check here what the largest departure is from resampling relative to local uncertainty, then apply
that uncertainty factor to all ages. The 95% confidence intervals then represent the values for
which, if we repeat the experiment many times, 95% of the cases the entire curve would fall within
that band, so that essentially we're 95% sure the entire curve lays within the confidence intervals.
If any point falls outside that confidence interval range, we have a significant departure from the
null. More ages to compare gives more opportunities for the maximum to be increased, but adjacent
values often are strongly correlated, so going from testing at every 5 ages to every age does not
widen confidence intervals much.

```{r predict_age_rate_bootstrap}
predict_age_rate_bootstrap <- function(
  df,
  model,
  type = "predict",
  covariate = NULL,
  reference_model = NULL,
  n_boot = params$n_boot,
  threads = max(1, min(params$n_threads, availableCores() - 1))
) {
  # Get grid of all ages
  ages <- seq(
    min(df[["age"]], na.rm = TRUE),
    max(df[["age"]], na.rm = TRUE),
    by = 1
  )

  # Add covariates to the prediction grid if needed
  if (is.null(covariate)) {
    newdata <- data.frame(age = ages)
  } else {
    newdata <- expand.grid(
      age = ages,
      covariate_level = unique(df[[covariate]])
    )

    names(newdata) <- c("age", covariate)
  }

  model_formula <- formula(model)

  # Store reference formula for the derivative difference
  reference_formula <- if (!is.null(reference_model)) {
    formula(reference_model)
  } else {
    NULL
  }

  # Point estimates from the full-sample models
  predictions <- extract_age_rate_prediction(
    model,
    newdata,
    df,
    type
  )

  if (!is.null(reference_model)) {
    predictions <- predictions - extract_age_rate_prediction(
      reference_model,
      newdata,
      df,
      type
    )
  }

  # Store PSU rows for bootstrap
  psu_id <- interaction(df$stratum, df$psu, drop = TRUE)
  psu_rows <- split(seq_len(nrow(df)), psu_id, drop = TRUE)
  stratum_psus <- split(names(psu_rows), df$stratum[match(names(psu_rows), as.character(psu_id))])

  set.seed(2)
  plan(multisession, workers = threads)

  bootstrap_predictions <- simplify2array(future_lapply(
    seq_len(n_boot),
    function(i) {
      indices <- unlist(lapply(stratum_psus, function(psus) {
        unlist(psu_rows[sample(psus, replace = TRUE)], use.names = FALSE)
      }), use.names = FALSE)

      data_subset <- df[indices, ]

      # Fit with the sampled weights
      fit <- function(model_formula) {
        bam(
          model_formula,
          data = data_subset,
          weights = weights / mean(weights),
          method = "REML",
          family = gaussian()
        )
      }

      pred <- extract_age_rate_prediction(
        fit(model_formula),
        newdata,
        data_subset,
        type
      )

      # Fit both models to the same sample for a paired difference
      if (!is.null(reference_formula)) {
        pred <- pred - extract_age_rate_prediction(
          fit(reference_formula),
          newdata,
          data_subset,
          type
        )
      }

      pred
    },
    future.seed = TRUE,
    future.globals = TRUE,
    future.packages = c(
      "emmeans",
      "mgcv",
      "tidytable",
      "gratia"
    )
  ))

  plan(sequential)

  SE_adj <- apply(bootstrap_predictions, 1, sd)

  # Simultaneous critical value over this prediction grid
  t_crit_max_t <- unname(quantile(
    apply(
      abs(sweep(bootstrap_predictions, 1, rowMeans(bootstrap_predictions), "-") / SE_adj),
      2,
      max
    ),
    0.95
  ))

  # Correct for the small number of PSUs per stratum
  SE_adj <- SE_adj / sqrt(1 - mean(1 / lengths(stratum_psus)))

  predictions <- newdata %>%
    as_tibble() %>%
    mutate(
      estimate = predictions,
      SE = SE_adj,
      ci_lower = estimate - t_crit_max_t * SE,
      ci_upper = estimate + t_crit_max_t * SE
    )

  if (!is.null(reference_model)) {
    predictions %>%
      transmute(
        derivative_adj = estimate,
        SE_adj = SE,
        t_crit_max_t = t_crit_max_t,
        ci_lower_adj = ci_lower,
        ci_upper_adj = ci_upper
      )
  } else if (type == "predict_derivative") {
    predictions %>%
      rename(derivative = estimate)
  } else {
    predictions
  }
}
```

### Run pipeline

We build named lists for the age rate and covariate curves, to which we append the point estimates
and confidence intervals.

```{r build_age_rate_predictions}
pred_age_rate <- list()

system.time(
  pred_age_rate <- append(
    pred_age_rate,
    list(
      linear = predict_age_rate_bootstrap(
        df,
        models_age_rate$linear,
        type = "predict"
      ),
      linear_derivative = predict_age_rate_bootstrap(
        df,
        models_age_rate$linear,
        type = "predict_derivative"
      ),
      quadratic = predict_age_rate_bootstrap(
        df,
        models_age_rate$quadratic,
        type = "predict"
      ),
      quadratic_derivative = predict_age_rate_bootstrap(
        df,
        models_age_rate$quadratic,
        type = "predict_derivative"
      ),
      smooth = predict_age_rate_bootstrap(
        df,
        models_age_rate$smooth,
        type = "predict"
      ),
      derivative_smooth = predict_age_rate_bootstrap(
        df,
        models_age_rate$smooth,
        type = "predict_derivative"
      )
    )
  )
)

system.time(
  pred_age_rate$derivative_smooth <- cbind(
    pred_age_rate$derivative_smooth,
    predict_age_rate_bootstrap(
      df,
      models_age_rate$smooth,
      type = "predict_derivative",
      reference_model = models_age_rate$quadratic
    )
  )
)

pred_covariates_full <- list()

system.time(
  pred_covariates_full <- append(
    pred_covariates_full,
    list(
      smooth_sex = predict_age_rate_bootstrap(
        df,
        models_covariates$smooth_sex,
        covariate = "sex"
      ),
      smooth_smoking = predict_age_rate_bootstrap(
        df,
        models_covariates$smooth_smoking,
        covariate = "smoking"
      ),
      smooth_race = predict_age_rate_bootstrap(
        df,
        models_covariates$smooth_race,
        covariate = "race"
      ),
      smooth_education = predict_age_rate_bootstrap(
        df,
        models_covariates$smooth_education,
        covariate = "education"
      )
    )
  )
)
```

The confidence intervals of the smooth curve wiggle considerably. Diagnostic tests (not shown)
revealed that these oscillations are the result of variation in the smoothing parameter calculated
over different bootstrap samples. The overall curve remains the same but the local slope (i.e., a
point on the derivative) can differ quite a bit. Small perturbations to the dataset can produce
considerable deviations in the local slope. Uncertainty in the smoothing parameter is a real
component of generalized additive models and should not be ignored. The gratia package, for example,
calculates confidence intervals with a fixed smoothing parameter, but that is not per se justified.
In our case, we have 4.7 effective degrees of freedom in the full dataset fit, whereas the median of
the bootstrap is 5.47. This uncertainty in the amount of small, local wiggliness is real and
important for hypothesis testing. If we would consider the smoothing parameter fixed/certain, we
could wrongly interpret a hump in the derivative as significant when that hump is actually
'incidental', many small local slopes after all can produce very similar complete curves. A straight
line up vs a wavy line up that only wiggles minutely would have completely different local slopes
even if the global slope is the same, and thus if we ask whether there are specific ages during
which the rate of hearing loss speeds up/slows down must consider that uncertainty.
