---
editor_options:
  markdown:
    wrap: 100
---

# Data Processing

## Split up bioage predictions into two

The reverse dataframe compares models with age and biological age to only biological age, to assess
the effect of age beyond biological age.

```{r split_bioage_preds}
split_bioage <- function(predictions) {
  predictions %>%
    filter(!predictors %in% c("phenoage", "kdm", "hd_log")) %>%
    mutate(predictors = fct_drop(predictors))
}

split_bioage_reverse <- function(predictions) {
  predictions %>%
    filter(predictors != "age") %>%
    mutate(predictors = fct_drop(predictors))
}

pred_bioage <- split_bioage(pred_bioage_bidirectional)

if (isFALSE(params$telolength)) {
  pred_bioage_reverse <- split_bioage_reverse(pred_bioage_bidirectional)
}
```

## Build predictions dataframe age rate

We sample 1 point for each age and predict the FI using the smooth-only model and generate
simultaneous confidence intervals with 'mvt'. We could ask for a single point and ask where the true
mean would be with 95% certainty, but if we'd apply that to all 60 points, there's a large chance
that somewhere we get a false positive. With the simultaneous interval or with multivariate t
distribution error correction, we can say that we're 95% sure that the **entire** derivative in the
true population that was sampled from (all people on earth if NHANES is representative) is within
the band.

To calculate the quadratic derivative, we can either calculate the slope at each point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same. However, using
emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

We add the uncertainty of the quadratic model derivative to that of the smooth (it's tiny compared
to that of the smooth). This is slightly conservative, because the variances are positively
correlated. However, the smooth term absorbs some of the variance normally explained by the linear
and quadratic terms, so the terms aren't fully uncorrelated. Because of this, the coefficients of
the linear and quadratic terms are different between the nested smooth and quadratic models and,
therefore, uncertainties are also different, which we need to account for. We back-calculate the
multivariate-corrected t-crit value for the smooth and recalculate the confidence intervals but with
the combined uncertainty from the quadratic and smooth.

### Main models

```{r predict_aging}
predict_age_rate <- function(
  df,
  model,
  predictor = "age",
  type = c("predict", "predict_derivative")
) {
  ages <- setNames(
    list(
      seq(
        min(df[[predictor]]),
        max(df[[predictor]]),
        by = 1
      )
    ),
    predictor
  )
  formula <- as.formula(
    paste("~", predictor)
  )

  if (type == "predict") {
    em <- emmeans(
      model,
      formula,
      at = ages,
      type = "response",
      data = df
    )
  } else {
    em <- emtrends(
      model,
      formula,
      var = predictor,
      at = ages,
      data = df
    )
  }

  predictions <- tibble(
    confint(em, adjust = "mvt")
  )

  if (type == "predict") {
    predictions %>%
      rename(
        "estimate" = emmean,
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  } else {
    predictions %>%
      rename(
        "derivative" = !!paste0(predictor, ".trend"),
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  }
}

pred_age_rate <- list()

pred_age_rate <- append(
  pred_age_rate,
  list(
    linear = predict_age_rate(
      df,
      models_age_rate$linear,
      type = "predict"
    ),
    linear_derivative = predict_age_rate(
      df,
      models_age_rate$linear,
      type = "predict_derivative"
    ),
    quadratic = predict_age_rate(
      df,
      models_age_rate$quadratic,
      type = "predict"
    ),
    quadratic_derivative = predict_age_rate(
      df,
      models_age_rate$quadratic,
      type = "predict_derivative"
    ),
    smooth = predict_age_rate(
      df,
      models_age_rate$smooth,
      type = "predict"
    ),
    derivative_smooth = predict_age_rate(
      df,
      models_age_rate$smooth,
      type = "predict_derivative"
    )
  )
)

pred_age_rate$derivative_smooth <- pred_age_rate$derivative_smooth %>%
  mutate(
    SE_adj = sqrt(SE^2 + pred_age_rate$quadratic_derivative$SE^2),
    derivative_adj = derivative - pred_age_rate$quadratic_derivative$derivative,
    t_crit_mvt = (ci_upper - ci_lower) / (2 * SE),
    ci_lower_adj = derivative_adj - t_crit_mvt * SE_adj,
    ci_upper_adj = derivative_adj + t_crit_mvt * SE_adj
  )
```

### Covariates

We predict the `hearing_var` for each age in the dataset (20-84) for each stratum level.

```{r}
predict_age_rate_covariate <- function(
  df,
  model,
  predictor = "age",
  covariate
) {
  ages <- setNames(
    list(
      seq(
        min(df[[predictor]]),
        max(df[[predictor]]),
        by = 1
      )
    ),
    predictor
  )

  em <- emmeans(
    model,
    as.formula(
      paste0("~ ", predictor, " + ", covariate)
    ),
    at = ages,
    by = covariate,
    type = "response",
    data = df
  )

  predictions <- as.data.frame(
    confint(em, adjust = "mvt")
  )

    predictions %>%
      rename(
        "estimate" = emmean,
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
}

pred_covariates_full <- list()

pred_covariates_full <- append(
  pred_covariates_full,
  list(
    smooth_gender = predict_age_rate_covariate(
      df,
      models_covariates$smooth_gender,
      covariate = "gender"
    ),
    smooth_smoking = predict_age_rate_covariate(
      df,
      models_covariates$smooth_smoking,
      covariate = "smoking"
    ),
    smooth_race = predict_age_rate_covariate(
      df,
      models_covariates$smooth_race,
      covariate = "race"
    ),
    smooth_education = predict_age_rate_covariate(
      df,
      models_covariates$smooth_education,
      covariate = "education"
    )
  )
)
```

## Calculate coefficients

We estimate the rate of change in dB/year and acceleration in dB/year² for the quadratic model and
GAM. For the GAM, we estimate the acceleration by fitting a linear model to the derivative and
calculating its coefficient.

```{r calc_coeffs}
with(pred_age_rate$linear, cov(age, estimate) / var(age))
with(pred_age_rate$quadratic_derivative, cov(age, derivative) / var(age))
coef(lm(derivative ~ age, data = pred_age_rate$derivative_smooth))[["age"]]
```
