---
editor_options:
  markdown:
    wrap: 100
---

# Statistics

## Calculate descriptive statistics

We turn our repeated design object back into a regular Survey object for calculating population
means. Because no PSUs are filtered out, we can reconstruct the object from the filtered dataset. We
use the full survey design to generate a descriptive statistics table on the filtered and unfiltered
audiometry samples across cycles. We calculate both weighted and unweighted proportions and check n
per stratum. We only include the strata in the unfiltered dataset that's in the filtered dataset and
remove NA values. Bioage descriptive statistics are also optionally calculated. Variance estimates
are calculated with Taylor linearization.

```{r calc_descriptives}
calculate_descriptive_statistics <- function(design, audiometry_cols, incl_bioage = FALSE) {
  design <- svydesign(
    ids = ~psu,
    strata = ~stratum,
    weights = ~weights,
    data = design$variables,
    nest = TRUE
  )

  design <- update(
    design,
    sex = dplyr::recode(
      sex,
      "1" = "Male",
      "2" = "Female",
      .default = NA_character_
    ),
    smoking = dplyr::recode(
      smoking,
      "1" = "Smoker",
      "2" = "Non-smoker",
      .default = NA_character_
    ),
    race = dplyr::recode(
      race,
      "1" = "Hispanic",
      "3" = "White",
      "4" = "Black",
      "5" = "Other",
      .default = NA_character_
    ),
    education = dplyr::recode(education,
      "1" = "< 9th grade",
      "2" = "9-11th grade",
      "3" = "High school",
      "4" = "Some college or AA degree",
      "5" = "College graduate or higher",
      .default = NA_character_
    )
  )

  groups <- c(Sex = "sex", Race = "race", Smoking = "smoking", Education = "education")
  age_groups <- factor(c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
  results <- list()

  design <- design %>%
    subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(audiometry_cols), ~ . == 888 | . == 666 | is.na(.))) %>%
        pull(keep) &
        age_group %in% age_groups &
        !is.na(sex) &
        !is.na(smoking) &
        !is.na(race) &
        !is.na(education)
    )


  for (age_group_val in age_groups) {
    design_subset <- if (age_group_val == "Overall") design else subset(design, age_group == age_group_val)

    if (age_group_val == "Overall") {
      proportion_w <- 1
    } else {
      proportion_w <- as.vector(
        coef(svymean(~age_group, design))[paste0("age_group", age_group_val)]
      )
    }

    rows <- bind_rows(
      data.frame(
        age_group = age_group_val,
        group = "Overall",
        category = "Overall",
        n = nrow(design_subset),
        proportion = nrow(design_subset) / nrow(design),
        proportion_w = proportion_w,
        mean_w_pta_hf = coef(svymean(~pta_hf, design_subset)),
        sd_w_pta_hf = sqrt(coef(svyvar(~pta_hf, design_subset))),
        mean_w_pta_lf = coef(svymean(~pta_lf, design_subset)),
        sd_w_pta_lf = sqrt(coef(svyvar(~pta_lf, design_subset))),
        mean_w_age = coef(svymean(~age, design_subset)),
        sd_w_age = sqrt(coef(svyvar(~age, design_subset)))
      ) %>%
        {
          if (isTRUE(incl_bioage)) {
            mutate(.,
              mean_w_kdm = coef(svymean(~kdm, design_subset)),
              mean_w_phenoage = coef(svymean(~phenoage, design_subset)),
              mean_w_hd_log = coef(svymean(~hd_log, design_subset)),
              mean_w_telolength_log = coef(svymean(~telolength_log, design_subset, na.rm = TRUE)),
              sd_w_kdm = sqrt(coef(svyvar(~kdm, design_subset))),
              sd_w_phenoage = sqrt(coef(svyvar(~phenoage, design_subset))),
              sd_w_hd_log = sqrt(coef(svyvar(~hd_log, design_subset))),
              sd_w_telolength_log = sqrt(coef(svyvar(~telolength_log, design_subset, na.rm = TRUE))),
            )
          } else {
            .
          }
        },
      lapply(names(groups), function(group) {
        grouping_variable <- as.formula(paste0("~", groups[[group]]))
        proportions_w <- coef(svymean(grouping_variable, design_subset))
        counts <- table(design_subset$variables[[groups[[group]]]])

        data.frame(
          age_group = age_group_val,
          group = group,
          category = svyby(~pta_hf, grouping_variable, design_subset, svymean)[[groups[[group]]]],
          n = as.vector(counts),
          proportion = as.vector(counts) / nrow(design_subset),
          proportion_w = as.vector(proportions_w),
          mean_w_pta_hf = svyby(~pta_hf, grouping_variable, design_subset, svymean)[["pta_hf"]],
          sd_w_pta_hf = sqrt(svyby(~pta_hf, grouping_variable, design_subset, svyvar)[["pta_hf"]]),
          mean_w_pta_lf = svyby(~pta_lf, grouping_variable, design_subset, svymean)[["pta_lf"]],
          sd_w_pta_lf = sqrt(svyby(~pta_lf, grouping_variable, design_subset, svyvar)[["pta_lf"]]),
          mean_w_age = svyby(~age, grouping_variable, design_subset, svymean)[["age"]],
          sd_w_age = sqrt(svyby(~age, grouping_variable, design_subset, svyvar)[["age"]])
        ) %>%
          {
            if (isTRUE(incl_bioage)) {
              mutate(.,
                mean_w_kdm = svyby(~kdm, grouping_variable, design_subset, svymean)$kdm,
                mean_w_phenoage = svyby(~phenoage, grouping_variable, design_subset, svymean)$phenoage,
                mean_w_hd_log = svyby(~hd_log, grouping_variable, design_subset, svymean)$hd_log,
                mean_w_telolength_log = svyby(~telolength_log, grouping_variable, design_subset, svymean, na.rm = TRUE)$telolength_log,
                sd_w_kdm = sqrt(svyby(~kdm, grouping_variable, design_subset, svyvar)$kdm),
                sd_w_phenoage = sqrt(svyby(~phenoage, grouping_variable, design_subset, svyvar)$phenoage),
                sd_w_hd_log = sqrt(svyby(~hd_log, grouping_variable, design_subset, svyvar)$hd_log),
                sd_w_telolength_log = sqrt(svyby(~telolength_log, grouping_variable, design_subset, svyvar, na.rm = TRUE)$telolength_log)
              )
            } else {
              .
            }
          }
      })
    )

    results[[age_group_val]] <- rows
  }

  bind_rows(results) %>%
    mutate(
      group = factor(group, levels = c("Overall", "Sex", "Race", "Smoking", "Education")),
      age_group = factor(age_group, levels = c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
    )
}

descriptives_unfiltered <- calculate_descriptive_statistics(design_no_missing, audiometry_cols = audiometry_cols)
descriptives <- calculate_descriptive_statistics(design, audiometry_cols = audiometry_cols, incl_bioage = TRUE)
```

## Calculate (delta) MSE

We set the age group and rename mse and rmse to `mse_w` and `rmse_w`. We then take all the results
with the reference predictor and add the reference group by the replicate and calculate delta MSE,
delta RMSE and partial R2².

```{r}
calculate_mse_per_replicate <- function(estimates, reference, covariate_analysis = FALSE) {
  results <- estimates %>%
    mutate(
      age_group = if_else(age_group == "Overall", "overall", age_group)
    ) %>%
    rename(
      mse_w = mse,
      rmse_w = rmse
    )

  results <- map_dfr(reference, \(ref) {
    results %>%
      filter(if (isTRUE(covariate_analysis)) TRUE else grepl(ref, predictors, fixed = TRUE)) %>%
      left_join(
        results %>%
          filter(predictors == ref) %>%
          mutate(
            mse_w_ref = mse_w,
            rmse_w_ref = rmse_w,
            reference = ref
          ) %>%
          select(reference, age_group, replicate, mse_w_ref, rmse_w_ref),
        by = c("age_group", "replicate")
      )
  })

  if (isFALSE(covariate_analysis)) {
    results %>%
      mutate(
        delta_mse_w = mse_w_ref - mse_w,
        delta_rmse_w = rmse_w_ref - rmse_w,
        r2p = delta_mse_w / mse_w_ref
      )
  } else {
    results %>%
      mutate(
        delta_mse_w = mse_w - mse_w_ref,
        delta_rmse_w = rmse_w - rmse_w_ref,
        r2p = delta_mse_w / mse_w
      )
  }
}

results_bioage <- pred_bioage$estimates %>%
  calculate_mse_per_replicate(
    reference = "age"
  )

if (isFALSE(params$telolength)) {
  results_bioage_reverse <- pred_bioage_reverse$estimates %>%
    calculate_mse_per_replicate(
      reference = c("phenoage", "kdm", "hd_log")
    )

  results_biomarkers <- pred_biomarkers$estimates %>%
    calculate_mse_per_replicate(
      reference = "age"
    )

  results_covariates <- pred_covariates$estimates %>%
    calculate_mse_per_replicate(
      reference = "age, sex, smoking, race, education",
      covariate_analysis = TRUE
    )
}
```

## Calculate summary statistics

The `calculate_variance()` function computes the jackknife variance. `x` is the variable of
interest, `replicate` is the replicate number. We group by age group, predictors, and the reference.
Therefore, `calculate_variance()` receives a matrix with only the variable of interest from one age
group and predictor with all replicates. `thetas` is the argument that takes this matrix (the
dataset). We ignore replicate 0 since this is the full dataset. `scale` turns how much the MSE moved
from the original dataset across jackknife results into a variation, `rscales` adjusts each
individual replicate since some jackknife repeats deleted one PSU from a stratum with 2 and in
another from a stratum with 3 PSUs, meaning the perturbations have different sizes and must be
scaled accordingly. `mse` is a boolean value that we defined in the asrepdesign object and says we
compare MSE values from the jackknife replicates with the original rather than with the average
across jackknife replicates. These are all computed by the Survey package. `coef` uses the variable
of interest from the reference dataset to compare to. `na.action` tells Survey to fail if any
replicate is missing.

**Total sum of squares calculation when binning R² per age group**: It's imperative that for the age
group stratified descriptives, the global mean is used. Otherwise it's as if the total sum of
squares is calculated within each age-bin, which is as if a linear regression is fitted through the
mean PTA per age group, explaining a lot of variation while it shouldn't. We want to assess the
predictive effect on age, so the total sum of squares should be calculated by taking the distance
from each PTA within an age group to the global mean. Of course, the models are trained on the whole
dataset, not just one age group. We're just assessing performance per age group.

Point estimates are calculated using the reference dataset. We convert variances to standard errors.

```{r calc_pooled_mse}
calculate_pooled_statistics <- function(results, design, data) {
  calculate_variance <- function(x, replicate) {
    as.numeric(
      svrVar(
        thetas = x[replicate > 0],
        scale = design$scale,
        rscales = design$rscales,
        mse = design$mse,
        coef = x[replicate == 0],
        na.action = "na.fail"
      )
    )
  }

  critical_value <- qt(
    0.975,
    df = degf(design)
  )

  # Make a matrix of weights to align with replicates
  cv_weights <- cbind(
    weights(design, type = "sampling"),
    weights(design, type = "analysis")
  )

  # Iterate over each replicate to calculate sum of squares per age group and overall
  sst <- map_dfr(seq_len(ncol(cv_weights)), \(j) {
    data <- data %>%
      mutate(weights = cv_weights[, j])

    weighted_mean <- sum(
      data$weights * data[[params$hearing_var]]
    ) / sum(data$weights)

    bind_rows(
      data %>%
        summarize(
          sst_w = sum(weights * (.data[[params$hearing_var]] - weighted_mean)^2) / sum(weights),
          .by = age_group
        ) %>%
        mutate(age_group = as.character(age_group)),
      data %>%
        summarize(
          sst_w = sum(weights * (.data[[params$hearing_var]] - weighted_mean)^2) / sum(weights)
        ) %>%
        mutate(age_group = "overall")
    ) %>%
      mutate(replicate = j - 1L)
  })

  results %>%
    left_join(
      sst,
      by = c("age_group", "replicate")
    ) %>%
    mutate(
      r2_w = 1 - (mse_w / sst_w)
    ) %>%
    arrange(replicate) %>%
    summarize(
      s2_mse_w = calculate_variance(mse_w, replicate),
      delta_s2_mse_w = calculate_variance(delta_mse_w, replicate),
      s2_rmse_w = calculate_variance(rmse_w, replicate),
      delta_s2_rmse_w = calculate_variance(delta_rmse_w, replicate),
      s2_r2_w = calculate_variance(r2_w, replicate),
      s2_r2p = calculate_variance(r2p, replicate),
      mse_w_ref = mse_w_ref[replicate == 0],
      rmse_w_ref = rmse_w_ref[replicate == 0],
      mse_w = mse_w[replicate == 0],
      delta_mse_w = delta_mse_w[replicate == 0],
      rmse_w = rmse_w[replicate == 0],
      delta_rmse_w = delta_rmse_w[replicate == 0],
      sst_w = sst_w[replicate == 0],
      r2_w = r2_w[replicate == 0],
      r2p = r2p[replicate == 0],
      .by = c(age_group, predictors, reference)
    ) %>%
    mutate(
      se_mse_w = sqrt(s2_mse_w),
      se_delta_mse_w = sqrt(delta_s2_mse_w),
      se_rmse_w = sqrt(s2_rmse_w),
      se_delta_rmse_w = sqrt(delta_s2_rmse_w),
      se_r2_w = sqrt(s2_r2_w),
      se_r2p = sqrt(s2_r2p),
      ci_low_mse = mse_w - critical_value * se_mse_w,
      ci_high_mse = mse_w + critical_value * se_mse_w,
      ci_low_delta_mse = delta_mse_w - critical_value * se_delta_mse_w,
      ci_high_delta_mse = delta_mse_w + critical_value * se_delta_mse_w,
      ci_low_rmse = rmse_w - critical_value * se_rmse_w,
      ci_high_rmse = rmse_w + critical_value * se_rmse_w,
      ci_low_delta_rmse = delta_rmse_w - critical_value * se_delta_rmse_w,
      ci_high_delta_rmse = delta_rmse_w + critical_value * se_delta_rmse_w,
      ci_low_r2 = r2_w - critical_value * se_r2_w,
      ci_high_r2 = r2_w + critical_value * se_r2_w,
      ci_low_r2p = r2p - critical_value * se_r2p,
      ci_high_r2p = r2p + critical_value * se_r2p
    )
}

summary_bioage <- results_bioage %>%
  calculate_pooled_statistics(
    design = design,
    data = df
  )

summary_bioage_overall <- summary_bioage %>%
  filter(age_group == "overall")
summary_bioage <- summary_bioage %>%
  filter(age_group != "overall")

if (isFALSE(params$telolength)) {
  summary_bioage_reverse <- results_bioage_reverse %>%
    calculate_pooled_statistics(
      design = design,
      data = df
    )

  summary_bioage_reverse_overall <- summary_bioage_reverse %>%
    filter(age_group == "overall")
  summary_bioage_reverse <- summary_bioage_reverse %>%
    filter(age_group != "overall")

  summary_biomarkers <- results_biomarkers %>%
    calculate_pooled_statistics(
      design = design,
      data = df
    )

  summary_biomarkers_overall <- summary_biomarkers %>%
    filter(age_group == "overall")
  summary_biomarkers <- summary_biomarkers %>%
    filter(age_group != "overall")

  summary_covariates <- results_covariates %>%
    calculate_pooled_statistics(
      design = design,
      data = df
    )

  summary_covariates_overall <- summary_covariates %>%
    filter(age_group == "overall")
  summary_covariates <- summary_covariates %>%
    filter(age_group != "overall")
}
```

## Repeated Nadeau-Bengio corrected t-tests

### Calculate p-values

Calculate t-statistic according to Nadeau Bengio (although we use a weighted adaptation of it),
which is `t = m/(√(variance(1/n + n2/n1)))`. `m` is the difference between the sample mean and
hypothesized population mean. In this case, we've normalized it to age, so the null hypothesis (the
hypothesized population mean) is 0, i.e., there is no difference from age only. The variance was
already calculated before. `n` is equivalent to the number of folds times the number of repeats.
`n2` is the number of testing/validation data and `n1` the number of training data. We then
calculate degrees of freedom by taking n - 1. Then we use `pt()` (which automatically calculates the
critical value) using a one-tailed t-test of the upper tail. This is because we're not interested in
whether an age + bioage model is different from age-only. We only want to know if it is better. We
then adjust p values with Holm error correction.

First, the question is, 'Does this bioage/biomarker add predictive value for hearing loss in the
population while controlling for chronological age?', which we test with the overall comparisons.
Only if it's significantly different do we look at per age-group tests. Another option is to ask
'Might bioage/biomarker only matter for specific age ranges, even if it doesn't have an overall
predictive effect?', in which case we could consider doing pairwise on all, but we'd lose a lot of
power and would have to use FDR and accept a 5% false discovery rate to salvage it.

```{r t_test_models}
jackknife_t_test <- function(summary,
                             design,
                             exceptions = NULL,
                             exception_include_exclude,
                             adjust_method) {
  if ("p" %in% colnames(summary)) {
    summary <- summary %>%
      select(
        -t, -df, -p, -p_adj, -signif
      )
  }
  if (exception_include_exclude == "include") {
    pwc <- summary %>%
      filter(predictors %in% exceptions)
  } else {
    pwc <- summary %>%
      filter(!predictors %in% exceptions)
  }

  pwc <- pwc %>%
    mutate(
      t = delta_mse_w / se_delta_mse_w,
      df = degf(design),
      p = pt(t, df = df, lower.tail = FALSE),
      p_adj = p.adjust(p, method = adjust_method),
      signif = case_when(
        p_adj <= 0.001 ~ "***",
        p_adj <= 0.01 ~ "**",
        p_adj <= 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(
      age_group, predictors, t, df, p, p_adj, signif
    )

  summary <- summary %>%
    left_join(pwc,
      by = c("age_group", "predictors")
    )

  summary
}

summary_bioage_overall <- jackknife_t_test(
  summary_bioage_overall,
  design = design,
  exceptions = c(
    "age",
    if (isTRUE(params$telolength)) "age, phenoage" else NULL,
    if (isTRUE(params$telolength)) "age, kdm" else NULL,
    if (isTRUE(params$telolength)) "age, hd" else NULL
  ),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)

summary_bioage <- jackknife_t_test(
  summary_bioage,
  design = design,
  exceptions = c("age, phenoage", "age, kdm"),
  exception_include_exclude = "include",
  adjust_method = "holm"
)

if (isFALSE(params$telolength)) {
  summary_bioage_reverse_overall <- jackknife_t_test(
    summary_bioage_reverse_overall,
    design = design,
    exceptions = c("phenoage", "kdm", "hd_log"),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_bioage_reverse <- jackknife_t_test(
    summary_bioage_reverse,
    design = design,
    exceptions = c("phenoage", "kdm", "hd_log"),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_biomarkers_overall <- jackknife_t_test(
    summary_biomarkers_overall,
    design = design,
    exceptions = c(
      "age"
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_biomarkers <- jackknife_t_test(
    summary_biomarkers,
    design = design,
    exceptions = c(
      "crp"
    ),
    exception_include_exclude = "include",
    adjust_method = "holm"
  )

  summary_covariates_overall <- jackknife_t_test(
    summary_covariates_overall,
    design = design,
    exceptions = c(
      "age, sex, smoking, race, education"
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_covariates <- jackknife_t_test(
    summary_covariates,
    design = design,
    exceptions = c(
      "age, sex, smoking, race, education",
      "age, sex, race, education",
      if (params$hearing_var == "pta_lf") "age, sex, smoking, education" else NULL,
      if (params$hearing_var == "pta_lf") "age, smoking, race, education" else NULL
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )
}
```

### Display summary tables

#### Make rounding function

We round to a max of 3 decimals (0.001) but scale it to the size, so 9.12312 becomes 9.12, 91.1231
becomes 91.1, 912.123123 becomes 912, and anything larger than or equal to 1000 gets rounded to an
integer (so that 12345 doesn't become 12300).

```{r round_function}
clean_rounding <- function(x, max_decimals = 3) {
  ifelse(
    abs(x) >= 10^max_decimals,
    round(x),
    round(signif(x, max_decimals), max_decimals)
  )
}
```

#### Bioage overall

```{r pwc_bioage_overall}
summary_bioage_overall %>%
  mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
  construct_table(
    title = "Overall Bioage Pairwise Comparisons"
  )
```

#### Bioage

```{r pwc_bioage}
summary_bioage %>%
  mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
  construct_table(
    title = "Per-Age Group Bioage Pairwise Comparisons"
  )
```

#### Bioage reverse overall

```{r pwc_bioage_reverse_overall}
if (isFALSE(params$telolength)) {
  summary_bioage_reverse_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Bioage Reverse Analysis Pairwise Comparisons"
    )
}
```

#### Bioage reverse

```{r pwc_bioage_reverse}
if (isFALSE(params$telolength)) {
  summary_bioage_reverse %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Bioage Reverse Analysis Pairwise Comparisons"
    )
}
```

#### Biomarkers overall

```{r pwc_biomarkers_overall}
if (isFALSE(params$telolength)) {
  summary_biomarkers_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Biomarker Pairwise Comparisons"
    )
}
```

#### Biomarkers

```{r pwc_biomarkers}
if (isFALSE(params$telolength)) {
  summary_biomarkers %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Biomarker Pairwise Comparisons"
    )
}
```

#### Covariates overall

```{r pwc_covariates_overall}
if (isFALSE(params$telolength)) {
  summary_covariates_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Covariate Pairwise Comparisons"
    )
}
```

#### Covariates

```{r pwc_covariates}
if (isFALSE(params$telolength)) {
  summary_covariates %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Covariate Pairwise Comparisons"
    )
}
```

### Add n training and testing datasets to dataframe

We calculate the number of participants in the dataset for each repeat excluding the testing fold
for the training dataset and for only the testing fold for the test datatset. We then show the
unique participant count with `n` and the `n_train_mean` and `n_test_mean` to show the mean across
folds.

```{r add_n}
add_n_age_groups <- function(summary_data, data, per_age_group = FALSE) {
  if (isFALSE(per_age_group)) {
    data <- data %>%
      mutate(age_group = "overall")
  }

  sample_sizes <- map_dfr(seq_len(params$global_repeats), function(r) {
    repeat_column <- paste0("fold_r", r)

    map_dfr(seq_len(params$global_folds), function(fold) {
      # Models train across all age groups
      train_n <- sum(data[[repeat_column]] != fold)

      data %>%
        summarize(
          n = n(),
          n_train = train_n,
          n_test = sum(.data[[repeat_column]] == fold),
          .by = age_group
        )
    })
  }) %>%
    summarize(
      n = first(n),
      n_train_mean = mean(n_train),
      n_test_mean = mean(n_test),
      .by = age_group
    )

  summary_data %>%
    left_join(sample_sizes, by = "age_group")
}

summary_bioage <- summary_bioage %>%
  add_n_age_groups(
    data = df,
    per_age_group = TRUE
  )
summary_bioage_overall <- summary_bioage_overall %>%
  add_n_age_groups(
    data = df
  )

if (isFALSE(params$telolength)) {
  summary_bioage_reverse <- summary_bioage_reverse %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_bioage_reverse_overall <- summary_bioage_reverse_overall %>%
    add_n_age_groups(
      data = df
    )

  summary_biomarkers <- summary_biomarkers %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_biomarkers_overall <- summary_biomarkers_overall %>%
    add_n_age_groups(
      data = df
    )

  summary_covariates <- summary_covariates %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_covariates_overall <- summary_covariates_overall %>%
    add_n_age_groups(
      data = df
    )
}
```

## Assess rate of aging

### Compare linear vs quadratic vs smooth age fits

We do a partial F-test to compare a full model with a constrained model, to see if the added term
significantly improves fit. We also calculate RMSEs with a weighted mean of residuals. The GAM does
not have explicit parametric linear and quadratic terms but instead has linear and quadratic terms
in the null space, so they're unpenalized, which means they do have the parametric terms, just not
spelled out in the formula. Adding the terms explicitly would make them double. We additionally
calculate partial omega-squared for the effect size of the additional terms. Note that the effect
size for the GAM is approximate, because the degrees of freedom are estimated.

```{r compare_aging_models}
aov_quadratic <- anova(
  models_age_rate$linear,
  models_age_rate$quadratic
)

aov_gam <- anova(
  models_age_rate$quadratic,
  models_age_rate$smooth
)

c(
  linear = sqrt(
    sum(residuals(models_age_rate$linear, type = "response")^2 * df$weights) /
      sum(df$weights)
  ),
  quadratic = sqrt(
    sum(residuals(models_age_rate$quadratic, type = "response")^2 * df$weights) /
      sum(df$weights)
  ),
  smooth = sqrt(
    sum(residuals(models_age_rate$smooth, type = "response")^2 * df$weights) /
      sum(df$weights)
  )
) %>%
  construct_table(
    title = "RMSE (In-Sample)"
  )

calculate_f_test_statistics <- function(anova) {
  anova_table <- tidy(anova) %>%
    rename(
      "numerator df" = df,
      "denominator df" = df.residual,
      "F statistic" = statistic,
      "p" = p.value
    ) %>%
    mutate(
      "Partial Omega2" = `numerator df` * (`F statistic` - 1) / (`numerator df` * `F statistic` + `denominator df` + 1)
    )

  report <- paste0(
    "<i>F</i><sub>(",
    round(anova_table$`numerator df`[2], 1),
    ", ",
    round(anova_table$`denominator df`[2], 1),
    ")</sub> = ",
    round(anova_table$F[2], 2),
    ", <i>p</i> ",
    if (anova_table$p[2] > 0.001) "= " else "< ",
    if (anova_table$p[2] > 0.001) signif(anova_table$p[2], 1) else "0.001",
    ", <i>ω²</i> = ",
    signif(anova_table$`Partial Omega2`[2], 1)
  )

  htmltools::HTML(report)
}

calculate_f_test_statistics(aov_quadratic)
calculate_f_test_statistics(aov_gam)
```

**We see that the quadratic curve fits better than the linear one, so the rate of aging is not
linear, and the smooth curve fits better than the quadratic one, so there are deviations from a
quadratic pattern.**

### Assess where rate of aging differs

By taking the derivative of the full model (with smooth) and subtracting the derivative of the
nested quadratic model, we can test against zero to test deviations from a quadratic pattern.

When making a smooth (or linear) model, we fit a line to noisy data as best we can, but other lines
with a different derivative might fit almost as well and be better for some data points than for
others. Therefore, there is an uncertainty. We can say that, for example, the best-guess slope is 5
(5 increase PTA per year). But because the data is noisy, assuming a normal distribution, we know
that the true slope that would actually fit the sampling population may be something different
within a certain range of uncertainty (say, 2), which we can express as a standard error. When
figuring out the derivative, we inherit this uncertainty. This standard error is what we can use to
assess the likelihood of the null-hypothesis:

**H0**: The GAM does not deviate from the quadratic age-rate function, so the GAM derivative minus
the quadratic derivative equals 0.

**H1**: The GAM does deviate from the quadratic age-rate function, so the GAM derivative minus the
quadratic derivative is not 0.

Depending on the variation/noise in the data, the difference between the quadratic and smooth
derivative needs to be larger or smaller compared to 0 to get a p-value of 0.05. If the variation in
the data is higher, the standard error is higher, and the derivative of the smooth term must be
higher to get a p-value of 0.05 or lower.

We want to know at which age ranges we have a non-zero derivative difference (two-tailed test).
However, rather than perform a t-test and correcting for multiple comparisons, we could use the
multiplicity-adjusted 95% confidence intervals, which directly uses the bootstrap covariance of the
whole plot rather than test per age and adjust for multiplicity. If the 95% confidence intervals do
not include 0, it is equivalent to rejection at family-wise alpha 0.05 for that age. However, it's
hard to tell just how significant something is. We just get a binary decision saying whether it is
above or is below the alpha. The upside is that we don't make assumptions about independence and do
a single test for inference on the whole model. We can report the effect size in the whole model and
report significance at specific ages.

To identify significant boundaries, check which pairwise comparisons are significant, then check if
the previous row was significant, then whenever a row is significant but the previous one was,
change the value by 1 (using `cumsum()`), or if the previous row was not significant but the current
one is, also change by 1 to identify stretches. Then take the first and last values of the stretches
to identify boundaries. Only keep the significant stretches.

```{r find_sig_ages}
pwc_age_rate <- pred_age_rate$derivative_smooth %>%
  mutate(
    signif = case_when(
      ci_lower_adj > 0 ~ "*",
      ci_upper_adj < 0 ~ "*",
      TRUE ~ "ns"
    )
  )

pwc_age_rate_boundaries <- pwc_age_rate %>%
  mutate(
    significant = (ci_lower_adj > 0 | ci_upper_adj < 0),
    run = cumsum(
      significant != lag(
        significant,
        default = first(significant)
      )
    )
  ) %>%
  group_by(run, significant) %>%
  summarise(
    x_min = first(age),
    x_max = last(age),
    .groups = "drop"
  ) %>%
  mutate(
    `Significantly non-linear age ranges` = paste0(x_min, "–", x_max)
  ) %>%
  filter(significant)

pwc_age_rate_boundaries %>%
  select(`Significantly non-linear age ranges`) %>%
  construct_table(
    title = "Boundaries"
  )
```

## Calculate coefficients

We estimate the rate of change in dB/year and (average) acceleration in dB/year² for the quadratic
model and GAM. For the GAM, we estimate the acceleration by fitting a linear model to the derivative
and calculating its coefficient.

```{r calc_coeffs}
with(pred_age_rate$linear, cov(age, estimate) / var(age))
with(pred_age_rate$quadratic_derivative, cov(age, derivative) / var(age))
coef(lm(derivative ~ age, data = pred_age_rate$derivative_smooth))[["age"]]
```

### Print highest/lowest rates of HL

```{r}
pred_age_rate$quadratic_derivative %>%
  filter(age == min(age) | age == max(age))

pred_age_rate$derivative_smooth %>%
  filter(age == min(age) | age == max(age))
```
