---
editor_options:
  markdown:
    wrap: 100
---

# Statistics

## Calculate descriptive statistics

We turn our repaeted design object back into a regular Survey object for calculating population
means. Because no PSUs are filtered out, we can reconstruct the object from the filtered dataset. We
use the full survey design to generate a descriptive statistics table on the filtered and unfiltered
audiometry samples across cycles. We calculate both weighted and unweighted proportions and check n
per stratum. We only include the strata in the unfiltered dataset that's in the filtered dataset and
remove NA values. Bioage descriptive statistics are also optionally calculated. Variance estimates
are calculated with Taylor linearization.

```{r calc_descriptives}
calculate_descriptive_statistics <- function(design, audiometry_cols, incl_bioage = FALSE) {
  design <- svydesign(
    ids = ~psu,
    strata = ~stratum,
    weights = ~weights,
    data = design$variables,
    nest = TRUE
  )
    
  design <- update(
    design,
    sex = dplyr::recode(
      sex,
      "1" = "Male",
      "2" = "Female",
      .default = NA_character_
    ),
    smoking = dplyr::recode(
      smoking,
      "1" = "Smoker",
      "2" = "Non-smoker",
      .default = NA_character_
    ),
    race = dplyr::recode(
      race,
      "1" = "Hispanic",
      "3" = "White",
      "4" = "Black",
      "5" = "Other",
      .default = NA_character_
    ),
    education = dplyr::recode(education,
      "1" = "< 9th grade",
      "2" = "9-11th grade",
      "3" = "High school",
      "4" = "Some college or AA degree",
      "5" = "College graduate or higher",
      .default = NA_character_
    )
  )

  groups <- c(Sex = "sex", Race = "race", Smoking = "smoking", Education = "education")
  age_groups <- factor(c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
  results <- list()

  design <- design %>%
    subset(
      design$variables %>%
        mutate(keep = !if_any(all_of(audiometry_cols), ~ . == 888 | . == 666 | is.na(.))) %>%
        pull(keep) &
        age_group %in% age_groups &
        !is.na(sex) &
        !is.na(smoking) &
        !is.na(race) &
        !is.na(education)
    )


  for (age_group_val in age_groups) {
    design_subset <- if (age_group_val == "Overall") design else subset(design, age_group == age_group_val)

    if (age_group_val == "Overall") {
      proportion_w <- 1
    } else {
      proportion_w <- as.vector(
        coef(svymean(~age_group, design))[paste0("age_group", age_group_val)]
      )
    }

    rows <- bind_rows(
      data.frame(
        age_group = age_group_val,
        group = "Overall",
        category = "Overall",
        n = nrow(design_subset),
        proportion = nrow(design_subset) / nrow(design),
        proportion_w = proportion_w,
        mean_w_pta_hf = coef(svymean(~pta_hf, design_subset)),
        sd_w_pta_hf = sqrt(coef(svyvar(~pta_hf, design_subset))),
        mean_w_pta_lf = coef(svymean(~pta_lf, design_subset)),
        sd_w_pta_lf = sqrt(coef(svyvar(~pta_lf, design_subset))),
        mean_w_age = coef(svymean(~age, design_subset)),
        sd_w_age = sqrt(coef(svyvar(~age, design_subset)))
      ) %>%
        {
          if (isTRUE(incl_bioage)) {
            mutate(.,
              mean_w_kdm = coef(svymean(~kdm, design_subset)),
              mean_w_phenoage = coef(svymean(~phenoage, design_subset)),
              mean_w_hd_log = coef(svymean(~hd_log, design_subset)),
              mean_w_telolength_log = coef(svymean(~telolength_log, design_subset, na.rm = TRUE)),
              sd_w_kdm = sqrt(coef(svyvar(~kdm, design_subset))),
              sd_w_phenoage = sqrt(coef(svyvar(~phenoage, design_subset))),
              sd_w_hd_log = sqrt(coef(svyvar(~hd_log, design_subset))),
              sd_w_telolength_log = sqrt(coef(svyvar(~telolength_log, design_subset, na.rm = TRUE))),
            )
          } else {
            .
          }
        },
      lapply(names(groups), function(group) {
        grouping_variable <- as.formula(paste0("~", groups[[group]]))
        proportions_w <- coef(svymean(grouping_variable, design_subset))
        counts <- table(design_subset$variables[[groups[[group]]]])

        data.frame(
          age_group = age_group_val,
          group = group,
          category = svyby(~pta_hf, grouping_variable, design_subset, svymean)[[groups[[group]]]],
          n = as.vector(counts),
          proportion = as.vector(counts) / nrow(design_subset),
          proportion_w = as.vector(proportions_w),
          mean_w_pta_hf = svyby(~pta_hf, grouping_variable, design_subset, svymean)[["pta_hf"]],
          sd_w_pta_hf = sqrt(svyby(~pta_hf, grouping_variable, design_subset, svyvar)[["pta_hf"]]),
          mean_w_pta_lf = svyby(~pta_lf, grouping_variable, design_subset, svymean)[["pta_lf"]],
          sd_w_pta_lf = sqrt(svyby(~pta_lf, grouping_variable, design_subset, svyvar)[["pta_lf"]]),
          mean_w_age = svyby(~age, grouping_variable, design_subset, svymean)[["age"]],
          sd_w_age = sqrt(svyby(~age, grouping_variable, design_subset, svyvar)[["age"]])
        ) %>%
          {
            if (isTRUE(incl_bioage)) {
              mutate(.,
                mean_w_kdm = svyby(~kdm, grouping_variable, design_subset, svymean)$kdm,
                mean_w_phenoage = svyby(~phenoage, grouping_variable, design_subset, svymean)$phenoage,
                mean_w_hd_log = svyby(~hd_log, grouping_variable, design_subset, svymean)$hd_log,
                mean_w_telolength_log = svyby(~telolength_log, grouping_variable, design_subset, svymean, na.rm = TRUE)$telolength_log,
                sd_w_kdm = sqrt(svyby(~kdm, grouping_variable, design_subset, svyvar)$kdm),
                sd_w_phenoage = sqrt(svyby(~phenoage, grouping_variable, design_subset, svyvar)$phenoage),
                sd_w_hd_log = sqrt(svyby(~hd_log, grouping_variable, design_subset, svyvar)$hd_log),
                sd_w_telolength_log = sqrt(svyby(~telolength_log, grouping_variable, design_subset, svyvar, na.rm = TRUE)$telolength_log)
              )
            } else {
              .
            }
          }
      })
    )

    results[[age_group_val]] <- rows
  }

  bind_rows(results) %>%
    mutate(
      group = factor(group, levels = c("Overall", "Sex", "Race", "Smoking", "Education")),
      age_group = factor(age_group, levels = c("Overall", "20–29", "30–39", "40–49", "50–59", "60–69", if (isFALSE(params$telolength)) "70–79" else NULL))
    )
}

descriptives_unfiltered <- calculate_descriptive_statistics(design_no_missing, audiometry_cols = audiometry_cols)
descriptives <- calculate_descriptive_statistics(design, audiometry_cols = audiometry_cols, incl_bioage = TRUE)
```

## Calculate (delta) MSE per fold

Sample sizes are first computed, either per age group or overall.

Squared residuals are calculated from predicted vs true values, weighted MSEs are calculated and
training and testing sample sizes are stored. Then if there is only one reference, select the
reference column and insert reference mse_w. Use `map_dfr()` to filter predictors to contain
reference, then insert reference in the right spot. We cannot use the filter for the covariate
analysis because it has a full model as reference, so the grepl would fail, therefore we bypass the
filter with the covariate_analysis boolean. There is only one reference variable in this case so we
don't need to filter the dataset first. We calculate `delta_mse_w` by subtracting MSE of the full
model from the MSE of the reduced model, where a positive number means a reduction in MSE over the
reference model (i.e., improvement in error). Because covariate_analysis contains both models that
are larger or smaller than the reference, we set the direction based on case_when.

```{r calculate_per_fold_mse}
calculate_mse_per_fold <- function(predictions, reference, per_age_group = FALSE, covariate_analysis = FALSE) {
  results <- predictions %>%
    mutate(
      squared_residuals = (.data[[params$hearing_var]] - predicted)^2,
      age_group = if (isTRUE(per_age_group)) age_group else "overall"
    ) %>%
    group_by(
      age_group, predictors, r, k
    ) %>%
    summarize(
      sum_weights = sum(weights),
      mse_w = sum(weights * squared_residuals) / sum(weights),
      .groups = "drop"
    )

  results <- map_dfr(reference, \(ref) {
    results %>%
      filter(if (isTRUE(covariate_analysis)) TRUE else grepl(ref, predictors, fixed = TRUE)) %>%
      left_join(
        results %>%
          filter(predictors == ref) %>%
          mutate(
            mse_w_ref = mse_w,
            reference = ref
          ) %>%
          select(reference, age_group, r, k, mse_w_ref),
        by = c("age_group", "r", "k")
      )
  })

  if (isFALSE(covariate_analysis)) {
    results <- results %>%
      mutate(
        delta_mse_w = mse_w_ref - mse_w
      )
  } else {
    results <- results %>%
      mutate(
        delta_mse_w = mse_w - mse_w_ref
      )
  }
  sample_sizes <- predictions %>%
    group_by(age_group, predictors, r, k) %>%
    summarize(
      n_train = first(n_train),
      n_test = first(n_test),
      .groups = "drop"
    )

  if (isFALSE(per_age_group)) {
    sample_sizes <- sample_sizes %>%
      group_by(predictors, r, k) %>%
      summarize(
        n_train = sum(n_train),
        n_test = sum(n_test),
        .groups = "drop"
      ) %>%
      mutate(
        age_group = "overall"
      )
  }
  results %>%
    left_join(
      sample_sizes,
      by = c("predictors", "age_group", "r", "k")
    )
}

results_bioage_overall <- pred_bioage %>%
  calculate_mse_per_fold(
    reference = "age"
  )
results_bioage <- pred_bioage %>%
  calculate_mse_per_fold(
    reference = "age",
    per_age_group = TRUE
  )

if (isFALSE(params$telolength)) {
  results_bioage_reverse_overall <- pred_bioage_reverse %>%
    calculate_mse_per_fold(
      reference = c("phenoage", "kdm", "hd_log")
    )
  results_bioage_reverse <- pred_bioage_reverse %>%
    calculate_mse_per_fold(
      reference = c("phenoage", "kdm", "hd_log"),
      per_age_group = TRUE
    )

  results_biomarkers_overall <- pred_biomarkers %>%
    calculate_mse_per_fold(
      reference = "age"
    )
  results_biomarkers <- pred_biomarkers %>%
    calculate_mse_per_fold(
      reference = "age",
      per_age_group = TRUE
    )

  results_covariates_overall <- pred_covariates %>%
    calculate_mse_per_fold(
      reference = "age, sex, smoking, race, education",
      covariate_analysis = TRUE
    )
  results_covariates <- pred_covariates %>%
    calculate_mse_per_fold(
      reference = "age, sex, smoking, race, education",
      per_age_group = TRUE,
      covariate_analysis = TRUE
    )
}
```

## Calculate summary statistics

### Calculate mean MSE and MSE variance across folds

The Nadeau Bengio corrected t-test estimates variance using the training and testing dataset sizes,
which is needed because cross validations are not independent, violating the assumption of a paired
t-test, inflating type 1 errors (false positives). MSE per fold is calculated using the weighted
mean of prediction errors. Average MSE across folds is calculated by weighting folds by the sum of
the participants' probability weights.

In <https://www.jmlr.org/papers/volume6/markatou05a/markatou05a.pdf>, Analysis of Variance of
Cross-Validation Estimators of the Generalization Error, the authors also report the variance
estimate as $S² * (1/J + n2/n1)$, where $S2$ is the variance (squared standard deviation), `J` is
the total number of folds, $n2$ is the validation/test set and $n1$ is the training set size. They
also show that when $J$ is appropriate ($10*10$), it provides a good variance estimate.

```{r calc_pooled_mse}
calculate_pooled_mse <- function(results) {
  results %>%
    group_by(age_group, predictors) %>%
    mutate(
      mse_w_mean = sum(sum_weights * mse_w) / sum(sum_weights),
      delta_mse_w_mean = sum(sum_weights * delta_mse_w) / sum(sum_weights)
    ) %>%
    summarize(
      reference = first(reference),
      mse_w_ref = sum(sum_weights * mse_w_ref) / sum(sum_weights),
      n = length(unique(r)) * length(unique(k)),
      sum_n_train = sum(n_train),
      sum_n_test = sum(n_test),
      mse_w = sum(sum_weights * mse_w) / sum(sum_weights),
      delta_mse_w = sum(sum_weights * delta_mse_w) / sum(sum_weights),
      s2_mse_w = (sum(sum_weights) / (sum(sum_weights)^2 - sum(sum_weights^2))) * sum(sum_weights * (mse_w - mse_w_mean)^2),
      delta_s2_mse_w = (sum(sum_weights) / (sum(sum_weights)^2 - sum(sum_weights^2))) * sum(sum_weights * (delta_mse_w - delta_mse_w_mean)^2),
      .groups = "drop"
    ) %>%
    mutate(
      se_delta_mse_w_nb = sqrt(
        delta_s2_mse_w * (1 / n + sum_n_test / sum_n_train)
      ),
      se_mse_w_nb = sqrt(
        s2_mse_w * (1 / n + sum_n_test / sum_n_train)
      ),
      n_train = sum_n_train / n,
      n_test = sum_n_test / n
    )
}
```

### Calculate RMSE, R2, and additionally explained variance across folds

In the previous chunk, `mse_w` was calculated by calculating MSE per fold, and calculating a mean
MSE by weighing the per-fold MSEs by the sum of the weights within that fold. This ensures that
folds with subjects with higher weight actually weigh more, rather than just adjusting for
differences in number of subjects per fold. The identical value can be obtained using
`mse_w = sum(weights * squared_residuals) / sum(weights)`, but we cannot derive a variance across
folds directly from residuals this way, so the mean weighted MSE is calculated from per-fold MSEs.
Because the `mse_w` that's computed across folds is already weighted, we can calculate RMSE directly
from it.

We import the weighted mean PTA from the dataframe. This is used as a reference for the R².
Thereafter, we group over the 100 (10x10) folds per model to calculate the R² (using MSE but we
could use also use sum of squares) and RMSE with survey weights taken into account.

**Total sum of squares calculation when binning R² per age group**: It's imperative that for the age
group stratified descriptives, the global mean is used. Otherwise it's as if the total sum of
squares is calculated within each age-bin, which is as if a linear regression is fitted through the
mean PTA per age group, explaining a lot of variation while it shouldn't. We want to assess the
predictive effect on age, so the total sum of squares should be calculated by taking the distance
from each PTA within an age group to the global mean. Of course, the models are trained on the whole
dataset, not just one age group. We're just assessing performance per age group.

Sample sizes, R², and RMSE are inserted into the summary dataframe, and delta RMSEs are calculated
relative to one or multiple reference variables across all ages or within age groups.

```{r calc_stats}
calculate_statistics_across_folds <- function(summary, predictions, data, per_age_group = FALSE, covariate_analysis = FALSE) {
  weighted_mean <- data %>%
    summarize(
      y = sum(weights * .data[[params$hearing_var]]) / sum(weights)
    ) %>%
    pull(y)

  sst <- predictions %>%
    mutate(
      squared_residuals = (.data[[params$hearing_var]] - predicted)^2,
      age_group = if (isTRUE(per_age_group)) age_group else "overall"
    ) %>%
    group_by(
      age_group, predictors
    ) %>%
    summarize(
      sst_w = sum(weights * (.data[[params$hearing_var]] - weighted_mean)^2) / sum(weights),
      .groups = "drop"
    )

  summary <- summary %>%
    left_join(
      sst,
      by = c("age_group", "predictors")
    ) %>%
    mutate(
      r2_w = 1 - (mse_w / sst_w),
      rmse_w = sqrt(mse_w),
      rmse_w_ref = sqrt(mse_w_ref)
    )

  if (isFALSE(covariate_analysis)) {
    summary %>%
      mutate(
        delta_rmse_w = rmse_w_ref - rmse_w,
        r2p = (mse_w_ref - mse_w) / mse_w_ref
      ) %>%
      select(-mse_w_ref, -rmse_w_ref)
  } else {
    summary %>%
      mutate(
        delta_rmse_w = rmse_w - rmse_w_ref,
        r2p = (mse_w - mse_w_ref) / mse_w
      ) %>%
      select(-mse_w_ref, -rmse_w_ref)
  }
}
```

### Calculate RMSE confidence intervals

Calculates squared residuals (error) to be used for RMSE calculations and takes mean MSE per subject
over the 10 repeats. Takes reference variable to be the full model or age model. Then for each
predictor (and optionally each age group) subsets the data to a dataframe with the reference
variable and with the target variable. These are equal length. Then takes random samples from the
dataset with number of samples equal to the length of the subset (i.e., age group). The indices of
each random sample is stored in a separate column, creating a matrix. The indices are then applied
to the vector with squared residuals. The column mean is then calculated for each column (i.e.,
bootstrap sample) within the matrix. The square root is then taken from each mean to get the RMSE
per sample. 2.5% and 97.5% quantiles are then computed to obtain 95% confidence intervals for both
raw RMSEs and delta RMSEs. The selected rows are always the same for target and reference variables.
Indices are calculated once for each age group and stored in a cache so that the next predictor uses
the same indices as the other predictor per age group. Each bootstrap run creates a different sample
of the data because observations are selected at random without requirement to be unique, so some
observations are taken twice and others are omitted (set by `replace = TRUE`). If
`per_age_group = TRUE`, iterate over each age group and subset df per age group. If
`per_age_group = FALSE`, only iterate over the single "all" item that is not used (that is, run the
code once and don't filter per age group). Compile into a dataframe (tidytable) at the end and
insert values into the summary dataframes.

We flip the sign and order of upper and lower CIs of covariates. Probability weights are applied to
calculate weighted column means. We do not need to use PSUs for sampling because we want to know the
uncertainty over observed individuals' prediction errors, not estimating model coefficient
uncertainty. We are bootstrapping based on already-computed prediction errors, not refitting GAMs
inside each bootstrap.

```{r calc_RMSE_CI}
compute_rmse_ci <- function(
  summary,
  predictions,
  n_boot,
  per_age_group = FALSE,
  covariate_analysis = FALSE
) {
  predictor_list <- unique(predictions$predictors)
  age_bins <- if (isTRUE(per_age_group)) unique(predictions$age_group) else "overall"

  predictor_ref_map <- summary %>%
    select(predictors, reference) %>%
    distinct()

  ci_list <- list()
  index_ci_list <- 1L
  index_cache <- list()

  predictions <- predictions %>%
    mutate(squared_residuals = (.data[[params$hearing_var]] - predicted)^2) %>%
    group_by(predictors, age_group, SEQN) %>%
    summarize(
      squared_residuals = mean(squared_residuals),
      weights = first(weights),
      .groups = "drop"
    )

  set.seed(43)

  for (predictor in predictor_list) {
    ref_var <- predictor_ref_map %>%
      filter(predictors == predictor) %>%
      pull(reference)
    for (age_bin in age_bins) {
      error_ref <- predictions %>%
        filter(
          predictors == ref_var,
          if (isTRUE(per_age_group)) age_group == age_bin else TRUE
        ) %>%
        pull(squared_residuals)
      w_ref <- predictions %>%
        filter(
          predictors == ref_var,
          if (isTRUE(per_age_group)) age_group == age_bin else TRUE
        ) %>%
        pull(weights)
      w_ref <- w_ref / mean(w_ref)

      error_target <- predictions %>%
        filter(
          predictors == predictor,
          if (isTRUE(per_age_group)) age_group == age_bin else TRUE
        ) %>%
        pull(squared_residuals)
      w_target <- predictions %>%
        filter(
          predictors == predictor,
          if (isTRUE(per_age_group)) age_group == age_bin else TRUE
        ) %>%
        pull(weights)
      w_target <- w_target / mean(w_target)

      if (length(error_ref) != length(error_target)) {
        stop("Length of target and reference predictor subsets are not equal")
      }
      if (any(w_target != w_ref)) {
        stop("Target and reference weights are different, SEQN likely not matching")
      }

      if (!age_bin %in% names(index_cache)) {
        index_cache[[age_bin]] <- matrix(
          sample(
            length(error_ref),
            length(error_ref) * n_boot,
            replace = TRUE
          ),
          length(error_ref)
        )
      }

      idx <- index_cache[[age_bin]]

      rmse_target <- sqrt(
        apply(idx, 2, function(i) sum(error_target[i] * w_target[i]) / sum(w_target[i]))
      )
      rmse_ref <- sqrt(
        apply(idx, 2, function(i) sum(error_ref[i] * w_ref[i]) / sum(w_ref[i]))
      )

      ci_list[[index_ci_list]] <- tibble(
        ci_low_rmse = quantile(rmse_target, 0.025),
        ci_high_rmse = quantile(rmse_target, 0.975),
        ci_low_delta_rmse = quantile(rmse_ref - rmse_target, 0.025),
        ci_high_delta_rmse = quantile(rmse_ref - rmse_target, 0.975),
        predictors = predictor
      ) %>%
        mutate(age_group = age_bin)

      index_ci_list <- index_ci_list + 1
    }
  }

  ci_df <- bind_rows(ci_list)

  if (isTRUE(covariate_analysis)) {
    ci_df <- ci_df %>%
      mutate(
        temp = ci_low_delta_rmse,
        ci_low_delta_rmse = ci_high_delta_rmse * -1,
        ci_high_delta_rmse = temp * -1
      ) %>%
      select(-temp)
  }

  summary %>%
    left_join(ci_df, by = c("predictors", "age_group"))
}
```

### Run descriptive statistics calculation pipeline

```{r descr_stat_pipeline}
summary_bioage_overall <- results_bioage_overall %>%
  calculate_pooled_mse() %>%
  calculate_statistics_across_folds(
    predictions = pred_bioage,
    data = df
  ) %>%
  compute_rmse_ci(
    predictions = pred_bioage,
    n_boot = params$n_boot
  )
summary_bioage <- results_bioage %>%
  calculate_pooled_mse() %>%
  calculate_statistics_across_folds(
    predictions = pred_bioage,
    data = df,
    per_age_group = TRUE
  ) %>%
  compute_rmse_ci(
    predictions = pred_bioage,
    n_boot = params$n_boot,
    per_age_group = TRUE
  )

if (isFALSE(params$telolength)) {
  summary_bioage_reverse_overall <- results_bioage_reverse_overall %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_bioage_reverse,
      data = df
    ) %>%
    compute_rmse_ci(
      predictions = pred_bioage_reverse,
      n_boot = params$n_boot
    )
  summary_bioage_reverse <- results_bioage_reverse %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_bioage_reverse,
      data = df,
      per_age_group = TRUE
    ) %>%
    compute_rmse_ci(
      predictions = pred_bioage_reverse,
      n_boot = params$n_boot,
      per_age_group = TRUE
    )

  summary_biomarkers_overall <- results_biomarkers_overall %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_biomarkers,
      data = df
    ) %>%
    compute_rmse_ci(
      predictions = pred_biomarkers,
      n_boot = params$n_boot
    )
  summary_biomarkers <- results_biomarkers %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_biomarkers,
      data = df,
      per_age_group = TRUE
    ) %>%
    compute_rmse_ci(
      predictions = pred_biomarkers,
      n_boot = params$n_boot,
      per_age_group = TRUE
    )

  summary_covariates_overall <- results_covariates_overall %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_covariates,
      data = df,
      covariate_analysis = TRUE
    ) %>%
    compute_rmse_ci(
      predictions = pred_covariates,
      n_boot = params$n_boot,
      covariate_analysis = TRUE
    )
  summary_covariates <- results_covariates %>%
    calculate_pooled_mse() %>%
    calculate_statistics_across_folds(
      predictions = pred_covariates,
      data = df,
      per_age_group = TRUE,
      covariate_analysis = TRUE
    ) %>%
    compute_rmse_ci(
      predictions = pred_covariates,
      n_boot = params$n_boot,
      per_age_group = TRUE,
      covariate_analysis = TRUE
    )
}
```

## Repeated Nadeau-Bengio corrected t-tests

### Calculate p-values

Calculate t-statistic according to Nadeau Bengio (although we use a weighted adaptation of it),
which is `t = m/(√(variance(1/n + n2/n1)))`. `m` is the difference between the sample mean and
hypothesized population mean. In this case, we've normalized it to age, so the null hypothesis (the
hypothesized population mean) is 0, i.e., there is no difference from age only. The variance was
already calculated before. `n` is equivalent to the number of folds times the number of repeats.
`n2` is the number of testing/validation data and `n1` the number of training data. We then
calculate degrees of freedom by taking n - 1. Then we use `pt()` (which automatically calculates the
critical value) using a one-tailed t-test of the upper tail. This is because we're not interested in
whether an age + bioage model is different from age-only. We only want to know if it is better. We
then adjust p values with Holm error correction.

First, the question is, 'Does this bioage/biomarker add predictive value for hearing loss in the
population while controlling for chronological age?', which we test with the overall comparisons.
Only if it's significantly different do we look at per age-group tests. Another option is to ask
'Might bioage/biomarker only matter for specific age ranges, even if it doesn't have an overall
predictive effect?', in which case we could consider doing pairwise on all, but we'd lose a lot of
power and would have to use FDR and accept a 5% false discovery rate to salvage it.

```{r t_test_models}
nb_corrected_repeated_t_test <- function(summary,
                                         exceptions = NULL,
                                         exception_include_exclude,
                                         adjust_method) {
  if ("p" %in% colnames(summary)) {
    summary <- summary %>%
      select(
        -t, -df, -p, -p_adj, -signif
      )
  }
  if (exception_include_exclude == "include") {
    pwc <- summary %>%
      filter(predictors %in% exceptions)
  } else {
    pwc <- summary %>%
      filter(!predictors %in% exceptions)
  }

  pwc <- pwc %>%
    mutate(
      t = delta_mse_w / se_delta_mse_w_nb,
      df = n - 1,
      p = pt(t, df = df, lower.tail = FALSE),
      p_adj = p.adjust(p, method = adjust_method),
      signif = case_when(
        p_adj <= 0.001 ~ "***",
        p_adj <= 0.01 ~ "**",
        p_adj <= 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(
      age_group, predictors, t, df, p, p_adj, signif
    )

  summary <- summary %>%
    left_join(pwc,
      by = c("age_group", "predictors")
    )

  summary
}

summary_bioage_overall <- nb_corrected_repeated_t_test(
  summary_bioage_overall,
  exceptions = c(
    "age",
    if (isTRUE(params$telolength)) "age, phenoage" else NULL,
    if (isTRUE(params$telolength)) "age, kdm" else NULL,
    if (isTRUE(params$telolength)) "age, hd" else NULL
  ),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)

summary_bioage <- nb_corrected_repeated_t_test(
  summary_bioage,
  exceptions = c("age, phenoage", "age, kdm"),
  exception_include_exclude = "include",
  adjust_method = "holm"
)

if (isFALSE(params$telolength)) {
  summary_bioage_reverse_overall <- nb_corrected_repeated_t_test(
    summary_bioage_reverse_overall,
    exceptions = c("phenoage", "kdm", "hd_log"),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_bioage_reverse <- nb_corrected_repeated_t_test(
    summary_bioage_reverse,
    exceptions = c("phenoage", "kdm", "hd_log"),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_biomarkers_overall <- nb_corrected_repeated_t_test(
    summary_biomarkers_overall,
    exceptions = c(
      "age"
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_biomarkers <- nb_corrected_repeated_t_test(
    summary_biomarkers,
    exceptions = c(
      "crp"
    ),
    exception_include_exclude = "include",
    adjust_method = "holm"
  )

  summary_covariates_overall <- nb_corrected_repeated_t_test(
    summary_covariates_overall,
    exceptions = c(
      "age, sex, smoking, race, education"
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )

  summary_covariates <- nb_corrected_repeated_t_test(
    summary_covariates,
    exceptions = c(
      "age, sex, smoking, race, education",
      "age, sex, race, education",
      if (params$hearing_var == "pta_lf") "age, sex, smoking, education" else NULL,
      if (params$hearing_var == "pta_lf") "age, smoking, race, education" else NULL
    ),
    exception_include_exclude = "exclude",
    adjust_method = "holm"
  )
}
```

### Display summary tables

#### Make rounding function

We round to a max of 3 decimals (0.001) but scale it to the size, so 9.12312 becomes 9.12, 91.1231
becomes 91.1, 912.123123 becomes 912, and anything larger than or equal to 1000 gets rounded to an
integer (so that 12345 doesn't become 12300).

```{r round_function}
clean_rounding <- function(x, max_decimals = 3) {
  ifelse(
    abs(x) >= 10^max_decimals,
    round(x),
    round(signif(x, max_decimals), max_decimals)
  )
}
```

#### Bioage overall

```{r pwc_bioage_overall}
summary_bioage_overall %>%
  mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
  construct_table(
    title = "Overall Bioage Pairwise Comparisons"
  )
```

#### Bioage

```{r pwc_bioage}
summary_bioage %>%
  mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
  construct_table(
    title = "Per-Age Group Bioage Pairwise Comparisons"
  )
```

#### Bioage reverse overall

```{r pwc_bioage_reverse_overall}
if (isFALSE(params$telolength)) {
  summary_bioage_reverse_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Bioage Reverse Analysis Pairwise Comparisons"
    )
}
```

#### Bioage reverse

```{r pwc_bioage_reverse}
if (isFALSE(params$telolength)) {
  summary_bioage_reverse %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Bioage Reverse Analysis Pairwise Comparisons"
    )
}
```

#### Biomarkers overall

```{r pwc_biomarkers_overall}
if (isFALSE(params$telolength)) {
  summary_biomarkers_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Biomarker Pairwise Comparisons"
    )
}
```

#### Biomarkers

```{r pwc_biomarkers}
if (isFALSE(params$telolength)) {
  summary_biomarkers %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Biomarker Pairwise Comparisons"
    )
}
```

#### Covariates overall

```{r pwc_covariates_overall}
if (isFALSE(params$telolength)) {
  summary_covariates_overall %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Overall Covariate Pairwise Comparisons"
    )
}
```

#### Covariates

```{r pwc_covariates}
if (isFALSE(params$telolength)) {
  summary_covariates %>%
    mutate(across(where(is.numeric), ~ clean_rounding(.x))) %>%
    construct_table(
      title = "Per-Age Group Covariate Pairwise Comparisons"
    )
}
```

### Add n training and testing datasets to dataframe

```{r add_n}
add_n_age_groups <- function(summary_data, data, per_age_group = FALSE) {
  if (per_age_group) {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      left_join(
        data %>%
          select(age_group) %>%
          group_by(age_group) %>%
          summarize(n = n()),
        by = "age_group"
      )
  } else {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      mutate(n = nrow(data))
  }

  summary_data
}

summary_bioage <- summary_bioage %>%
  add_n_age_groups(
    data = df,
    per_age_group = TRUE
  )
summary_bioage_overall <- summary_bioage_overall %>%
  add_n_age_groups(
    data = df
  )

if (isFALSE(params$telolength)) {
  summary_bioage_reverse <- summary_bioage_reverse %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_bioage_reverse_overall <- summary_bioage_reverse_overall %>%
    add_n_age_groups(
      data = df
    )

  summary_biomarkers <- summary_biomarkers %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_biomarkers_overall <- summary_biomarkers_overall %>%
    add_n_age_groups(
      data = df
    )

  summary_covariates <- summary_covariates %>%
    add_n_age_groups(
      data = df,
      per_age_group = TRUE
    )
  summary_covariates_overall <- summary_covariates_overall %>%
    add_n_age_groups(
      data = df
    )
}
```

## Assess rate of aging

### Compare linear vs quadratic vs smooth age fits

We do a partial F-test to compare a full model with a constrained model, to see if the added term
significantly improves fit. We also calculate RMSEs with a weighted mean of residuals. The GAM does
not have explicit parametric linear and quadratic terms but instead has linear and quadratic terms
in the null space, so they're unpenalized, which means they do have the parametric terms, just not
spelled out in the formula. Adding the terms explicitly would make them double. We additionally
calculate partial omega-squared for the effect size of the additional terms. Note that the effect
size for the GAM is approximate, because the degrees of freedom are estimated.

```{r compare_aging_models}
aov_quadratic <- anova(
  models_age_rate$linear,
  models_age_rate$quadratic
)

aov_gam <- anova(
  models_age_rate$quadratic,
  models_age_rate$smooth
)

c(
  linear = sqrt(
    sum(residuals(models_age_rate$linear, type = "response")^2 * df$weights) /
      sum(df$weights)
  ),
  quadratic = sqrt(
    sum(residuals(models_age_rate$quadratic, type = "response")^2 * df$weights) /
      sum(df$weights)
  ),
  smooth = sqrt(
    sum(residuals(models_age_rate$smooth, type = "response")^2 * df$weights) /
      sum(df$weights)
  )
) %>%
  construct_table(
    title = "RMSE (In-Sample)"
  )

calculate_f_test_statistics <- function(anova) {
  anova_table <- tidy(anova) %>%
    rename(
      "numerator df" = df,
      "denominator df" = df.residual,
      "F statistic" = statistic,
      "p" = p.value
    ) %>%
    mutate(
      "Partial Omega2" = `numerator df` * (`F statistic` - 1) / (`numerator df` * `F statistic` + `denominator df` + 1)
    )

  report <- paste0(
    "<i>F</i><sub>(",
    round(anova_table$`numerator df`[2], 1),
    ", ",
    round(anova_table$`denominator df`[2], 1),
    ")</sub> = ",
    round(anova_table$F[2], 2),
    ", <i>p</i> ",
    if (anova_table$p[2] > 0.001) "= " else "< ",
    if (anova_table$p[2] > 0.001) signif(anova_table$p[2], 1) else "0.001",
    ", <i>ω²</i> = ",
    signif(anova_table$`Partial Omega2`[2], 1)
  )

  htmltools::HTML(report)
}

calculate_f_test_statistics(aov_quadratic)
calculate_f_test_statistics(aov_gam)
```

**We see that the quadratic curve fits better than the linear one, so the rate of aging is not
linear, and the smooth curve fits better than the quadratic one, so there are deviations from a
quadratic pattern.**

### Assess where rate of aging differs

By taking the derivative of the full model (with smooth) and subtracting the derivative of the
nested quadratic model, we can test against zero to test deviations from a quadratic pattern.

When making a smooth (or linear) model, we fit a line to noisy data as best we can, but other lines
with a different derivative might fit almost as well and be better for some data points than for
others. Therefore, there is an uncertainty. We can say that, for example, the best-guess slope is 5
(5 increase PTA per year). But because the data is noisy, assuming a normal distribution, we know
that the true slope that would actually fit the sampling population may be something different
within a certain range of uncertainty (say, 2), which we can express as a standard error. When
figuring out the derivative, we inherit this uncertainty. This standard error is what we can use to
assess the likelihood of the null-hypothesis:

**H0**: The GAM does not deviate from the quadratic age-rate function, so the GAM derivative minus
the quadratic derivative equals 0.

**H1**: The GAM does deviate from the quadratic age-rate function, so the GAM derivative minus the
quadratic derivative is not 0.

Depending on the variation/noise in the data, the difference between the quadratic and smooth
derivative needs to be larger or smaller compared to 0 to get a p-value of 0.05. If the variation in
the data is higher, the standard error is higher, and the derivative of the smooth term must be
higher to get a p-value of 0.05 or lower.

We want to know at which age ranges we have a non-zero derivative difference (two-tailed test).
However, rather than perform a t-test and correcting for multiple comparisons, we could use the
multiplicity-adjusted 95% confidence intervals, which directly uses the bootstrap covariance of the
whole plot rather than test per age and adjust for multiplicity. If the 95% confidence intervals do
not include 0, it is equivalent to rejection at family-wise alpha 0.05 for that age. However, it's
hard to tell just how significant something is. We just get a binary decision saying whether it is
above or is below the alpha. The upside is that we don't make assumptions about independence and do
a single test for inference on the whole model. We can report the effect size in the whole model and
report significance at specific ages.

To identify significant boundaries, check which pairwise comparisons are significant, then check if
the previous row was significant, then whenever a row is significant but the previous one was,
change the value by 1 (using `cumsum()`), or if the previous row was not significant but the current
one is, also change by 1 to identify stretches. Then take the first and last values of the stretches
to identify boundaries. Only keep the significant stretches.

```{r find_sig_ages}
pwc_age_rate <- pred_age_rate$derivative_smooth %>%
  mutate(
    signif = case_when(
      ci_lower_adj > 0 ~ "*",
      ci_upper_adj < 0 ~ "*",
      TRUE ~ "ns"
    )
  )

pwc_age_rate_boundaries <- pwc_age_rate %>%
  mutate(
    significant = (ci_lower_adj > 0 | ci_upper_adj < 0),
    run = cumsum(
      significant != lag(
        significant,
        default = first(significant)
      )
    )
  ) %>%
  group_by(run, significant) %>%
  summarise(
    x_min = first(age),
    x_max = last(age),
    .groups = "drop"
  ) %>%
  mutate(
    `Significantly non-linear age ranges` = paste0(x_min, "–", x_max)
  ) %>%
  filter(significant)

pwc_age_rate_boundaries %>%
  select(`Significantly non-linear age ranges`) %>%
  construct_table(
    title = "Boundaries"
  )
```

## Calculate coefficients

We estimate the rate of change in dB/year and (average) acceleration in dB/year² for the quadratic
model and GAM. For the GAM, we estimate the acceleration by fitting a linear model to the derivative
and calculating its coefficient.

```{r calc_coeffs}
with(pred_age_rate$linear, cov(age, estimate) / var(age))
with(pred_age_rate$quadratic_derivative, cov(age, derivative) / var(age))
coef(lm(derivative ~ age, data = pred_age_rate$derivative_smooth))[["age"]]
```

### Print highest/lowest rates of HL

```{r}
pred_age_rate$quadratic_derivative %>%
  filter(age == min(age) | age == max(age))

pred_age_rate$derivative_smooth %>%
  filter(age == min(age) | age == max(age))
```

