---
editor_options:
  markdown:
    wrap: 100
---

# Visualization

This section contains the main results and figures. Note that some supplementary figures are in
02_data_preprocessing or 03_modeling.

## Tables

#### Descriptives table

#### Format table

Round R² and delta RMSE, rename variables, set p values to 'below 0.001' when very small.

```{r format_table}
format_table <- function(summary_data, per_age_group = FALSE) {
  table <- summary_data %>%
    mutate(
      r2 = round(r2_w, 3),
      delta_rmse = round(delta_rmse_w, 2)
    ) %>%
    rename(
      "Predictors" = predictors,
      "∆RMSE (vs age, dB HL)" = delta_rmse,
      "R²" = r2,
      "p (vs age)" = p_adj
    ) %>%
    mutate(
      `p (vs age)` = as.character(round(`p (vs age)`, 3)),
      `p (vs age)` = case_when(
        signif == "***" ~
          cell_spec("< 0.001", "html", bold = TRUE),
        suppressWarnings(
          !is.na(as.numeric(`p (vs age)`)) &
            as.numeric(`p (vs age)`) < 0.05
        ) &
          `p (vs age)` != "< 0.001" ~
          cell_spec(
            suppressWarnings(
              round(as.numeric(`p (vs age)`), 3)
            ),
            "html",
            bold = TRUE
          ),
        TRUE ~ `p (vs age)`
      ),
      `p (vs age)` = case_when(
        Predictors == "age" ~ "",
        TRUE ~ `p (vs age)`
      ),
      `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
      `∆RMSE (vs age, dB HL)` = case_when(
        Predictors == "age" ~ "",
        TRUE ~ `∆RMSE (vs age, dB HL)`
      )
    )

  if (isTRUE(per_age_group)) {
    table <- table %>%
      rename("Age" = age_group) %>%
      select(Predictors, Age, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  } else {
    table <- table %>%
      select(Predictors, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  }

  table
}
```

#### Store biomarker labels

```{r}
biomarker_labels <- summary_biomarkers_overall %>%
  mutate(
    label =
      c(
        "Age",
        "GLU",
        "ALB",
        "ALP",
        "HbA1c",
        "CREA",
        "BASO",
        "Ca",
        "Cl",
        "HDL-C",
        "BUN",
        "Hb",
        "HCT",
        "Fe",
        "LDH",
        "LYMPH",
        "MCH",
        "MCHC",
        "Mono",
        "NEUT",
        "P",
        "K",
        "RBC",
        "RDW",
        "MCV",
        "Na",
        "TC",
        "TP",
        "UA",
        "WBC",
        "SBP",
        "ALT",
        "AST",
        "GGT",
        "CRP",
        if (isTRUE(params$run_biomarker_add_analysis)) "PhenoAge" else NULL,
        if (isTRUE(params$run_biomarker_add_analysis)) "CRP + PhenoAge" else NULL
      )
  ) %>%
  arrange(desc(delta_rmse_w)) %>%
  select(predictors, label)
```

#### Order, filter, and rename tables

```{r order_table}
rename_bioage_table <- function(bioage_table) {
  bioage_table <- bioage_table %>%
    mutate(
      Predictors = case_when(
        Predictors == "age" ~ "Age",
        Predictors == "age, hd_log" ~ "Age + HD",
        Predictors == "age, kdm" ~ "Age + KDM",
        Predictors == "age, phenoage" ~ "Age + PhenoAge",
        isTRUE(params$telolength) & Predictors == "age, telolength_log" ~ "Age + Telolength",
        TRUE ~ Predictors
      ),
      Predictors = factor(
        Predictors,
        levels = if (isTRUE(params$telolength)) {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge", "Age + Telolength")
        } else {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge")
        }
      )
    ) %>%
    arrange(Predictors)

  bioage_table
}

rename_biomarker_table <- function(biomarker_table) {
  biomarker_table <- biomarker_table %>%
    filter(
      !(Predictors %in% c(
        "age",
        "age, phenoage",
        "age, lncrp, phenoage"
      ))
    ) %>%
    mutate(
      Predictors = factor(Predictors, levels = biomarker_labels$predictors, labels = paste0("Age + ", biomarker_labels$label))
    ) %>%
    arrange(Predictors)

  biomarker_table
}

rename_covariate_table <- function(covariates_table) {
  covariates_table <- covariates_table %>%
    filter(
      Predictors != "age, sex, smoking, race, education"
    ) %>%
    mutate(
      Predictors = case_when(
        Predictors == "age, sex, smoking, race, education, phenoage" ~ "Phenoage",
        Predictors == "age, smoking, race, education" ~ "Gender",
        Predictors == "age, sex, race, education" ~ "Smoking",
        Predictors == "age, sex, smoking, education" ~ "Race",
        Predictors == "age, sex, smoking, race" ~ "Education"
      ),
      Predictors = factor(Predictors, levels = c(
        "Phenoage",
        "Gender",
        "Smoking",
        "Race",
        "Education",
        "Gender * Race"
      ))
    ) %>%
    rename(
      "p (vs full model)" = `p (vs age)`,
      "∆RMSE (vs full model, dB HL)" = `∆RMSE (vs age, dB HL)`,
    ) %>%
    arrange(Predictors)

  covariates_table
}
```

#### Run table formatting pipeline

```{r table_format_pipe}
table_bioage <- summary_bioage %>%
  format_table(
    per_age_group = TRUE
  ) %>%
  rename_bioage_table()

table_bioage_overall <- summary_bioage_overall %>%
  format_table() %>%
  rename_bioage_table()

if (isFALSE(params$telolength)) {
  table_biomarkers <- summary_biomarkers %>%
    format_table(
      per_age_group = TRUE
    ) %>%
    rename_biomarker_table()

  table_biomarkers_overall <- summary_biomarkers_overall %>%
    format_table() %>%
    rename_biomarker_table()

  table_covariates <- summary_covariates %>%
    format_table(
      per_age_group = TRUE
    ) %>%
    rename_covariate_table()

  table_covariates_overall <- summary_covariates_overall %>%
    format_table() %>%
    rename_covariate_table()
}
```

#### Render table

```{r store_tables}
render_and_save_table <- function(table, path) {
  html_table <- table %>%
    kable(
      format = "html",
      align = c("l", "l", "r", "r", "r", "r", "r", "r", "r"),
      escape = FALSE
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "center"
    ) %>%
    collapse_rows(columns = 1, valign = "top")

  if (isTRUE(params$save_figures)) {
    save_kable(html_table, path)
  }

  html_table
}

render_and_save_table(
  table = table_bioage,
  path = here("output", "figures", params$hearing_var, "table_bioage.html")
)

render_and_save_table(
  table = table_bioage_overall,
  path = here("output", "figures", params$hearing_var, "table_bioage_overall.html")
)

if (isFALSE(params$telolength)) {
  render_and_save_table(
    table = table_biomarkers,
    path = here("output", "figures", params$hearing_var, "table_biomarkers.html")
  ) %>%
    scroll_box(width = "100%", height = "400px")

  render_and_save_table(
    table = table_biomarkers_overall,
    path = here("output", "figures", params$hearing_var, "table_biomarkers_overall.html")
  )

  render_and_save_table(
    table = table_covariates,
    path = here("output", "figures", params$hearing_var, "table_covariates.html")
  ) %>%
    scroll_box(width = "100%", height = "400px")

  render_and_save_table(
    table = table_covariates_overall,
    path = here("output", "figures", params$hearing_var, "table_covariates_overall.html")
  )
}
```

### Covariates table

#### Create covariate table and format covariate table

We recode variables, combine means and SDs, pivot table, and format it

```{r pivot_conf_table}
if (isFALSE(params$telolength)) {
  table_covariates_descriptive <- descriptives %>%
    filter(
      age_group != "Overall",
      group != "Overall"
    ) %>%
    mutate(
      stat = sprintf("%.2f (%.2f)", mean_w_pta, sd_w_pta),
      category = factor(category, levels = c(
        "Male",
        "Female",
        "White",
        "Hispanic",
        "Black",
        "Other",
        "Smoker",
        "Non-smoker",
        "< 9th grade",
        "9-11th grade",
        "High school",
        "Some college or AA degree",
        "College graduate or higher"
      ))
    ) %>%
    select(
      age_group, group, category, stat
    ) %>%
    pivot_wider(names_from = age_group, values_from = stat) %>%
    group_by(group) %>%
    mutate(group = if_else(row_number() == 1L, group, "")) %>%
    ungroup() %>%
    rename(" " = group, "  " = category)
}
```

#### Save table

```{r store_conf_table}
if (isFALSE(params$telolength)) {
  render_and_save_table(
    table_covariates_descriptive,
    path = here(
      "output",
      "figures",
      params$hearing_var,
      "table_covariates_descriptives.html"
    )
  )
}
```

## Figures

### Set plot config

#### Specify plot styling variables

We store a separate configuration depending on output type of figure for easy switching between
output types. Only the currently selected output type variables in params + global variables for
plotting we may need later is returned. These variables are used to construct ggplot protoobjects
and can be accessed either with `plot_config()`, which returns all, or with
`plot_config("parameter")` to return a specific parameter.

```{r fun_plot_cfg}
plot_config <- function(variable = NULL, global_params = params) {
  p_cfg_vars <- list(
    global = list(
      axis_unit_color = "#4D4D4D" # used later when constructing axes from HTML
    ),
    poster = list(),
    paper = list(
      figure_dir = here(
        "output",
        "figures",
        if (isTRUE(global_params$telolength)) {
          paste0(global_params$hearing_var, "_telolength_log")
        } else {
          global_params$hearing_var
        }
      ),
      text_size_primary = 8,
      text_size_secondary = 9,
      signif_size = 14
    )
  )

  p_cfg_vars_comb <- c(p_cfg_vars$global, p_cfg_vars[[global_params$plot_output_type]])

  if (!is.null(variable)) p_cfg_vars_comb[[variable]] else p_cfg_vars_comb
}
```

#### Specify general plot theme

We first store the theme preset (theme_few), then theme overrides, then other theme items like
guides. All items specified in `plot_theme()` must be ggplot proto objects.

```{r fun_plot_theme}
plot_theme <- function(global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  list(
    theme_few(),
    theme(
      axis.text.x = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      axis.text.y = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      strip.text.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      strip.text.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ),
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = config$text_size_secondary,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
  )
}
```

#### Specify axis scales

```{r fun_plot_axes}
axes <- function(x = NULL, y = NULL, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  if (!is.null(x)) {
    x <- switch(x,
      age = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(20, 80, 10),
        limits = c(20, 80),
        name = "Age",
        unit = "Years"
      ),
      age_group = define_axis(
        config = config,
        axis = "x",
        type = "discrete",
        limits = c(1.4, 5.7),
        name = "Age",
        unit = "Years"
      ),
      age_accel_phenoage = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-20, 20),
        name = "PhenoAge Age Acceleration",
        unit = "PhenoAge – Age (Years)"
      ),
      age_accel_kdm = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-20, 20),
        name = "KDM Age Acceleration",
        unit = "KDM – Age (Years)"
      )
    )
  }

  if (!is.null(y)) {
    y <- switch(y,
      age_group = define_axis(
        config = config,
        axis = "y",
        type = "discrete",
        limits = c(1, 7),
        name = "Age",
        unit = "Years"
      ),
      hl = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 90, 30),
        limits = c(-10, 100),
        name = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA"),
        unit = "dB HL"
      ),
      hl_rate = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(-1, 2, 1),
        limits = c(-1.1, 2.6),
        name = "Rate of HL",
        unit = "dB HL/Year"
      ),
      hl_diff = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(-1, 2, 1),
        limits = c(-1.1, 2.8),
        name = "Deviation Rate of HL",
        unit = "dB HL/Year"
      ),
      dmse_bioage = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 8, 2),
        limits = c(-2, 10),
        name = "Additional Predictive Capacity",
        unit = "ΔMSE (dB HL)<sup>2</sup>"
      ),
      drmse_pred = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 0.3, 0.1),
        limits = c(-0.1, 0.35),
        name = "Additional Predictive Capacity",
        unit = "ΔRMSE (dB HL)"
      ),
      drmse_biomark = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 0.1, 0.05),
        limits = c(-0.025, 0.1),
        name = "Additional Predictive Capacity",
        unit = "ΔRMSE (dB HL)"
      ),
      drmse_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 2, 0.5),
        limits = c(-0.05, 2.5),
        name = "Additional Predictive Capacity",
        unit = "ΔRMSE (dB HL)"
      ),
      dmse_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 40, 5),
        limits = c(-1, 40),
        name = "Additional Predictive Capacity",
        unit = "ΔMSE (dB HL)<sup>2</sup>"
      ),
      add_pred_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 14, 2),
        limits = c(-1, 15),
        name = "Additional Predictive Capacity",
        unit = "% Change MSE"
      )
    )
  }

  build_axes(x = x, y = y)
}
```

#### Specify colors

```{r fun_plot_colors}
color_scales <- function(color_list_name) {
  switch(color_list_name,
    bioage = build_color_scale(
      colors = c(
        "age" = "#000000",
        "age, hd_log" = if (isFALSE(params$telolength)) color("highcontrast")(3)[1] else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) color("highcontrast")(3)[2] else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) color("highcontrast")(3)[3] else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "#7B3A96" else NULL
      ),
      fill_colors = c(
        "age" = "#000000",
        "age, hd_log" = if (isFALSE(params$telolength)) color("highcontrast")(3)[1] else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) color("highcontrast")(3)[2] else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) color("highcontrast")(3)[3] else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "#B276B2" else NULL
      ),
      labels = c(
        "age" = "Age",
        "age, hd_log" = if (isFALSE(params$telolength)) "Age + HD" else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) "Age + KDM" else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) "Age + PhenoAge" else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "Age + Telolength" else NULL
      )
    ),
    covariates = build_color_scale(
      colors = c(
        "age, smoking, race, education" = color("bright")(4)[1],
        "age, sex, smoking, education" = color("bright")(4)[2],
        "age, sex, smoking, race" = color("bright")(4)[3],
        "age, sex, race, education" = color("bright")(4)[4]
      ),
      fill_colors = c(
        "age, smoking, race, education" = color("bright")(4)[1],
        "age, sex, smoking, education" = color("bright")(4)[2],
        "age, sex, smoking, race" = color("bright")(4)[3],
        "age, sex, race, education" = color("bright")(4)[4]
      ),
      labels = c(
        "age, smoking, race, education" = "Sex",
        "age, sex, smoking, education" = "Race",
        "age, sex, smoking, race" = "Education",
        "age, sex, race, education" = "Smoking"
      )
    ),
    covariates_overall = build_color_scale(
      colors = c(
        "Sex" = color("bright")(4)[1],
        "Smoking" = color("bright")(4)[4],
        "Race" = color("bright")(4)[2],
        "Education" = color("bright")(4)[3]
      )
    ),
    sex = build_color_scale(
      colors = c(
        "1" = color("sunset")(2)[1],
        "2" = color("sunset")(2)[2]
      ),
      fill = c(
        "1" = color("sunset")(2)[1],
        "2" = color("sunset")(2)[2]
      ),
      labels = c(
        "1" = "Male",
        "2" = "Female"
      )
    ),
    smoking = build_color_scale(
      colors = c(
        "1" = color("PRGn")(2)[1],
        "2" = color("PRGn")(2)[2]
      ),
      fill = c(
        "1" = color("PRGn")(2)[1],
        "2" = color("PRGn")(2)[2]
      ),
      labels = c(
        "1" = "Smokers",
        "2" = "Non-smokers"
      )
    ),
    race = build_color_scale(
      colors = c(
        "1" = color("vibrant")(4)[1],
        "3" = color("vibrant")(4)[2],
        "4" = color("vibrant")(4)[3],
        "5" = color("vibrant")(4)[4]
      ),
      fill = c(
        "1" = color("vibrant")(4)[1],
        "3" = color("vibrant")(4)[2],
        "4" = color("vibrant")(4)[3],
        "5" = color("vibrant")(4)[4]
      ),
      labels = c(
        "1" = "Hispanic",
        "3" = "White",
        "4" = "Black",
        "5" = "Other"
      )
    ),
    education = build_color_scale(
      colors = c(
        "1" = color("nightfall")(5)[1],
        "2" = color("nightfall")(5)[2],
        "3" = color("nightfall")(5)[3],
        "4" = color("nightfall")(5)[4],
        "5" = color("nightfall")(5)[5]
      ),
      fill = c(
        "1" = color("nightfall")(5)[1],
        "2" = color("nightfall")(5)[2],
        "3" = color("nightfall")(5)[3],
        "4" = color("nightfall")(5)[4],
        "5" = color("nightfall")(5)[5]
      ),
      labels = c(
        "1" = "< 9th grade",
        "2" = "9-11th grade",
        "3" = "High school",
        "4" = "Some college or AA degree",
        "5" = "College graduate or higher"
      )
    ),
    biomarker_diff = build_color_scale(
      colors = c(
        "age, lncrp, phenoage" = "#CB2027",
        "age, lncrp" = "#059748",
        "age, phenoage" = "grey65"
      ),
      fill_colors = c(
        "age, lncrp, phenoage" = "#F07E6E",
        "age, lncrp" = "#90CD97",
        "age, phenoage" = "grey80"
      ),
      labels = c(
        "age, lncrp, phenoage" = "Age + CRP + PhenoAge",
        "age, lncrp" = "Age + CRP",
        "age, phenoage" = "Age + PhenoAge"
      ),
      breaks = c(
        "age, lncrp, phenoage",
        "age, lncrp",
        "age, phenoage"
      )
    ),
    best_biomarkers = build_color_scale(
      colors = c(
        "age, lncrp" = "#256DAB"
      ),
      fill_colors = c(
        "age, lncrp" = "#88BDE6"
      )
    )
  )
}
```

#### Specify geom style

Note that using `update_geom_defaults()` is global and requires `reset_geom_defaults()` to be reset.
`update_geom_defaults()` is only set once when switching to a different plot type and subsequent
plots with the same styling will not have repeated `update_geom_defaults()` calls to avoid overhead
and clutter. Plots must therefore be run in order.

```{r fun_plot_theme}
set_geom_theme <- function(geom_theme_name, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  switch(geom_theme_name,
    ribbon = list(
      alpha = 0.1
    ),
    highlight_rect = list(
      alpha = 0.125,
      fill = "#ff0000"
    ),
    hline = list(
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    ),
    age_rate_point = list(
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ),
    age_rate_line = list(
      linewidth = 0.6,
      alpha = 1
    ),
    mse_point = list(
      size = 1.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    ),
    mse_line = list(
      linewidth = 0.7,
      alpha = 1
    ),
    biomarker_linerange = list(
      linewidth = 0.75
    ),
    biomarker_point = list(
      size = 2.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    )
  )
}
```

#### Set default geoms

The geom themes below are used in all plots. Plot-specific geoms are called on the first plot in a
sequence of plots that share the same geom theme.

```{r set_default_geoms}
update_geom_defaults("ribbon", set_geom_theme("ribbon"))
update_geom_defaults("rect", set_geom_theme("highlight_rect"))
update_geom_defaults("hline", set_geom_theme("hline"))
```

### Rate of aging

#### Main panel (paper)

```{r plot_aging, warning=FALSE}
update_geom_defaults("point", set_geom_theme("age_rate_point"))
update_geom_defaults("line", set_geom_theme("age_rate_line"))

p0 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_line(
    data = pred_age_rate$linear,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$linear,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.x = element_blank()
  )

p1 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_line(
    data = pred_age_rate$quadratic,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$quadratic,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p2 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_ribbon(
    data = pred_age_rate$smooth,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  geom_line(
    data = pred_age_rate$smooth,
    aes(
      x = age,
      y = estimate
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p3 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = pred_age_rate$linear_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$linear_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p4 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = pred_age_rate$quadratic_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$quadratic_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank()
  )

p5 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  geom_line(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      y = derivative
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank()
  )

p6 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower_adj,
      ymax = ci_upper_adj
    )
  ) +
  geom_line(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      y = derivative_adj
    )
  ) +
  axes(
    x = "age",
    y = "hl_diff"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 1, 0, unit = "pt"))


p <- align_plots(p0, p1, p2, p3, p4, p5, p6, align = "hv")

p <- plot_grid(
  plot_grid(
    p[[1]], NULL,
    p[[2]], NULL,
    p[[3]],
    labels = c("A", "", "B", "", "C"),
    nrow = 1,
    rel_widths = c(1, -0.01, 1, -0.01, 1),
    label_x = 0.04,
    label_y = 1.02
  ),
  NULL,
  plot_grid(
    p[[4]], NULL,
    p[[5]], NULL,
    p[[6]],
    labels = c("D", "", "E", "", "F"),
    nrow = 1,
    rel_widths = c(1, -0.01, 1, -0.01, 1),
    label_x = 0.04,
    label_y = 1.02
  ),
  NULL,
  plot_grid(
    NULL, p[[7]], NULL,
    nrow = 1,
    rel_widths = c(0.25, 0.5, 0.25),
    labels = c("", "G", ""),
    label_x = 0.04,
    label_y = 1.02
  ),
  ncol = 1,
  rel_heights = c(1, 0, 1, 0, 1)
)

sameplot(
  p,
  file = here(plot_config("figure_dir"), "age_rate_panel.png"),
  width = params$double_column_width,
  height = 6,
  res = 600,
  save = params$save_figures
)
```

#### Additional plots

##### Plot aging mean without covariates

```{r plot_aging_mean}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_age_rate$smooth,
      aes(
        x = age,
        y = estimate
      )
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]]
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_age_rate$smooth,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper
      )
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    plot_theme()

  sameplot(
    p0,
    file = here(plot_config("figure_dir"), "Extras", "age_rate_mean.png"),
    width = params$single_column_width,
    height = 3,
    res = 600,
    save = params$save_figures
  )
}
```

##### Plot aging GAM with covariates with raw means (supplementary)

```{r plot_covariate_model_fits_means}
if (isFALSE(params$telolength)) {
  p1 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = sex,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_sex,
      aes(
        x = age,
        y = estimate,
        color = sex
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ sex * age, design, svymean),
      aes(
        x = as.numeric(age),
        y = .data[[params$hearing_var]],
        color = as.factor(sex)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_sex,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = sex
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("sex") +
    plot_theme() +
    theme(
      axis.title.x = element_blank()
    )

  p2 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = smoking,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_smoking,
      aes(
        x = age,
        y = estimate,
        color = smoking
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ smoking * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(smoking)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_smoking,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = smoking
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("smoking") +
    plot_theme() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )

  p3 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = race,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_race,
      aes(
        x = age,
        y = estimate,
        color = race
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ race * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(race)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_race,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = race
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("race") +
    plot_theme()

  p4 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = education,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_education,
      aes(
        x = age,
        y = estimate,
        color = education
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ education * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(education)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_education,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = education
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("education") +
    plot_theme() +
    theme(
      axis.title.y = element_blank()
    )

  p <- align_plots(p1, p2, p3, p4, align = "hv")

  p <- plot_grid(
    plot_grid(
      p[[1]], NULL,
      p[[2]],
      labels = c("A", "", "B"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1),
      label_x = 0.02,
      label_y = 1.02
    ),
    NULL,
    plot_grid(
      p[[3]], NULL,
      p[[4]],
      labels = c("C", "", "D"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1),
      label_x = 0.02,
      label_y = 1.02
    ),
    ncol = 1,
    rel_heights = c(1, -0.1, 1)
  )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "Extras", "covariate_strata_panel.png"),
    width = params$double_column_width,
    height = 4,
    res = 600,
    save = params$save_figures
  )
}
```

### Bioage

#### Bioage RMSE plot (paper)

```{r plot_bioage_rmse}
update_geom_defaults("point", set_geom_theme("mse_point"))
update_geom_defaults("line", set_geom_theme("mse_line"))
p <- summary_bioage %>%
  {
    if (isTRUE(params$telolength)) filter(., !predictors %in% c("age, phenoage", "age, kdm", "age, hd_log")) else .
  } %>%
  mutate(predictors = factor(predictors, levels = c("age, phenoage", "age, kdm", "age, hd_log", "age, telolength_log", "age"))) %>%
  ggplot(
    aes(
      x = age_group, y = delta_rmse_w
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_rmse,
      ymax = ci_high_delta_rmse
    ),
    alpha = 0.1,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(signif != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_rmse + 0.01,
      label = signif,
      color = predictors
    ),
    size = plot_config("signif_size") / (14 / 5)
  ) +
  axes(
    x = "age_group",
    y = "drmse_pred"
  ) +
  color_scales("bioage") +
  plot_theme()

sameplot(
  p,
  file = here(plot_config("figure_dir"), "bioage_panel.png"),
  width = params$single_column_width,
  height = 2.15,
  res = 600,
  save = params$save_figures
)
```

#### Additional Plots

##### Advance bioage plot (supplementary)

```{r plot_bioage_advance}
p1 <- df %>%
  ggplot() +
  geom_density_ridges(
    aes(
      x = phenoage_advance,
      y = age_group
    )
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkred"
  ) +
  axes(
    x = "age_accel_phenoage",
    y = "age_group"
  ) +
  color_scales("bioage") +
  plot_theme()

p2 <- df %>%
  ggplot() +
  geom_density_ridges(
    aes(
      x = kdm_advance,
      y = age_group
    )
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkred"
  ) +
  axes(
    x = "age_accel_kdm",
    y = "age_group"
  ) +
  color_scales("bioage") +
  plot_theme()

p <- ggarrange(
  p1,
  NULL,
  p2,
  nrow = 1,
  widths = c(1, 0.1, 1),
  labels = c("A", "", "B"),
  label.x = 0.02,
  label.y = 1.01,
  align = "v"
)

sameplot(
  p,
  file = here(plot_config("figure_dir"), "Extras", "age_acceleration.png"),
  width = params$double_column_width,
  height = 2.4,
  res = 600,
  save = params$save_figures
)
```

##### covariates predictive capacity plot

```{r plot_covariates, warning=FALSE}
if (isFALSE(params$telolength)) {
  p <- summary_covariates %>%
    filter(!predictors %in% c("age, sex, smoking, race, education", "age, sex, smoking, race, education, phenoage")) %>%
    ggplot(
      aes(
        x = age_group,
        y = delta_rmse_w
      )
    ) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        color = predictors,
        group = predictors
      ),
      show.legend = TRUE
    ) +
    geom_point(
      aes(
        color = predictors
      ),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = age_group,
        y = ci_high_delta_rmse,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5)
    ) +
    axes(
      x = "age_group",
      y = "drmse_conf"
    ) +
    color_scales("covariates") +
    plot_theme()

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "covariate_panel.png"),
    width = params$single_column_width,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```

##### covariates predictive capacity overall plot

```{r plot_covariates_overall, warning=FALSE}
if (isFALSE(params$telolength)) {
  p <- summary_covariates_overall %>%
    filter(!predictors %in% c("age, sex, smoking, race, education", "age, sex, smoking, race, education, phenoage")) %>%
    mutate(predictors = factor(
      predictors,
      levels = c(
        "age, smoking, race, education",
        "age, sex, smoking, education",
        "age, sex, smoking, race",
        "age, sex, race, education"
      ),
      labels = c(
        "Sex",
        "Race",
        "Education",
        "Smoking"
      )
    )) %>%
    ggplot(
      aes(
        x = predictors,
        y = delta_rmse_w
      )
    ) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        color = predictors,
        group = predictors
      ),
      show.legend = FALSE
    ) +
    geom_linerange(
      aes(
        color = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      position = position_dodge(0.8),
      lineend = "round",
      show.legend = FALSE
    ) +
    geom_point(
      aes(
        color = predictors
      ),
      show.legend = FALSE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = predictors,
        y = ci_high_delta_rmse,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5),
      show.legend = FALSE
    ) +
    axes(
      y = "drmse_conf"
    ) +
    color_scales("covariates_overall") +
    plot_theme() +
    theme(
      axis.title.x = element_blank()
    )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "covariate_panel_overall.png"),
    width = params$single_column_width,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```

##### Biomarker phenoage variance plots

###### Plot Delta mse Biomarkers (only relevant ones)

```{r plot_biomarkers, include=FALSE}
if (isFALSE(params$telolength) && isTRUE(params$run_biomarker_add_analysis)) {
  p1 <- summary_biomarkers %>%
    filter(predictors %in% c("age, lncrp")) %>%
    mutate(
      label = predictors,
      label = case_when(
        label == "age, lncrp" ~ "Age + CRP",
        TRUE ~ NA
      ),
      predictors = factor(
        predictors,
        levels = c("age, lncrp")
      )
    ) %>%
    ggplot(
      aes(
        x = age_group,
        y = delta_rmse_w
      )
    ) +
    geom_hline(
      yintercept = 0
    ) +
    geom_label(
      data = . %>%
        filter(age_group == "20-30"),
      aes(
        x = "20-30",
        y = 3.65,
        label = label,
        fill = predictors
      ),
      size = 8 / (14 / 5),
      hjust = 0,
      label.padding = unit(0.3, "lines"),
      nudge_x = -0.1,
      color = "white",
      show.legend = FALSE
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        color = predictors,
        group = predictors
      ),
      show.legend = FALSE
    ) +
    geom_point(
      aes(
        color = predictors
      ),
      show.legend = FALSE
    ) +
    facet_wrap(
      ~predictors,
      ncol = 1,
      nrow = 5,
      labeller = labeller(
        predictors = c(
          "age, lncrp" = "Age + CRP"
        )
      )
    ) +
    axes(
      x = "age_group",
      y = "drmse_pred"
    ) +
    color_scales("best_biomarkers") +
    plot_theme() +
    theme(strip.text.x = element_blank())

  sameplot(
    p1,
    file = here(plot_config("figure_dir"), "biomarkers_biomarkers_panel.png"),
    width = 2.5,
    height = 3,
    res = 600,
    save = FALSE
  )
}
```

###### Plot biomarkers together vs biomarkers + phenoage

```{r plot_explained_var, include=FALSE}
if (isFALSE(params$telolength) && isTRUE(params$run_biomarker_add_analysis)) {
  p2 <- summary_biomarkers %>%
    filter(
      predictors %in% c(
        "age, phenoage",
        "age, lncrp",
        "age, lncrp, phenoage"
      )
    ) %>%
    mutate(predictors = factor(
      predictors,
      levels = c(
        "age, phenoage",
        "age, lncrp",
        "age, lncrp, phenoage"
      )
    )) %>%
    ggplot(
      aes(
        x = age_group,
        y = delta_rmse_w
      )
    ) +
    geom_hline(
      yintercept = 0,
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        color = predictors,
        group = predictors
      ),
      show.legend = TRUE
    ) +
    geom_point(
      aes(
        color = predictors
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age_group",
      y = "drmse_pred"
    ) +
    color_scales("biomarker_diff") +
    plot_theme() +
    theme(
      axis.title.y = element_blank(),
      plot.margin = margin(1, 1, 0, 0, unit = "pt")
    )

  sameplot(
    p2,
    file = here(plot_config("figure_dir"), "biomarkers_v_phenoage.png"),
    width = 3,
    height = 4,
    res = 600,
    save = FALSE
  )
}
```

###### Composite plot

```{r plot_biomark_panel}
if (isFALSE(params$telolength) && isTRUE(params$run_biomarker_add_analysis)) {
  p <- ggarrange(
    p1,
    NULL,
    p2,
    nrow = 1,
    widths = c(0.6, 0, 1),
    labels = c("A", "", "B"),
    label.x = 0.02,
    label.y = 1.01,
    align = "v"
  )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "biomarkers_fig_panel.png"),
    width = params$double_column_width,
    height = 3,
    res = 600,
    save = params$save_figures
  )
}
```

##### Biomarkers supplemental plots

```{r plot_biomarker_all, warning=FALSE}
if (isFALSE(params$telolength)) {
  update_geom_defaults("linerange", set_geom_theme("biomarker_linerange"))
  update_geom_defaults("point", set_geom_theme("biomarker_point"))

  p <- summary_biomarkers_overall %>%
    filter(!(
      predictors %in% c(
        "age",
        "age, phenoage",
        "age, lncrp, phenoage"
      )
    )) %>%
    mutate(predictors = factor(
      predictors,
      levels = rev(biomarker_labels$predictors),
      labels = rev(biomarker_labels$label)
    )) %>%
    ggplot(
      aes(
        x = predictors,
        y = delta_rmse_w
      )
    ) +
    geom_point(
      aes(
        color = predictors,
        fill = predictors
      ),
      position = position_dodge(0.8),
      show.legend = FALSE
    ) +
    geom_linerange(
      aes(
        color = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      position = position_dodge(0.8),
      lineend = "round",
      show.legend = FALSE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = predictors,
        y = ci_high_delta_rmse + 0.01,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5),
      show.legend = FALSE
    ) +
    axes(
      y = "drmse_biomark"
    ) +
    plot_theme() +
    theme(
      panel.grid.major.y = element_line(color = "grey95", linewidth = 0.25),
      axis.text.x = element_markdown(
        "sans",
        size = 8,
        angle = 45,
        hjust = 1,
        vjust = 1
      ),
      axis.title.x = element_blank()
    )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "biomarkers_panel.png"),
    width = params$double_column_width,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```
