---
editor_options:
  markdown:
    wrap: 100
---

# Visualization

This section contains the main results and figures. Note that some supplementary figures are in
02_data_preprocessing or 03_modeling.

## Tables

### Cross-validation tables

#### Store biomarker labels

```{r biomarker_labels}
if (isFALSE(params$telolength)) {
  biomarker_labels <- summary_biomarkers_overall %>%
    mutate(
      label =
        c(
          "Age",
          "GLU",
          "ALB",
          "ALP",
          "HbA1c",
          "CREA",
          "BASO",
          "Ca",
          "Cl",
          "HDL-C",
          "BUN",
          "Hb",
          "HCT",
          "Fe",
          "LDH",
          "LYMPH",
          "MCH",
          "MCHC",
          "Mono",
          "NEUT",
          "P",
          "K",
          "RBC",
          "RDW",
          "MCV",
          "Na",
          "TC",
          "TP",
          "UA",
          "WBC",
          "SBP",
          "ALT",
          "AST",
          "GGT",
          "CRP"
        )
    ) %>%
    arrange(desc(delta_rmse_w)) %>%
    select(predictors, label)
}
```

#### Merge tables

```{r merge_tables}
merge_tables <- function(per_age_table, overall_table) {
  per_age_table %>%
    bind_rows(
      overall_table %>% mutate(
        age_group = "Overall"
      )
    ) %>%
    mutate(
      age_group = factor(age_group, levels = c("Overall", unique(per_age_table$age_group)))
    ) %>%
    arrange(predictors, age_group)
}
```

#### Order, filter, and rename tables

```{r order_table}
rename_bioage_table <- function(bioage_table) {
  bioage_table <- bioage_table %>%
    mutate(
      predictors = case_when(
        predictors == "age" ~ "Age",
        predictors == "age, hd_log" ~ "Age + HD",
        predictors == "age, kdm" ~ "Age + KDM",
        predictors == "age, phenoage" ~ "Age + PhenoAge",
        isTRUE(params$telolength) & predictors == "age, telolength_log" ~ "Age + Telomere Length",
        TRUE ~ predictors
      ),
      predictors = factor(
        predictors,
        levels = if (isTRUE(params$telolength)) {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge", "Age + Telomere Length")
        } else {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge")
        }
      )
    ) %>%
    arrange(predictors)

  bioage_table
}

rename_bioage_reverse_table <- function(bioage_table) {
  bioage_table <- bioage_table %>%
    mutate(
      predictors = case_when(
        predictors == "phenoage" ~ "PhenoAge",
        predictors == "kdm" ~ "KDM",
        predictors == "hd_log" ~ "HD",
        predictors == "phenoage, age" ~ "PhenoAge + Age",
        predictors == "kdm, age" ~ "KDM + Age",
        predictors == "hd_log, age" ~ "HD + Age",
        TRUE ~ predictors
      ),
      predictors = factor(
        predictors,
        levels =
          c("PhenoAge", "PhenoAge + Age", "KDM", "KDM + Age", "HD", "HD + Age")
      )
    ) %>%
    arrange(predictors)

  bioage_table
}

rename_biomarker_table <- function(biomarker_table) {
  biomarker_table <- biomarker_table %>%
    filter(
      !(predictors %in% c(
        "age"
      ))
    ) %>%
    mutate(
      predictors = factor(predictors, levels = biomarker_labels$predictors, labels = paste0("Age + ", biomarker_labels$label))
    ) %>%
    arrange(predictors)

  biomarker_table
}

rename_covariate_table <- function(covariates_table) {
  covariates_table <- covariates_table %>%
    filter(
      predictors != "age, sex, smoking, race, education"
    ) %>%
    mutate(
      predictors = case_when(
        predictors == "age, smoking, race, education" ~ "Sex",
        predictors == "age, sex, race, education" ~ "Smoking",
        predictors == "age, sex, smoking, education" ~ "Race",
        predictors == "age, sex, smoking, race" ~ "Education"
      ),
      predictors = factor(predictors, levels = c(
        "Sex",
        "Smoking",
        "Race",
        "Education"
      ))
    ) %>%
    arrange(predictors)

  covariates_table
}
```

#### Format table

Round R² and delta RMSE, rename variables, set p values to 'below 0.001' when very small.

```{r format_table}
format_table <- function(summary_data) {
  table <- summary_data %>%
    mutate(
      r2 = round(r2_w, 3),
      delta_rmse = round(delta_rmse_w, 2),
      r2p = round(r2p, 2),
      p_adj = round(p_adj, 3)
    ) %>%
    mutate(
      p_adj = as.character(p_adj),
      p_adj = case_when(
        signif == "***" ~ cell_spec("< 0.001", "html", bold = TRUE),
        suppressWarnings(as.numeric(p_adj) < 0.05) ~ cell_spec(p_adj, "html", bold = TRUE),
        is.na(p_adj) ~ "",
        TRUE ~ p_adj
      ),
      delta_rmse = as.character(delta_rmse),
      delta_rmse = case_when(
        predictors %in% c("Age", "PhenoAge", "KDM", "HD") ~ "",
        TRUE ~ delta_rmse
      ),
      r2p = as.character(r2p),
      r2p = case_when(
        predictors %in% c("Age", "PhenoAge", "KDM", "HD") ~ "",
        TRUE ~ r2p
      )
    ) %>%
    rename(
      "Age" = age_group,
      "Predictors" = predictors,
      "∆RMSE (dB HL)*" = delta_rmse,
      "<i>n</i>" = n,
      "<i>R</i>²<i><sub>CV</sub></i>" = r2,
      "<i>R</i>²<i><sub>p,CV</sub></i>*" = r2p,
      "<i>p</i> value*" = p_adj
    ) %>%
    select(Predictors, Age, `<i>n</i>`, `<i>R</i>²<i><sub>CV</sub></i>`, `<i>R</i>²<i><sub>p,CV</sub></i>*`, `∆RMSE (dB HL)*`, `<i>p</i> value*`)
}
```

#### Run table formatting pipeline

```{r table_format_pipe}
if (isTRUE(params$telolength)) {
  table_bioage <- summary_bioage_overall %>%
    mutate(age_group = "Overall") %>%
    filter(predictors %in% c("age", "age, telolength_log")) %>%
    rename_bioage_table() %>%
    format_table()
} else {
  table_bioage <- merge_tables(summary_bioage, summary_bioage_overall) %>%
    rename_bioage_table() %>%
    format_table()

  table_bioage_reverse <- merge_tables(summary_bioage_reverse, summary_bioage_reverse_overall) %>%
    rename_bioage_reverse_table() %>%
    format_table()

  table_biomarkers <- merge_tables(summary_biomarkers, summary_biomarkers_overall) %>%
    rename_biomarker_table() %>%
    format_table() %>%
    filter(
      Age == "Overall"
    )

  table_covariates <- merge_tables(summary_covariates, summary_covariates_overall) %>%
    rename_covariate_table() %>%
    format_table()
}
```

#### Render table

```{r store_tables}
render_and_save_table <- function(table, path, covariate = FALSE) {
  column_names <- names(table)

  if (isTRUE(covariate)) {
    column_names[1:2] <- c("&nbsp;", "&nbsp;")
  }

  html_table <- table %>%
    kable(
      format = "html",
      col.names = column_names,
      align = rep("c", ncol(table)),
      escape = FALSE,
      table.attr = paste0(
        'border="1" cellspacing="0" cellpadding="0" ',
        'style="width:159mm; table-layout:fixed; border-collapse:collapse; ',
        "font-family:'Times New Roman', Times, serif; ",
        'font-size:10pt; color:black;"'
      )
    ) %>%
    column_spec(
      1,
      bold = TRUE
    ) %>%
    row_spec(
      0,
      color = "black",
      extra_css = "height:5.29mm; padding:0; vertical-align:middle;"
    ) %>%
    row_spec(
      1:nrow(table),
      extra_css = "height:5.29mm; padding:0; vertical-align:middle;"
    ) %>%
    collapse_rows(
      columns = 1,
      valign = "middle"
    )

  if (isTRUE(params$save_figures)) {
    save_kable(html_table, path)
  }

  html_table
}

render_and_save_table(
  table = table_bioage,
  path = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "table_bioage.html"
  )
)

if (isFALSE(params$telolength)) {
  render_and_save_table(
    table = table_biomarkers,
    path = here("output", "figures", params$hearing_var, "table_biomarkers.html")
  ) %>%
    scroll_box(width = "100%", height = "400px")

  render_and_save_table(
    table = table_bioage_reverse,
    path = here("output", "figures", params$hearing_var, "table_bioage_reverse.html")
  ) %>%
    scroll_box(width = "100%", height = "400px")

  render_and_save_table(
    table = table_covariates,
    path = here("output", "figures", params$hearing_var, "table_covariates.html")
  ) %>%
    scroll_box(width = "100%", height = "400px")
}
```

### Covariates table

#### Create covariate table and format covariate table

We recode variables, combine means and SDs, pivot table, and format it

```{r pivot_conf_table}
if (isFALSE(params$telolength)) {
  table_covariates_descriptive <- descriptives %>%
    filter(
      age_group != "Overall"
    ) %>%
    mutate(
      stat = sprintf("%.1f (%.1f)", .data[[paste0("mean_w_", params$hearing_var)]], .data[[paste0("sd_w_", params$hearing_var)]]),
      category = case_when(
        category == "Overall" ~ "All",
        TRUE ~ category
      ),
      category = factor(category, levels = c(
        "All",
        "Male",
        "Female",
        "White",
        "Hispanic",
        "Black",
        "Other",
        "Smoker",
        "Non-smoker",
        "< 9th grade",
        "9-11th grade",
        "High school",
        "Some college or AA degree",
        "College graduate or higher"
      ))
    ) %>%
    select(
      age_group, group, category, stat
    ) %>%
    pivot_wider(names_from = age_group, values_from = stat) %>%
    as.data.frame()
}
```

#### Save table

```{r store_conf_table}
if (isFALSE(params$telolength)) {
  render_and_save_table(
    table_covariates_descriptive,
    covariate = TRUE,
    path = here(
      "output",
      "figures",
      if (isTRUE(params$telolength)) {
        paste0(params$hearing_var, "_telolength_log")
      } else {
        params$hearing_var
      },
      "table_covariates_descriptives.html"
    )
  )
}
```

## Descriptive tables

### Show descriptives tables

```{r show_descr_table}
descriptives_unfiltered %>%
  construct_table(
    title = "Descriptive statistics (before filtering)"
  )

descriptives %>%
  construct_table(
    title = "Descriptive statistics (after filtering)"
  )
```

### Show summary descriptives table

```{r summary_descr_table}
descriptive_summary_table <- bind_rows(
  descriptives %>%
    filter(age_group == "Overall", group == "Sex") %>%
    arrange(match(category, c("Male", "Female"))) %>%
    transmute(
      Characteristic = as.character(category),
      Value = sprintf(
        "%.1f%% (<i>n</i> = %s)",
        100 * proportion_w,
        format(n, big.mark = ",")
      )
    ) %>%
    add_row(Characteristic = "<b>Sex</b>", Value = "", .before = 1),
  descriptives %>%
    filter(group == "Overall", category == "Overall") %>%
    transmute(
      Characteristic = if_else(
        age_group == "Overall",
        "<b>Chronological Age (years)</b>",
        as.character(age_group)
      ),
      Value = if_else(
        age_group == "Overall",
        sprintf("%.1f ± %.1f", mean_w_age, sd_w_age),
        sprintf(
          "%.1f%% (<i>n</i> = %s)",
          100 * proportion_w,
          format(n, big.mark = ",")
        )
      )
    ),
  descriptives %>%
    filter(age_group == "Overall", group == "Race") %>%
    arrange(match(category, c("White", "Black", "Hispanic", "Other"))) %>%
    transmute(
      Characteristic = recode(
        as.character(category),
        White = "White, non-Hispanic",
        Black = "Black, non-Hispanic"
      ),
      Value = sprintf(
        "%.1f%% (<i>n</i> = %s)",
        100 * proportion_w,
        format(n, big.mark = ",")
      )
    ) %>%
    add_row(
      Characteristic = "<b>Race/Ethnicity</b>",
      Value = "",
      .before = 1
    ),
  descriptives %>%
    filter(
      age_group == "Overall",
      group == "Overall",
      category == "Overall"
    ) %>%
    transmute(
      Characteristic = list(sprintf("<b>%s</b>", c(
        "HF-PTA (dB HL)",
        "LF-PTA (dB HL)",
        "KDM",
        "PhenoAge",
        "HD (log)",
        "Telomere Mean Length (log)"
      ))),
      Value = list(c(
        sprintf("%.1f ± %.1f", mean_w_pta_hf, sd_w_pta_hf),
        sprintf("%.1f ± %.1f", mean_w_pta_lf, sd_w_pta_lf),
        sprintf("%.1f ± %.1f", mean_w_kdm, sd_w_kdm),
        sprintf("%.1f ± %.1f", mean_w_phenoage, sd_w_phenoage),
        sprintf("%.1f ± %.1f", mean_w_hd_log, sd_w_hd_log),
        sprintf(
          "%.2f ± %.2f (<i>n</i> = %s)",
          mean_w_telolength_log,
          sd_w_telolength_log,
          format(
            design$variables %>%
              filter(!is.na(telolength_log)) %>%
              nrow(),
            big.mark = ","
          )
        )
      ))
    ) %>%
    unnest(c(Characteristic, Value))
) %>%
  kable(
    format = "html",
    col.names = c(
      "Characteristic",
      "Percentage or<br>Mean ± Standard Deviation"
    ),
    align = c("c", "c"),
    escape = FALSE,
    table.attr = paste0(
      'cellspacing="0" cellpadding="0" ',
      'style="width:159mm; table-layout:fixed; border-collapse:collapse; ',
      "border:1px solid black; margin-left:auto; margin-right:auto; ",
      "font-family:'Times New Roman', Times, serif; ",
      'font-size:10pt; color:black;"'
    )
  ) %>%
  row_spec(
    0,
    extra_css = paste0(
      "border-bottom:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  row_spec(
    if (isFALSE(params$telolength)) c(3, 10, 15:21) else c(3, 9, 14:20),
    extra_css = paste0(
      "border-bottom:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  column_spec(
    1,
    include_thead = TRUE,
    extra_css = paste0(
      "border-right:1px solid black;",
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  ) %>%
  column_spec(
    2,
    extra_css = paste0(
      "padding:0;",
      "font-family:'Times New Roman', Times, serif;",
      "font-size:10pt;",
      "color:black;"
    )
  )

if (isTRUE(params$save_figures)) {
  save_kable(descriptive_summary_table, here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "descriptive_summary.html"
  ))
}

descriptive_summary_table
```

## Figures

### Set plot config

#### Specify plot styling variables

We store a separate configuration depending on output type of figure for easy switching between
output types. Only the currently selected output type variables in params + global variables for
plotting we may need later is returned. These variables are used to construct ggplot protoobjects
and can be accessed either with `plot_config()`, which returns all, or with
`plot_config("parameter")` to return a specific parameter.

```{r fun_plot_cfg}
plot_config <- function(variable = NULL, global_params = params) {
  p_cfg_vars <- list(
    global = list(
      axis_unit_color = "#4D4D4D" # used later when constructing axes from HTML
    ),
    paper = list(
      figure_dir = here(
        "output",
        "figures",
        if (isTRUE(global_params$telolength)) {
          paste0(global_params$hearing_var, "_telolength_log")
        } else {
          global_params$hearing_var
        }
      ),
      text_size_primary = 8,
      text_size_secondary = 9,
      signif_size = 14
    )
  )

  p_cfg_vars_comb <- c(p_cfg_vars$global, p_cfg_vars[[global_params$plot_output_type]])

  if (!is.null(variable)) p_cfg_vars_comb[[variable]] else p_cfg_vars_comb
}
```

#### Specify general plot theme

We first store the theme preset (theme_few), then theme overrides, then other theme items like
guides. All items specified in `plot_theme()` must be ggplot proto objects.

```{r fun_plot_theme}
plot_theme <- function(global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  list(
    theme_few(),
    theme(
      axis.text.x = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      axis.text.y = element_markdown(
        "sans",
        size = config$text_size_primary
      ),
      strip.text.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      strip.text.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = config$text_size_secondary,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ),
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = config$text_size_secondary,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
  )
}
```

#### Specify axis scales

```{r fun_plot_axes}
axes <- function(x = NULL, y = NULL, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  if (!is.null(x)) {
    x <- switch(x,
      age = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(20, 80, 10),
        limits = c(20, 80),
        name = "Age",
        unit = "Years"
      ),
      age_group = define_axis(
        config = config,
        axis = "x",
        type = "discrete",
        limits = c(1.3, 5.7),
        name = "Age",
        unit = "Years"
      ),
      age_accel_phenoage = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-25, 25),
        name = "PhenoAge Age Acceleration",
        unit = "PhenoAge – Age (Years)"
      ),
      age_accel_kdm = define_axis(
        config = config,
        axis = "x",
        type = "continuous",
        breaks = seq(-20, 20, 20),
        limits = c(-25, 25),
        name = "KDM Age Acceleration",
        unit = "KDM – Age (Years)"
      )
    )
  }

  if (!is.null(y)) {
    y <- switch(y,
      age_group = define_axis(
        config = config,
        axis = "y",
        type = "discrete",
        limits = c(1, 7.3),
        name = "Age",
        unit = "Years"
      ),
      hl = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 90, 30),
        limits = c(-10, 100),
        name = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA"),
        unit = "dB HL"
      ),
      hl_rate = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(-1, 2, 1),
        limits = c(-0.3, 2),
        name = "Rate of HL",
        unit = "dB HL/Year"
      ),
      hl_diff = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(-1, 2, 1),
        limits = c(-1.25, 2),
        name = "Deviation Rate of HL",
        unit = "dB HL/Year"
      ),
      drmse_pred = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 1.5, 0.25),
        limits = c(0, 1.545),
        name = "Additional Predictive Performance",
        unit = "&Delta;RMSE<span style='color:transparent;'>xi</span>(dB HL)"
      ),
      drmse_reverse_pred = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 15, 2.5),
        limits = c(0, 15.45),
        name = "Additional Predictive Performance",
        unit = "ΔRMSE (dB HL)"
      ),
      drmse_biomark = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 0.1, 0.05),
        limits = c(-0.025, 0.1015),
        name = "Additional Predictive Performance",
        unit = "&Delta;RMSE<span style='color:transparent;'>xi</span>(dB HL)"
      ),
      drmse_conf = define_axis(
        config = config,
        axis = "y",
        type = "continuous",
        breaks = seq(0, 2, 0.5),
        limits = c(-0.05, 2.3),
        name = "Additional Predictive Performance",
        unit = "ΔRMSE (dB HL)"
      )
    )
  }

  build_axes(x = x, y = y)
}
```

#### Specify colors

```{r fun_plot_colors}
color_scales <- function(color_list_name) {
  switch(color_list_name,
    bioage = build_color_scale(
      colors = c(
        "age" = "#000000",
        "age, hd_log" = if (isFALSE(params$telolength)) color("highcontrast")(3)[1] else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) color("highcontrast")(3)[2] else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) color("highcontrast")(3)[3] else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "#7B3A96" else NULL
      ),
      fill_colors = c(
        "age" = "#000000",
        "age, hd_log" = if (isFALSE(params$telolength)) color("highcontrast")(3)[1] else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) color("highcontrast")(3)[2] else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) color("highcontrast")(3)[3] else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "#B276B2" else NULL
      ),
      labels = c(
        "age" = "Age",
        "age, hd_log" = if (isFALSE(params$telolength)) "Age + HD" else NULL,
        "age, kdm" = if (isFALSE(params$telolength)) "Age + KDM" else NULL,
        "age, phenoage" = if (isFALSE(params$telolength)) "Age + PhenoAge" else NULL,
        "age, telolength_log" = if (isTRUE(params$telolength)) "Age + Telomere Length" else NULL
      )
    ),
    bioage_reverse = build_color_scale(
      colors = c(
        "hd_log, age" = color("highcontrast")(3)[1],
        "kdm, age" = color("highcontrast")(3)[2],
        "phenoage, age" = color("highcontrast")(3)[3]
      ),
      fill_colors = c(
        "hd_log, age" = color("highcontrast")(3)[1],
        "kdm, age" = color("highcontrast")(3)[2],
        "phenoage, age" = color("highcontrast")(3)[3]
      ),
      labels = c(
        "hd_log, age" = "HD + Age",
        "kdm, age" = "KDM + Age",
        "phenoage, age" = "PhenoAge + Age"
      )
    ),
    covariates = build_color_scale(
      colors = c(
        "age, smoking, race, education" = color("bright")(4)[1],
        "age, sex, smoking, education" = color("bright")(4)[2],
        "age, sex, smoking, race" = color("bright")(4)[3],
        "age, sex, race, education" = color("bright")(4)[4]
      ),
      fill_colors = c(
        "age, smoking, race, education" = color("bright")(4)[1],
        "age, sex, smoking, education" = color("bright")(4)[2],
        "age, sex, smoking, race" = color("bright")(4)[3],
        "age, sex, race, education" = color("bright")(4)[4]
      ),
      labels = c(
        "age, smoking, race, education" = "Sex",
        "age, sex, smoking, education" = "Race",
        "age, sex, smoking, race" = "Education",
        "age, sex, race, education" = "Smoking"
      )
    ),
    covariates_overall = build_color_scale(
      colors = c(
        "Sex" = color("bright")(4)[1],
        "Smoking" = color("bright")(4)[4],
        "Race" = color("bright")(4)[2],
        "Education" = color("bright")(4)[3]
      )
    ),
    sex = build_color_scale(
      colors = c(
        "1" = color("sunset")(2)[1],
        "2" = color("sunset")(2)[2]
      ),
      fill_colors = c(
        "1" = color("sunset")(2)[1],
        "2" = color("sunset")(2)[2]
      ),
      labels = c(
        "1" = "Male",
        "2" = "Female"
      )
    ),
    smoking = build_color_scale(
      colors = c(
        "1" = color("PRGn")(2)[1],
        "2" = color("PRGn")(2)[2]
      ),
      fill_colors = c(
        "1" = color("PRGn")(2)[1],
        "2" = color("PRGn")(2)[2]
      ),
      labels = c(
        "1" = "Smokers",
        "2" = "Non-smokers"
      )
    ),
    race = build_color_scale(
      colors = c(
        "1" = color("vibrant")(4)[1],
        "3" = color("vibrant")(4)[2],
        "4" = color("vibrant")(4)[3],
        "5" = color("vibrant")(4)[4]
      ),
      fill_colors = c(
        "1" = color("vibrant")(4)[1],
        "3" = color("vibrant")(4)[2],
        "4" = color("vibrant")(4)[3],
        "5" = color("vibrant")(4)[4]
      ),
      labels = c(
        "1" = "Hispanic",
        "3" = "White",
        "4" = "Black",
        "5" = "Other"
      )
    ),
    education = build_color_scale(
      colors = c(
        "1" = color("nightfall")(5)[1],
        "2" = color("nightfall")(5)[2],
        "3" = color("nightfall")(5)[3],
        "4" = color("nightfall")(5)[4],
        "5" = color("nightfall")(5)[5]
      ),
      fill_colors = c(
        "1" = color("nightfall")(5)[1],
        "2" = color("nightfall")(5)[2],
        "3" = color("nightfall")(5)[3],
        "4" = color("nightfall")(5)[4],
        "5" = color("nightfall")(5)[5]
      ),
      labels = c(
        "1" = "< 9th grade",
        "2" = "9-11th grade",
        "3" = "High school",
        "4" = "Some college or AA degree",
        "5" = "College graduate or higher"
      )
    )
  )
}
```

#### Specify geom style

Note that using `update_geom_defaults()` is global and requires `reset_geom_defaults()` to be reset.
`update_geom_defaults()` is only set once when switching to a different plot type and subsequent
plots with the same styling will not have repeated `update_geom_defaults()` calls to avoid overhead
and clutter. Plots must therefore be run in order.

```{r fun_geom_theme}
set_geom_theme <- function(geom_theme_name, global_params = params) {
  config <- if (!is.null(global_params)) plot_config(global_params = global_params)

  switch(geom_theme_name,
    ribbon = list(
      alpha = 0.1
    ),
    highlight_rect = list(
      alpha = 0.125,
      fill = "#ff0000"
    ),
    hline = list(
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    ),
    age_rate_point = list(
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ),
    age_rate_line = list(
      linewidth = 0.6,
      alpha = 1
    ),
    mse_point = list(
      size = 1.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    ),
    mse_line = list(
      linewidth = 0.7,
      alpha = 1
    ),
    biomarker_linerange = list(
      linewidth = 0.75
    ),
    biomarker_point = list(
      size = 2.25,
      alpha = 1,
      shape = 16,
      stroke = 0.5
    )
  )
}
```

#### Set default geoms

The geom themes below are used in all plots. Plot-specific geoms are called on the first plot in a
sequence of plots that share the same geom theme.

```{r set_default_geoms}
update_geom_defaults("ribbon", set_geom_theme("ribbon"))
update_geom_defaults("rect", set_geom_theme("highlight_rect"))
update_geom_defaults("hline", set_geom_theme("hline"))
```

### Rate of aging

#### Main panel (paper)

```{r plot_aging, warning=FALSE}
update_geom_defaults("point", set_geom_theme("age_rate_point"))
update_geom_defaults("line", set_geom_theme("age_rate_line"))

p0 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_line(
    data = pred_age_rate$linear,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$linear,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.x = element_blank()
  )

p1 <- ggplot() +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_line(
    data = pred_age_rate$quadratic,
    aes(
      x = age,
      y = estimate
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$quadratic,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p2 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_line(
    data = svyby(as.formula(paste0("~", params$hearing_var)), ~age, design, svymean),
    aes(
      x = age,
      y = .data[[params$hearing_var]]
    ),
    alpha = 0.4
  ) +
  geom_point(
    data = df,
    aes(
      x = age,
      y = .data[[params$hearing_var]],
      alpha = 0.05 * weights / mean(weights)
    ),
    show.legend = FALSE
  ) +
  geom_ribbon(
    data = pred_age_rate$smooth,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  geom_line(
    data = pred_age_rate$smooth,
    aes(
      x = age,
      y = estimate
    )
  ) +
  axes(
    x = "age",
    y = "hl"
  ) +
  scale_alpha_identity() +
  plot_theme() +
  theme(
    plot.margin = margin(1, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p3 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = pred_age_rate$linear_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$linear_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p4 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = pred_age_rate$quadratic_derivative,
    aes(
      x = age,
      y = derivative
    )
  ) +
  geom_ribbon(
    data = pred_age_rate$quadratic_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank()
  )

p5 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    )
  ) +
  geom_line(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      y = derivative
    )
  ) +
  axes(
    x = "age",
    y = "hl_rate"
  ) +
  plot_theme() +
  theme(
    plot.margin = margin(0, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank()
  )

p6 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      ymin = ci_lower_adj,
      ymax = ci_upper_adj
    )
  ) +
  geom_line(
    data = pred_age_rate$derivative_smooth,
    aes(
      x = age,
      y = derivative_adj
    )
  ) +
  axes(
    x = "age",
    y = "hl_diff"
  ) +
  plot_theme() +
  theme(plot.margin = margin(0, 0, 1, 0, unit = "pt"))


p <- align_plots(p0, p1, p2, p3, p4, p5, p6, align = "hv")

p <- plot_grid(
  plot_grid(
    p[[1]], NULL,
    p[[2]], NULL,
    p[[3]],
    labels = c("A", "", "B", "", "C"),
    nrow = 1,
    rel_widths = c(1, -0.01, 1, -0.01, 1),
    label_x = 0.04,
    label_y = 1.02
  ),
  NULL,
  plot_grid(
    p[[4]], NULL,
    p[[5]], NULL,
    p[[6]],
    labels = c("D", "", "E", "", "F"),
    nrow = 1,
    rel_widths = c(1, -0.01, 1, -0.01, 1),
    label_x = 0.04,
    label_y = 1.02
  ),
  NULL,
  plot_grid(
    NULL, p[[7]], NULL,
    nrow = 1,
    rel_widths = c(0.25, 0.5, 0.25),
    labels = c("", "G", ""),
    label_x = 0.04,
    label_y = 1.02
  ),
  ncol = 1,
  rel_heights = c(1, 0, 1, 0, 1)
)

sameplot(
  p,
  file = here(plot_config("figure_dir"), "age_rate_panel.tif"),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures
)
```

### Covariates

#### Plot aging GAM with covariates (paper)

```{r plot_covariate_model_fits_means}
if (isFALSE(params$telolength)) {
  p1 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = sex,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_sex,
      aes(
        x = age,
        y = estimate,
        color = sex
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ sex * age, design, svymean),
      aes(
        x = as.numeric(age),
        y = .data[[params$hearing_var]],
        color = as.factor(sex)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_sex,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = sex
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("sex") +
    plot_theme() +
    theme(
      axis.title.x = element_blank()
    )

  p2 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = smoking,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_smoking,
      aes(
        x = age,
        y = estimate,
        color = smoking
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ smoking * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(smoking)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_smoking,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = smoking
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("smoking") +
    plot_theme() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )

  p3 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = race,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_race,
      aes(
        x = age,
        y = estimate,
        color = race
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ race * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(race)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_race,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = race
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("race") +
    plot_theme()

  p4 <- ggplot() +
    geom_point(
      data = df,
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = education,
        alpha = 0.05 * weights / mean(weights)
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = pred_covariates_full$smooth_education,
      aes(
        x = age,
        y = estimate,
        color = education
      ),
      show.legend = TRUE
    ) +
    geom_line(
      data = svyby(as.formula(paste0("~", params$hearing_var)), ~ education * age, design, svymean),
      aes(
        x = age,
        y = .data[[params$hearing_var]],
        color = as.factor(education)
      ),
      alpha = 0.4
    ) +
    geom_ribbon(
      data = pred_covariates_full$smooth_education,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = education
      ),
      show.legend = FALSE
    ) +
    axes(
      x = "age",
      y = "hl"
    ) +
    scale_alpha_identity() +
    color_scales("education") +
    plot_theme() +
    theme(
      axis.title.y = element_blank()
    )

  p <- align_plots(p1, p2, p3, p4, align = "hv")

  p <- plot_grid(
    plot_grid(
      p[[1]], NULL,
      p[[2]],
      labels = c("A", "", "B"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1),
      label_x = 0.02,
      label_y = 1.02
    ),
    NULL,
    plot_grid(
      p[[3]], NULL,
      p[[4]],
      labels = c("C", "", "D"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1),
      label_x = 0.02,
      label_y = 1.02
    ),
    ncol = 1,
    rel_heights = c(1, -0.1, 1)
  )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "age_rate_covariate_panel.tif"),
    width = params$double_column_width,
    height = 4,
    res = 1200,
    save = params$save_figures
  )
}
```

#### covariates predictive capacity plot (paper)

```{r plot_covariates, warning=FALSE}
update_geom_defaults("point", set_geom_theme("mse_point"))
update_geom_defaults("line", set_geom_theme("mse_line"))

if (isFALSE(params$telolength)) {
  p <- summary_covariates %>%
    filter(!predictors %in% c("age, sex, smoking, race, education")) %>%
    ggplot(
      aes(
        x = age_group,
        y = delta_rmse_w
      )
    ) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        color = predictors,
        group = predictors
      ),
      show.legend = TRUE
    ) +
    geom_point(
      aes(
        color = predictors
      ),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = age_group,
        y = ci_high_delta_rmse,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5)
    ) +
    axes(
      x = "age_group",
      y = "drmse_conf"
    ) +
    color_scales("covariates") +
    plot_theme()

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "covariate_panel.tif"),
    width = params$single_column_width,
    height = 2.35,
    res = 1200,
    save = params$save_figures
  )
}
```

### Bioage

#### Bioage delta RMSE plot (supplementary)

```{r plot_bioage_rmse}
p1 <- summary_bioage %>%
  {
    if (isTRUE(params$telolength)) filter(., !predictors %in% c("age, phenoage", "age, kdm", "age, hd_log")) else .
  } %>%
  mutate(predictors = factor(predictors, levels = c("age, phenoage", "age, kdm", "age, hd_log", "age, telolength_log", "age"))) %>%
  ggplot(
    aes(
      x = age_group, y = delta_rmse_w
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_rmse,
      ymax = ci_high_delta_rmse
    ),
    alpha = 0.1,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(signif != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_rmse + 0.01,
      label = signif,
      color = predictors
    ),
    size = plot_config("signif_size") / (14 / 5)
  ) +
  axes(
    x = "age_group",
    y = "drmse_pred"
  ) +
  color_scales("bioage") +
  plot_theme()

sameplot(
  p1,
  width = params$single_column_width,
  height = 2.35,
  res = 1200
)

if (isTRUE(params$telolength)) {
  if (params$hearing_var == "pta_lf") {
    p1 <- p1 + theme(axis.title.y = element_markdown(
      "sans",
      size = 9,
      lineheight = 1.1,
      color = "white"
    ))
  }
  p1 <- ggarrange(
    p1,
    labels = if (params$hearing_var == "pta_hf") "A" else "B",
    label.x = 0.025,
    label.y = 1.01
  )

  p1 <- p1 +
    draw_label(
      if (params$hearing_var == "pta_hf") "PTA-HF" else "PTA-LF",
      x = 0.99,
      y = 0.99,
      hjust = 1,
      vjust = 1,
      size = 9
    )

  sameplot(
    p1,
    file = here(plot_config("figure_dir"), "bioage_panel_A.tif"),
    width = params$single_column_width,
    height = 2.35,
    res = 1200,
    save = params$save_figures
  )
}
```

#### Age delta RMSE plot

```{r plot_bioage_reverse_rmse}
if (isFALSE(params$telolength)) {
  p2 <- summary_bioage_reverse %>%
    filter(!predictors %in% c("phenoage", "kdm", "hd_log")) %>%
    mutate(predictors = factor(predictors, levels = c("phenoage, age", "kdm, age", "hd_log, age"))) %>%
    ggplot(
      aes(
        x = age_group, y = delta_rmse_w
      )
    ) +
    geom_hline(yintercept = 0) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      alpha = 0.1,
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = TRUE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = age_group,
        y = ci_high_delta_rmse + 0.01,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5)
    ) +
    axes(
      x = "age_group",
      y = "drmse_reverse_pred"
    ) +
    color_scales("bioage_reverse") +
    plot_theme() +
    theme(
      axis.title.y = element_blank()
    )

  sameplot(
    p2,
    file = here(plot_config("figure_dir"), "bioage_panel_B_reverse.tif"),
    width = params$single_column_width,
    height = 2.35,
    res = 1200,
    save = FALSE
  )
}
```

#### Composite plot (paper)

```{r composite_bioage_plot}
if (isFALSE(params$telolength)) {
  p <-
    ggarrange(
      p1,
      NULL,
      p2,
      nrow = 1,
      widths = c(1, 0, 1),
      labels = c("A", "", "B"),
      label.x = 0.03,
      label.y = 1.01,
      align = "v"
    )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "bioage_panel.tif"),
    width = params$double_column_width,
    height = 2.35,
    res = 1200,
    save = params$save_figures
  )
}
```

### Additional Plots

#### Advance bioage plot (supplementary)

```{r plot_bioage_advance}
p1 <- df %>%
  ggplot() +
  geom_density_ridges(
    aes(
      x = phenoage_advance,
      y = age_group
    )
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkred"
  ) +
  axes(
    x = "age_accel_phenoage",
    y = "age_group"
  ) +
  color_scales("bioage") +
  plot_theme()

p2 <- df %>%
  ggplot() +
  geom_density_ridges(
    aes(
      x = kdm_advance,
      y = age_group
    )
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = plot_config("text_size_primary") / (14 / 5),
    color = "darkred"
  ) +
  axes(
    x = "age_accel_kdm",
    y = "age_group"
  ) +
  color_scales("bioage") +
  plot_theme()

p <- ggarrange(
  p1,
  NULL,
  p2,
  nrow = 1,
  widths = c(1, 0.1, 1),
  labels = c("A", "", "B"),
  label.x = 0.02,
  label.y = 1.01,
  align = "v"
)

sameplot(
  p,
  file = here(plot_config("figure_dir"), "age_acceleration.tif"),
  width = params$double_column_width,
  height = 3,
  res = 1200,
  save = params$save_figures
)
```

#### Biomarkers (supplementary)

```{r plot_biomarker_all, warning=FALSE}
if (isFALSE(params$telolength)) {
  update_geom_defaults("linerange", set_geom_theme("biomarker_linerange"))
  update_geom_defaults("point", set_geom_theme("biomarker_point"))

  p <- summary_biomarkers_overall %>%
    filter(!(
      predictors %in% c(
        "age"
      )
    )) %>%
    mutate(predictors = factor(
      predictors,
      levels = rev(biomarker_labels$predictors),
      labels = rev(biomarker_labels$label)
    )) %>%
    ggplot(
      aes(
        x = predictors,
        y = delta_rmse_w
      )
    ) +
    geom_point(
      aes(
        color = predictors,
        fill = predictors
      ),
      position = position_dodge(0.8),
      show.legend = FALSE
    ) +
    geom_linerange(
      aes(
        color = predictors,
        ymin = ci_low_delta_rmse,
        ymax = ci_high_delta_rmse
      ),
      position = position_dodge(0.8),
      lineend = "round",
      show.legend = FALSE
    ) +
    geom_text(
      data = . %>%
        filter(signif != "ns"),
      aes(
        x = predictors,
        y = ci_high_delta_rmse + 0.01,
        label = signif,
        color = predictors
      ),
      size = plot_config("signif_size") / (14 / 5),
      show.legend = FALSE
    ) +
    axes(
      y = "drmse_biomark"
    ) +
    plot_theme() +
    theme(
      panel.grid.major.y = element_line(color = "grey95", linewidth = 0.25),
      axis.text.x = element_markdown(
        "sans",
        size = 8,
        angle = 45,
        hjust = 1,
        vjust = 1
      ),
      axis.title.x = element_blank()
    )

  p <- ggarrange(
    p,
    labels = if (params$hearing_var == "pta_hf") "A" else "B",
    label.x = 0.015,
    label.y = 1.01
  )

  p <- p +
    draw_label(
      if (params$hearing_var == "pta_hf") "PTA-HF" else "PTA-LF",
      x = 0.5,
      y = 0.99,
      hjust = 0.5,
      vjust = 1,
      size = 9
    )

  sameplot(
    p,
    file = here(plot_config("figure_dir"), "biomarkers_panel.tif"),
    width = params$double_column_width,
    height = 2.35,
    res = 1200,
    save = params$save_figures
  )
}
```

#### Biological age correlations (supplementary)

We check the correlations between chronological and biological age and calculate Cox hazard ratios
on the differences/advances (corrected for age, sex, race, education, and smoking) and normalize to
the standard deviation so that 1 unit away means 1 standard deviation. The hazard ratios are
expressed as hazard per standard deviation away. The hazard ratios are unweighted.

```{r plot_aging_measures}
plot_bioage_correlations <- function(design, predictors, labels, predictors_advance, covariates = c("age", "sex", "race", "education", "smoking")) {
  data <- design$variables

  data <- data %>%
    mutate(
      across(
        all_of(predictors_advance),
        ~ .x / sd(.x, na.rm = TRUE)
      )
    )

  correlate <- function(p) cor(data$age, data[[p]], use = "complete.obs")
  correlations <- vapply(predictors, correlate, numeric(1))

  cox_hr <- function(p) {
    form <- as.formula(
      paste0(
        "survival::Surv(time, status) ~ ", p, " + ",
        paste(covariates, collapse = " + ")
      )
    )
    test_set <- data %>%
      filter(
        !is.na(time),
        !is.na(status),
        p != "telolength_log" | !is.na(telolength_log)
      ) %>%
      mutate(
        "{p}" := if (p == "telolength_log") {
          -.data[[p]]
        } else {
          .data[[p]]
        }
      )

    message(p, " test set size = ", nrow(test_set))

    fit <- survival::coxph(form, data = test_set)
    setNames(exp(c(coef(fit)[p], confint(fit)[p, ])), c("HR", "ci_low", "ci_high"))
  }
  hazard_ratios <- vapply(predictors_advance, cox_hr, numeric(3))

  labs <- setNames(
    sprintf(
      "<b>%s</b><br><i>r</i> = %.2f<br>HR = %.2f<br>HR 95%% CI [%.2f, %.2f]",
      labels,
      correlations,
      hazard_ratios["HR", ],
      hazard_ratios["ci_low", ],
      hazard_ratios["ci_high", ]
    ),
    predictors
  )

  data %>%
    pivot_longer(all_of(predictors), names_to = "metric", values_to = "value") %>%
    mutate(sex = factor(sex)) %>%
    ggplot(aes(x = age, y = value)) +
    geom_point(aes(color = sex, alpha = 0.04 * weights / mean(weights)), shape = 20) +
    geom_smooth(method = "lm", se = TRUE, color = "black", formula = y ~ x) +
    facet_wrap(~metric, scales = "free_y", labeller = as_labeller(labs), nrow = 1) +
    scale_alpha_identity() +
    scale_color_manual(
      values = c("1" = color("sunset")(2)[1], "2" = color("sunset")(2)[2]),
      labels = c("1" = "Male", "2" = "Female")
    ) +
    labs(y = "Biological Age", x = "Age (Years)") +
    theme_few() +
    theme(
      axis.text.x = element_markdown(
        "sans",
        size = 8
      ),
      axis.text.y = element_markdown(
        "sans",
        size = 9
      ),
      strip.text.x = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = 9,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = 9,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
}

sameplot(
  plot_bioage_correlations(design,
    predictors = c("kdm", "phenoage", "hd_log", "telolength_log"),
    labels = c("KDM", "PhenoAge", "HD (Log)", "Telomere Length (Log)"),
    predictors_advance = c("kdm_advance", "phenoage_advance", "hd_log", "telolength_log")
  ),
  width = params$double_column_width,
  height = 4,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Correlations.tif"
  )
)
```

#### Survey proportion representation

```{r plot_props}
plot_proportions <- function(data, y_variable, title) {
  ggplot(
    data,
    aes(x = category, y = .data[[y_variable]], group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_text(
      aes(y = .data[[y_variable]] / 2, label = signif(.data[[y_variable]], 2)),
      size = 2
    ) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = "Proportion"
    ) +
    theme_few() +
    scale_y_continuous(breaks = seq(0, 1, 0.5), labels = c("0", "0.5", "1")) +
    coord_cartesian(ylim = c(0, 1)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion_w",
    title = "Weighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted unfiltered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion_w",
    title = "Weighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Weighted filtered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives_unfiltered,
    y_variable = "proportion",
    title = "Unweighted unfiltered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted unfiltered strata proportions.tif"
  )
)

sameplot(
  plot_proportions(
    data = descriptives,
    y_variable = "proportion",
    title = "Unweighted filtered strata proportions"
  ),
  width = params$double_column_width,
  height = 6,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    "Unweighted filtered strata proportions.tif"
  )
)
```

**Proportions remain roughly the same before and after filtering. Gender skews slightly towards men
after filtering due to excluding pregnant women. Relative black people drops by \~12% after
filtering but absolute change is under 3% for any stratum.**

#### Descriptive statistics before and after filtering

We plot survey design-adjusted hearing means and variances per-stratum of covariates in the full and
filtered dataset to assess whether the filtered dataset is representative of the general population.

```{r survey_balance_check}
plot_descriptives <- function(data, title) {
  paste0("mean_w_", params$hearing_var)
  ggplot(
    data,
    aes(x = category, y = .data[[paste0("mean_w_", params$hearing_var)]], group = group)
  ) +
    geom_col(aes(fill = category, color = category)) +
    geom_errorbar(aes(ymin = .data[[paste0("mean_w_", params$hearing_var)]] - .data[[paste0("sd_w_", params$hearing_var)]], ymax = .data[[paste0("mean_w_", params$hearing_var)]] + .data[[paste0("sd_w_", params$hearing_var)]]), width = 0.2) +
    facet_grid(age_group ~ group, scales = "free_x", space = "free_x") +
    labs(
      title = title,
      y = paste0(if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "\n (dB HL)")
    ) +
    theme_few() +
    coord_cartesian(ylim = c(0, 80)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      axis.title.x = element_blank()
    )
}

sameplot(
  plot_descriptives(
    data = descriptives_unfiltered,
    title = paste0("Mean Weighted Unfiltered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted unfiltered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".tif")
  )
)

sameplot(
  plot_descriptives(
    data = descriptives,
    title = paste0("Mean Weighted Filtered ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA")
  ),
  width = params$double_column_width,
  height = 8,
  res = 1200,
  save = params$save_figures,
  file = here(
    "output",
    "figures",
    if (isTRUE(params$telolength)) {
      paste0(params$hearing_var, "_telolength_log")
    } else {
      params$hearing_var
    },
    "Extras",
    paste0("Weighted filtered strata ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", ".tif")
  )
)
```

#### Assessment of selection bias

```{r check_selection_bias}
plot_selection_bias <- function(ref, test) {
  ref <- ref %>% filter(group == "Overall")
  test <- test %>% filter(group == "Overall")

  diff <- ref[[paste0("mean_w_", params$hearing_var)]] - test[[paste0("mean_w_", params$hearing_var)]]

  ref %>%
    mutate(
      delta = diff
    ) %>%
    filter(
      age_group != "Overall"
    ) %>%
    ggplot(aes(x = age_group, y = delta)) +
    geom_point(aes(color = age_group), size = 4, show.legend = FALSE) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_few() +
    labs(x = "Age (Years)", y = paste0("Underestimation ", if (params$hearing_var == "pta_hf") "HF-PTA" else "LF-PTA", "", "\n(dB HL)"))
}

sameplot(
  plot_selection_bias(
    ref = descriptives_unfiltered,
    test = descriptives
  ),
  width = params$double_column_width,
  height = 4,
  res = 1200
)
```

**Selection bias is minimal in both LF-PTA and HF-PTA.**
